### 2021-09-03 src/data.prep.Rmd
initial commit

---
title: "data.prep"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tidyverse)
library(rprojroot)

# This is the project path
path <- find_rstudio_root_file()
```

The internet archive data is saved on /hdd

```{r}
data.folder <- "/hdd/internet_archive/archive/data/"
data.path <- paste0(data.folder, "all.csv")

all <- fread(data.path)
#glimpse(all)
unique(all$year)

all$id <- seq.int(nrow(all))

# for test at the end
test.all <- dim(all[1])

# data.path_ <- paste0(data.folder, "all_sample.csv")
# all_ <-fread(data.path_)
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- fread(path.lookup)
#glimpse(lookup)
```
Merge

```{r}
setkey(lookup, pcds)
setkey(all, pc)
all.msoa <- lookup[all]

sapply(all.msoa, function(x) sum(is.na(x)))
```

The next snippet drops URLS from the same host referring to same postcode, which
were observed more than once per year. 
In other words, it drops URLs, which were archived multiple times per year.

```{r}
all.msoa <- unique(all.msoa, by = c("host", "year", "pcds"))
```

Aggregate

```{r}
msoa <- all.msoa[, .(count.url=.N), by = .(year, msoa11cd)]
```

Test to see if i have missed any of the original urls

```{r}
test.msoa <- sum(msoa$count.url)
ifelse(test.all==test.msoa, yes = "OK", no = "problem")

missing <- msoa[is.na(msoa$msoa11cd)]
print(sum(missing$count.url)) #
```

Drop the observations without MSOA code

```{r}
msoa <- msoa[!is.na(msoa11cd)]
```

Export

```{r}
path.out <- paste0(path, "/data/temp/msoa.csv")
fwrite(msoa, path.out) # for all obs
```