### 2022-04-13 src/data.prep.2011.2012.Rmd
update

---
title: "data.prep.2011-2012"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(urltools)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, msoa11cd) %>% 
  dplyr::rename(pc = pcds)
  # the below is for a 'fuzzy' pc match for missing pc
  #mutate(pc_1 = str_sub(pc, start = 1, end = nchar(pc)-1))

#glimpse(lookup)
#problems(lookup)
#lookup[400296:400300,]
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd

# old pc file to fill a few hundred NA
old.pc.path <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
old.pc <- read_csv(old.pc.path) %>% 
  dplyr::select(pcds, msoa11) %>% 
  rename(msoa11.help = msoa11)

```

Download the 2011-2012 data.

**TODO** move to second drive.

```{r eval=F}
pre <- "https://www.webarchive.org.uk/datasets/ukwa.ds.2/geo/2011-201304/arcs/part-"
post <- ".bz2"  
numbers <- 0:49 
numbers <- ifelse(numbers < 10, paste0("0000", numbers), paste0("000", numbers))
urls <- lapply(numbers, function(x) paste0(pre, numbers, post))
urls <- urls[1]

for (url in urls) {
    #download.file(url, destfile = basename(url), method="curl", extra="-k")
    download.file(url, destfile = paste0(path,"/data/temp/", basename(url)))
}
```

Loop over the 50 .bz2 files of raw data for 2011-2012 and create a tibble.

```{r}

#bz2.path <- paste0(path,"/data/temp/part-00000.bz2")

all2011_12 <- tibble()

files <- list.files(path=paste0(path,"/data/temp"), pattern="*.bz2", full.names=TRUE, recursive=FALSE)

# files <-c("/home/nw19521/projects/web.diffusion/data/temp/part-00000.bz2",
#            "/home/nw19521/projects/web.diffusion/data/temp/part-00001.bz2")
# 
# x <- files[2]

#lapply(files, function(x) {

for (x in files){
  df <- read_delim(x, delim = "\t", col_names = F) %>% 
    rename(url = X1,
         pc = X2) %>%
    mutate(year = str_sub(url, start = 1, end = 4),
           newurl = substring(url,16),
           domain.ext = domain(newurl)) %>% 
    #suffix = suffix_extract(domain.ext)[,4])
    left_join(lookup, by = "pc", suffix =c("","")) %>% 
    left_join(old.pc, by=c("pc" = "pcds")) %>% 
    mutate(msoa11cd = ifelse(is.na(msoa11cd), msoa11.help,
                             msoa11cd)) %>% 
    select(-msoa11.help) %>% 
    filter(!is.na(msoa11cd))

  help <- suffix_extract(df$domain.ext) %>% 
    select(host, domain, suffix)

  #file.name = paste0(path,"/data/temp/data20112012_", substr(x[1], 56, 57), ".csv")

  df <- df %>% 
    bind_cols(help) %>% 
    select(-newurl, -domain.ext) %>% 
    filter(suffix == "co.uk")  # keep the .co.uk
    #write_csv(file = file.name)

  all2011_12 <- all2011_12 %>% bind_rows(df)

  rm(help)
 }

path.all2011_12 <- paste0(path, "/data/temp/all2011_2012.csv")
write_csv(all2011_12, path.all2011_12)
```
Create PC frequencies and bring them in all2011_2021.

As per the `loop_read_bb_new.R`:

`f <- a[, uniqueN(pc), by = .(host,year)][order(-V1)]`
`f.old <- a[, .N, by = .(host,year)][order(-N)] # old f`

I need to replicate V1 as this is what I have been using in other papers too.

```{r}
# # help code
# f <- test %>% 
#   group_by(domain.ext, year) %>%
#   summarise(V1 = n_distinct(pc))
# 
# f.old <- test %>% 
#   group_by(domain.ext, year) %>%
#   summarise(N = n())
# # equivalent to:
# test %>% count(domain.ext, year)
# 
# left_join(test, f, by = c("domain.ext" = "domain.ext",
#                           "year" = "year"))

f <- all2011_12 %>% 
  group_by(host, year) %>%
  summarise(V1 = n_distinct(pc))

all2011_12 <- all2011_12 %>% 
  left_join(f, by = c("host" = "host",
                      "year" = "year"))
#sapply(all2011_12, function(x) sum(is.na(x)))
```
Aggregate

```{r}
all2011_12 <- all2011_12 %>%

  # strips out the timestamp from the url
  mutate(url.no.date = substring(url, 16)) %>%

  # OLD 
  # Drops URLS from the same host referring to same postcode, which
  # are observed more than once per year. 
  # In other words, it drops URLs, which were archived multiple times per year.
  #distinct(host, pc,  .keep_all = TRUE) %>%

  # Drops URLs (no timestamp) referring to same postcode, which
  # are observed more than once per year. 
  # In other words, it drops pairs of URLs and postcodes, which were archived multiple times per year.
  distinct(url.no.date, pc, year, .keep_all = TRUE) %>% 
  # from 162,716,272 to 106,755,879 rows

  # Aggregate:
  # Counts how many unique pairs of url.no.date and pc exist in every MSOA per year.
  # In other words, it provides the number of georeferenced archived webpages per MSOA and year
  count(msoa11cd, year) %>%

  # Drops 2013 as we only have data for the 1st quarter
  filter(year!=2013)

# exports one file per year
all2011_12 %>% 
  group_by(year) %>%
  group_map(~ write_csv(.x, paste0(path,"/data/temp/msoa", .y$year, ".csv")))
```

Test with other PC files from https://geoportal.statistics.gov.uk/.
None of the below decreases NA substantially. The old file adds 
helps a bit, so I use this above.

```{r eval=F}
# I can only find a few more hunderd PC from the 46k missing.
old.pc.path <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
old.pc <- read_csv(old.pc.path)
missing_ <- left_join(missing, old.pc, by=c("pc" = "pcds"))
sapply(missing_, function(x) sum(is.na(x)))

# New pc file
#https://geoportal.statistics.gov.uk/datasets/national-statistics-postcode-lookup-august-2011/about
new.pc.file <- "/home/nw19521/projects/web.diffusion/data/raw/NSPL_AUG_2011_UK/Data/NSPL_AUG_2011.csv"
new.pc <- read_csv(new.pc.file, col_names = F)
missing_ <- left_join(missing, new.pc, by=c("pc" = "X1"))
sapply(missing_, function(x) sum(is.na(x)))

# NHS pc file
nhs.pc.file <- "/home/nw19521/projects/web.diffusion/data/raw/NHSPD_UK_FULL/Data/nhg21aug.csv"
nhs.pc <- read_csv(nhs.pc.file, col_names = F)
missing_ <- left_join(missing, nhs.pc, by=c("pc" = "X2"))
sapply(missing_, function(x) sum(is.na(x)))
```