### 2022-04-13 src/data.prep.Rmd

---
title: "data.prep"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)

# This is the project path
path <- find_rstudio_root_file()
```

The internet archive data is saved on /hdd.
Split to yearly files.

```{r}
data.folder <- "/hdd/internet_archive/archive/data/"
data.path <- paste0(data.folder, "all.csv")

read_csv(data.path) %>% 
  #distinct(host, year)
  group_by(year) %>%
  #group_walk(~ write_csv(.x, paste0(.y$year, ".csv"))) #
  group_map(~ write_csv(.x, paste0(path,"/data/temp/msoa", .y$year, ".csv")))
```
Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, msoa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

Bring MSOA code and then aggregate per MSOA.

```{r}

# all <- read_csv(data.path)
# all %>% 
#   #mutate(id = 1:n()) %>% #seq.int(nrow())) %>% 
#   #mutate(id = id(.)) %>% 
#   left_join(lookup, by = "pc", suffix =c("","")) %>% 
#   distinct(host, year, pc, .keep_all = TRUE) %>% 
#   group_by(msoa11cd) %>% 
#   summarise(count.url=sum(N))
#   write_csv(paste0(path,"/data/temp/data", "X", ".csv"))

files <- list.files(path=paste0(path,"/data/temp"), pattern="*.csv", full.names=TRUE, recursive=FALSE)
lapply(files, function(x) {
  read_csv(x, col_names = T) %>% 
    #mutate(id = id(.)) %>%  
    left_join(lookup, by = "pc", suffix =c("","")) %>%

    # strips out the timestamp from the url
    mutate(url.no.date = substring(url, 16)) %>%

    # OLD
    # Drops URLS from the same host referring to same postcode, which
    # are observed more than once per year. 
    # In other words, it drops URLs, which were archived multiple times per year.
    #distinct(host, pc,  .keep_all = TRUE) %>%

    # Drops URLs (no timestamp) referring to same postcode, which
    # are observed more than once per year. 
    # In other words, it drops pairs of URLs and postcodes, which were archived multiple times per year.
    distinct(url.no.date, pc, .keep_all = TRUE) %>%

    # OLD Aggregate
    #group_by(msoa11cd) %>% 
    #summarise(count.url=sum(N)) %>%

    # Aggregate:
    # Counts how many unique pairs of url.no.date and pc exist in every MSOA per year.
    # In other words, it provides the number of georeferenced archived webpages per MSOA and year
    count(msoa11cd) %>%

    write_csv(x)
})

# for test at the end
#test.all <- dim(all[1])

# data.path_ <- paste0(data.folder, "all_sample.csv")
# all_ <-fread(data.path_)
```

Test to see if i have missed any of the original urls

```{r}
test.msoa <- sum(msoa$count.url)
ifelse(test.all==test.msoa, yes = "OK", no = "problem")

missing <- msoa[is.na(msoa$msoa11cd)]
print(sum(missing$count.url)) #
```