### 2022-04-13 src/domain1112.Rmd

---
title: "domain1112"
output: html_document
---

The following creates a file with a unique pc per domain per year. 
The `all.csv` was based on all pc per host per year.
Here, I am using the domain, which in essense is the 3rd level DN.

It is based on the `/archive/data/domain.r` but for the last 2011-2012 data.

```{r setup, include=FALSE}
library(data.table)
library(kableExtra)
library(rprojroot)

path <- find_rstudio_root_file()
```

```{r}

data.path2011_2012 <- paste0(path, "/data/temp/all2011_2012.csv")
#add comments
all1112 <- read_csv(data.path2011_2012)
all1112 <- as.data.table(all1112)

f <- all1112[, uniqueN(pc), by = .(domain,year)][order(-V1)]

setkey(all1112, domain, year) # table, column
setkey(f, domain, year) #

# perform the join, eliminating not matched rows from Right
dim(all1112)
dim(f)
all1112<-all1112[f] #, nomatch=0

dim(all1112)
setnames(all1112, old=("V1"), new=("V1.domain"))

# unique rows: 1 row per domain, postcode and year. No URLs
all1112.unique <- unique(all1112, by = c("domain", "pc", "year"))
dim(all1112.unique)
all1112.unique$url <- NULL
all1112.unique$host <-NULL
all1112.unique$V1 <-NULL
all1112.unique$N <- NULL
all1112.unique$msoa11cd <- NULL
all1112.unique$suffix <- NULL

# export data
path.out <- paste0(path, "/data/temp/domain_pc_year1112.csv")
fwrite(all1112.unique, path.out)
```