### 2022-04-13 src/s_la.Rmd

---
title: "s.function"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(rprojroot)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    V1.domain < m,             # for multiple pc
    #V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  summarise(n = n()) %>%            # for multiple pc  
  #summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## Help, not really used

```{r, eval=FALSE}

df.test <- df %>% filter(oa11cd == "E00000007" |
                         oa11cd == "E00000021" |
                         oa11cd == "E00000025") %>% 
  mutate(rownum = row_number(),
         value2 = n/max(n))

# function to select unique groups
sample_n_groups = function(grouped_df, size, replace = FALSE, weight=NULL) {
  grp_var <- grouped_df %>% 
    groups %>%
    unlist %>% 
    as.character
  random_grp <- grouped_df %>% 
    summarise() %>% 
    sample_n(size, replace, weight) %>% 
    mutate(unique_id = 1:NROW(.))
  grouped_df %>% 
    right_join(random_grp, by=grp_var) %>% 
    group_by_(grp_var) 
}

df.test <- df %>% group_by(oa11cd) %>% 
  sample_n_groups(10000) %>% 
  arrange(oa11cd, year) %>% 
  select(-unique_id)

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)

fd.nls <- df.test %>% 
  group_by(oa11cd) %>%
  select(year, oa11cd, n) %>% 
  do(fitloop = tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .))) %>% 
  unnest(fitloop)

fd.nls %>% filter(term=="xmid") %>% 
  summarise(rng = range(estimate))
```

## Test for S function estimation

To interpret the coefficients, check [SSlogis](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SSlogis).
This is a self-starting model, which estimates the starting values.

```{r}
df.test <- df %>% filter(ladcd == "S12000033")

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
summary(fit)

df.test$prednls = predict(fit)

ggplot(df.test, aes(x = year, y = n) ) +
  geom_point() +
  geom_line(aes(y = prednls), size = 1) +
  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
  ylab("N. of websites")
  #scale_x_date(breaks = df.test$year) 
```

## S function estimation for all the country

```{r}
df.country <- df %>% group_by(year) %>% 
  summarise(n = sum(n))

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.country)
summary(fit)

df.country$prednls = predict(fit)

ggplot(df.country, aes(x = year, y = n) ) +
  geom_point() +
  geom_line(aes(y = prednls), size = 1) +
  scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
  ylab("N. of websites") +
  ggtitle("Country")
  #scale_x_date(breaks = df.test$year) 
```

## Loop for S function estimation for Local Authorities

```{r}
a.la <- as_tibble()
all.la <- df %>% filter(ladcd != "NA") %>% distinct(ladcd) %>% simplify2array()
simplify2array(all.la)

for (i in all.la){
  data <- df %>% filter(ladcd==i) %>%
    select(year, ladcd, n) 
    #lm(n~year, data = .)
    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
  tryCatch(b.la <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
  tryCatch(a.la <-rbind(a.la,b.la), error=function(e) NULL)
}

a.la %>% filter(term=="xmid") %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  select(estimate) %>% 
  summarise(r = range(estimate),
            median = median(estimate))

a.la %>% filter(term=="xmid") %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
  select(estimate) %>% 
  ggplot(aes(x=estimate)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 2003, colour = "red")
  scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) 
  #scale_x_discrete(labels = as.factor(a$estimate))

a.la <- a.la %>% filter(term=="xmid") %>% 
  bind_cols(as.tibble(all.la)) %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  filter(p.value < 0.01) %>%    # 376 / 376, all sig
  select(ladcd, estimate)
  #filter(estimate>2011)
```

write the object with all the models

```{r eval=F}
path.out <- paste0(path, "/data/temp/s_la.csv")
write_csv(a.la, path.out)
```

## Map the xmid

Spatial data

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))

la.f <- fortify(la, region = "LAD21CD")

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
```

```{r}

la.f <- la.f %>% left_join(a.la, by = c("id" = "ladcd")) %>% 
  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps

# oa.gb.f %>% mutate(fast = ifelse(estimate < 2003, 1, 0)) %>% 
#   arrange(order) %>% 
ggplot(data = la.f, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = estimate)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  scale_fill_continuous(type = "viridis") +
  theme_void()

ggplot(data = la.f, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = fast)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
    geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
  #scale_fill_continuous(type = "viridis") +
  theme_void()

```

```{r}
la.f %>% left_join(df %>% filter(year==2012), by = c("id" = "ladcd")) %>% 
  arrange(order) %>% # if i don't order merge.nuts.f loses order and the map has gaps
  ggplot(aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = n)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  scale_fill_continuous(type = "viridis") +
  theme_void()

```