### 2022-05-26 paper/paper.Rmd
papers

---
title: Title of submission to PLOS journal
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliation: University of Bristol and The Alan Turing Institute
    corresponding: e.tranos@bristol.ac.uk
address:
  - code: University of Bristol and The Alan Turing Institute
    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
abstract: |
  blah blah

author_summary: |
  blah blah

bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
header-includes:
- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
---

```{r echo=FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
#library(rmarkdown)
```

# Introduction {#sec:1}

@beardsell1999spatial
To characterize evolution of the computer industry, we examine the distribution
of relative employment across cities in 1977 and how that distribution
changes over time to 1992

There is no
tendency of relative size distributions of urban computer employment to collapse,
go bimodal, or fully spread. Overall computers exhibit some turbulence,
with dramatic big winners and losers among cities. In attracting or repelling an
industry, urban heterogeneity is important. Large, well educated cities near San
Jose have much greater chances of attracting high-tech employment and less of
losing it

@bednarz2020pulled
spatial innovation diffusion
By using Bayesian survival models with time-dependent data of wind turbine deployment
and firm foundation for 402 German regions between the years 1970 and 2015, we show that the spatial
evolution of the German wind energy industry was more strongly influenced by local demand–
pull than local supply–push processes.

The industry’s initial locations are distributed relatively arbitrarily
and unpredictably, as their needs in terms of resources and skills are diverse and distinct from the older existing
industries (Boschma and Lambooy, 1999). Consequently, emerging industries are characterized by relatively high
degrees of freedom in terms of location. In later extensions of the concept, the assumption of the randomness of locations
was revised with greater importance assigned to regional conditions (Boschma and Lambooy, 1999; Fornahl
et al., 2012).

larger local technological niches (at least from a demand perspective) are
more likely to be formed in urban areas than in rural ones, which is in line with the work of Ha¨gerstrand (1965a,b).
However, size is not the only regional characteristic that may matter in this context. Regions also differ in accumulated
experiences and the presence of tacit knowledge with respect to products, as well as in actors’ propensity share
this knowledge among producers (Martin et al., 2019)

@bento2015spatial
The concept of absorptive capacity has also been applied in the TIS context in terms of the con-struction of capacity needed to transfer and implement a new technology in the receiving country (cf.Van Alphen, 2011), with repercussions for the nature of system functions

@das2022diffusion [not good paper]
Invention, innovation
and diffusion are considered as the three pillars of
the theory of technological change and Schumpeter is considered
as the father of the study of diffusion

Epidemic approach: The approach emphasizes that the adoption of a technology
spreads like an infection among the potential adopters.
By taking a homogeneous group of potential adopters as
an assumption, it explains the adoption of any innovation.

To summarize
the discussion, first, here the diffusion process is selfperpetuating
as initial use stimulates further use. Second, it
follows a disequilibrium path as the level of users is
always lower than the number of potential users (Stoneman
2002).

The
criticisms are mainly based on the fact that although the
approach gives an idea of aggregate (industry or household)
behaviour, it does not focus on the individual’s
(firm or household) adoption process

Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) Early adopters: next
13.5%, (III.) Early Majority: following 34%, (IV.) Late Majority: next 34%, and the rest as (V.) Laggards

@morrill2020spatial
Spatial diffusion is the process by which behavior or characteristics of the landscape change as a result of
what happens elsewhere earlier. Spatial diffusion is the spread of the phenomenon, over space and timed,
from limited origins

Contagious diffusion=> spatial autocorelation

expansion-type diffusion

the three geographic diffusion processes are (1) purely contagious where distance or adjacency
is the absolute controlling factor, (2) purely hierarchical where size or urban position in the central place
hierarchy is the absolute controlling factor, or (3) where the location of change is purely random. These are
the pure or ideal forms of the diffusion process; all real diffusion processes are the result of a combination of
these extremes.

The work by the communication sociologist Everett Rogers (1962, 1971, 1983) has emphasized the role of
information, communication, formal and informal media, opinion leaders and social networks, and economic
and psychological constraints on acceptance. Rogers’s work stresses the decision mechanism of the potential
adopter; the work of Rogers partitions the adopter’s process of choosing to accept a new phenomenon or
trait into five stages: stage 1–the potential adopter gains knowledge or awareness of an innovation; stage
2–persuasion is exercised to adopt; stage 3–a decision is made to adopt; stage 4–the decision to adopt is
implemented; stage 5–the adopter confirms the decision to adopt. In Rogers’s stage theory there may be a
substantial time lag between when a potential adopter becomes aware of the new characteristic and when a
decision is actually made to adopt.

Hägerstrand considered diffusion to be a fundamental geographic process: Whatever the phenomenon being
diffused might be, one may consider it in the context of a larger universal process of spatial diffusion.

@zook2022mapping
Because new digital activities are rarely—if
ever—captured in official state data, researchers
must rely on information gathered from alternative
sources. With this in mind, the overarching aim of
this article is outlining an approach for analyzing the
“new spaces and geographies” of digital phenomena
and practices.

@leibowicz2016representing for energy
There is no generally accepted theory that explains diffusion
rate heterogeneity across technologies, but several factors are
considered important. Greater unit scale and larger market size
contribute to slower diffusion. Requirements for interrelated
technologies or complex infrastructures also hinder the diffusion
process (Grubler, 2012).

Mobile phones benefited from early
deployment in recreational boats and automobiles, where the
traditional competitor was not a viable option. In the early
stages of diffusion, performance is a more important driver of
adoption than cost competitiveness. Typically, significant cost
reductions only occur once the technology reaches a deployment
level capable of supporting standardization and mass
production (Wilson, 2012).

Diffusion is a spatial (Hagerstrand, 1967) as well as a
temporal process, and historical evidence confirms that
technologies diffuse at different times, at different rates, and
to different extents in different places, and can be significantly
influenced by policies (Victor, 1993). An illustrative empirically
based theory for technology diffusion acrossmultiple regions is
suggested by “Schmidt's Law” (Grubler, 1990). With respect to a
specific technology, it divides regions into three groups: core,
rim, and periphery. The core iswhere the technology is invented,
subject to experimentation, and first deployed commercially.
To develop the technology, the core must have both sufficient
innovative ability and economic motivation to make the
necessary resource commitment. As the technology spreads
from the core to the rim and eventually to the periphery, later
adopters have the advantage of benefiting from the experiences
of earlier adopters via knowledge spillover effects (Verdolini and
Galeotti, 2011). At this point, the technology has likely been
refined and costs have likely declined significantly. Schmidt's
Law claims that, as a result, diffusion proceeds at a faster rate in
regions which adopt the technology later. However, diffusion
tends to be less pervasive in regions where adoption is delayed,
ultimately saturating at a lower extent than in regions where
adoption occurred earlier. This is often attributed to a lack of
proper institutions and infrastructures in later adopters, markets
whichmay not be as favorable to a technology, or the availability
of yet newer options that invite technological leapfrogging.

Empirical evidence supports the validity of Schmidt's Law
over a wide range of technologies, time periods, and geographical
contexts. A recent meta-analysis of technology up-scaling
found that diffusion accelerated moving from the core to the
rim and periphery for technologies as diverse as natural gas
power, oil refineries, and automobiles (Wilson, 2009). One
historical example that conforms particularly well to Schmidt's
Law is the diffusion of coal power in Europe (Grubler, 2012).
England emerged as the core region for coal power because it
had legal and economic institutions that incentivized scientific
pursuits, domestic coal reserves, and a clear industrial motivation
to replace water power with coal. 
...

Lastly, it can be globally optimal for innovative economies
to deploy advanced technologies more than what is locally
optimal if this enables faster diffusion in less advanced regions
that could benefit from the technology.

@wilson201281
Logistic growth describes an initial period of gradual diffusion as a technology is introduced as a new commercial application, moving then through a rapid, exponential growth phase, before slowing and eventually saturating [@grubler1999dynamics]. The substitution of incumbent technologies by new competitors leads to subsequent decline and eventual obsolescence.

Early on in their lifecycle, new technologies are crude, imperfect,
and expensive (Rosenberg, 1994). New energy technologies
are attractive for their ability to perform a particular task or
deliver a new or improved energy service (Fouquet, 2010). This is
often circumscribed by a particular set of needs in a particular
context: a market ‘niche’. End-users in niche markets are generally
less sensitive to the effective price of the energy service
provided or have a higher willingness to pay for its performance
advantages (Fouquet, 2010). Thus initially, performance dominates
cost competitiveness (Wilson and Grubler, 2011). Market
niches afford some protection from competitive pressures, allowing
technologies to be tested and improved in applied settings,
reducing uncertainties with performance or market demand
(Kemp et al., 1998). Costs may only fall substantively after an
extended period of commercial experimentation, concurrent with
the establishment of an industrial base and characteristic moves
towards standardisation and mass production (Grubler, 1998).
The influence of accumulating production experience on costs is
captured by the concept of learning.

@grubler1990rise
Following the definition given by Rogers, 1983, diffusion is the
process by which an innovation is communicated through certain communication
channels over time among the members of a social system.
An innovation may be an idea, object or practice which is perceived
as new by an individual or another unit of adoption (e.g., an
organization) .
Communication via communication channels refers to how
knowledge about an innovation is transferred between the members of
a social system. Communication takes time, and is therefore an
important aspect of innovation diffusion. In fact, all theoretical and
empirical diffusion studies agree that an innovation does not instantly
diffuse into a social system. Instead a typical time pattern of diffusion
along a S-shaped trajectory seems to be the rule. The S-shaped pattern
of diffusion appears to be a basic anthropologic phenomenon, as
it is also confirmed by a large number of studies· of primitive preindustrial
societies.

The scale of spatial innovation diffusion is similar* to that of
temporal diffusion models: hierarchically decomposed.

Spatial innovation diffusion in Hagerstrand's model consists of
three stages.
Stage 1: Local concentrations of initial acceptances (initial
agglomeration) .
Stage 2: Radial dissemination outward from the initial agglomeration,
accompanied by the rise of secondary agglomerations.
Stage 3: The condensing and saturation stage, in which filling-in
occurs and diffusion eventually ceases.

Thus, three main characteristics define the spatial diffusion process:
(1) As in temporal diffusion models, an S-shaped pattern in the
cumulative level of adoption.
(2) A "hierarchy effect" by which diffusion proceeds from initial
centers to secondary ones. Within the context of urban systems
(central places) this implies a hierarchical spread of an innovation
from larger to smaller centers, and from the centers to the surrounding
hinterland.
(3) A "neighborhood effect" Le., diffusion proceeds outwards from
innovation centers, first "hitting" (Brown, 1981) nearby rather
than far-away locations.

Later Hagerstrand conceptualized
physical "barrier" effects like lakes or uninhabited areas, which, in
addition to distance, act as further retarding effects on diffusion.
These are formalized in the form of "zero" or "half" contact multiplicators
on the (distance decaying) message flows.

With respect to the formalization of the communication flows
Hagerstrand defines a "mean information field" (MIF), in which the
probability of communication is a negative function of distance
between individuals

@ding2010modeling model the spatial diffusion of mobile telecommunications in China
as well as its determinants. regional socioeconomic characteristics play an
important role in determining the timing,
speed, and level of mobile telecommunications
diffusion in China

Rogers
(1995) characterized early adopters as knowledgeable
risk takers. Griliches (1957) looked
at the rate of return for early adopters and
characterized them as profit maximizers. The
geographer Hagerstrand (1967) was focusing
on the character of early adopters in the transmittal
process that produces follow-on participants
in the process. Space was treated as a
contiguity system with embedded characteristics
producing barrier elements to the diffusion
process.

REVIEW ON Technological Diffusion

Perkins and Neumayer (2005)
examined whether latecomers have a higher diffusion
rate of three technologies including digital
telephone mainline

@bento2018time explore 
What determines the duration of
formative phases for energy innovations in different markets? We are interested
both in initial markets (also: core, lead, first mover, early
adopter) where formative phases prepare technologies for mass commercialization,
and in follower markets (also: periphery, lag, late
adopter) where accelerated formative phases may benefit from diffusion
and spillovers.

Hagerstrand (1968) was the first
to demonstrate that diffusion typically starts from innovative centers
(‘core´) and disseminates, through a hierarchy of subcenters, to the
periphery. This process is sequential rather than simultaneous, and
tends to accelerate from core to periphery, where diffusion reaches a
lower intensity than in the core (Hagerstrand, 1968; Grubler, 1990;
Morrill, 2005). A similar effect of spatial acceleration has been identified
in marketing research and is known as the “lead-lag effect” (Peres
et al., 2010).

@perkins2005international analyze whether the rate at which new producer
technologies diffuse is significantly influenced by
(1) latecomer advantage and (2) engagement with the
global economy via trade and foreign investment

Indeed, precisely
because of these latecomer advantages, developing
countries1 are believed to be well placed to catch up with
developed ones (Gerschenkron 1962; Abramovitz
1986).

Geographers, of course, were at the forefront of
quantitative diffusion research. Ha¨gerstrand’s (1967)
early work on understanding innovation diffusion in
Sweden using simulation modeling is particularly noteworthy
in this respect. However, already long before the
‘‘cultural turn’’ in economic geography (Barnes 2001),
geographers’ interest in formally modeling the diffusion
of new innovations waned (Gregory 1985). Indeed,
much of the recent quantitative research into the diffusion
of new technology has been undertaken by
economists, sociologists, and business scholars.

Theoretical models disagree as to why firms adopt
innovations at different times. Epidemic models emphasize
information (Griliches 1957). Certain firms are hypothesized
to adopt earlier because they come into
contact with, and learn from, adopters of the new technology
before others. Economic models, on the other
hand, predominantly emphasize firm heterogeneity (Ireland
and Stoneman 1986). Firms adopt technologies at
different times because they differ with respect to various
organizational and environmental variables influencing
the economic returns from adoption (Blackman 1999).

Either
way, the strong assumption is that developing countries
can acquire modern technology innovated in developed
economies, often at a fraction of the original research
and development (R&D) costs, thereby leapfrogging
many decades of technological progress (Teece 2000).

The first involves a country’s geographical location.
Recent empirical work suggests that diffusion is ‘‘geographically
localized’’ (Globerman, Kokko, and Sjo¨holm
2000; Keller 2002; Milner 2003) in that a technology
diffuses faster in a country where it is already more
widely diffused in neighboring countries. Underlying
these regional effects are contagion and contact with
prior users or producers of technology.

@perkins2011internet
Two main mechanisms are
identified in the literature to explain diffusion:
(1) epidemic-type dynamics whereby contact
with previous adopters stimulates uptake as po
tential
adopters learn about a new innovation;
and (2) economic-type mechanisms whereby
potential users adopt a new innovation as it be
comesmore
profitable, useful, or valuable, with
uptake characteristically spreading as costs be
come
lower, performance improves, or the po
tential
uses of the innovation grow over time.

To this extent,
our results contribute to a growing body of
work that has sought to caution against claims
about the supposed novelty of the Internet and
the suggestion that it is somehow different.

@meade2021modelling
There are numerous agents who would benefit from accurate predictions of introduction times.
These include: suppliers of network equipment for production planning; suppliers of mobile phones who need to have inventories of
appropriately configured handsets; international network planners who need to provide sufficient bandwidth for communications
between different countries; network operators and regulatory authorities.

Classifying adopters of a technology along a time axis was pioneered by
Rogers (1962), who categorized adopters into five groups as innovators, early adopters, early majority, late majority and laggards.

An a priori approach groups similar countries into segments, where similarity is defined using economic, social, or
political covariates. Examples include Hofstede (2001) and Lee (1990). An a posteriori analysis considers segments based on realized
market behaviour, for example see Sood et al. (2009). Some studies investigate changing segment membership over time, see Cannon
and Yaprak (2011).

@fritsch2015new

@fritsch2015new analyze the spatial diffusion of laser technology research in West Germany from
1960, when this technology began, until 2005.

Hägerstrand (1952, 1967) was the first scholar to intensively analyze spatial diffusion pro
cesses.
Based on the empirical observations of the regional spread of process innovations,
he hypothesized that new technology is first implemented in the ‘center’ – the large agglom
erations
– and then moves ‘down’ the spatial hierarchy, the last stop being the ‘periphery’,
i.e. remote and sparsely populated regions. There are at least two factors that may explain
this pattern: first, the transfer of tacit knowledge via face-to-face contact and second, the
absorptive capacity of a region.

# Literature review {#sec2} MAYBE SKIP THIS??

# Materials and Methods {#sec3}

@wilson201281 growth function description in p. 86

from R: Asym/(1+exp((xmid-input)/scal))
using the terms from Wilson k/(1 + e^(t_0 - t)/scal)

Wilson: k(1 + e^-b(t - t_0))
So, scal = - 1/b

The curve is symmetric* around t_0

**minimum R2 = 95%**, see also @grubler1990rise

The literature usually uses the saturation level as the asymptote. I am using the total
number of websites as we cannot compute a rate.

Moran's I as @ding2010modeling
**TODO** for t_0,  diffusion speed

## Multi-country analysis {#sec3.1}

## 3.2	Case study approach {#sec3.2}

# Digital technologies and spatial structure: a global view {#sec4}

# The impact of ICT on the US and the UK spatial structures {#sec5}

# Discussion and conclusions {#sec6}

contrary to results from future studies regarding social media [@lengyel2020role], web technologies 
did not exclusively spread from a central location.

@lengyel2020role
We see that the OSN
spread almost exclusively from the original location (the capital Budapest, with an order of magnitude more inhabitants than
the next size town) to various parts of the country in the early phase of the life-cycle. Later, diffusion became less mono-centric
and other towns also emerged as spreaders. Our findings support the idea that spreading initially happens to large distances and
becomes more local over time. This is illustrated and discussed later in Figure 2I.

# References {#references .unnumbered}