### 2022-05-26 presentations/gceg.Rmd

---
title: 'Digital economy in the UK: an evolutionary story'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: mybibfile.bib
biblio-style: "apalike"
css: style.css
---

```{r, echo=FALSE, results=FALSE, include=FALSE}
library(knitr)
library(rprojroot)

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size

# This is the project path
path <- find_rstudio_root_file()

#bibliography: ../bibliography.bib
```

## Contents
- Introduction
- Web data and spatial research
- Empirical strategy
- Descriptive statistics
- Results
- Conclusions
<br>
<br>
[etranos.info/post/sad2021](https://etranos.info/post/sad2021)

# Introduction

## Aims

- Map the active engagement with the *digital*
- Over time, early stages of the internet
- Granular spatial level
- Importance: understand how the adoption of technologies evolves
- Predictions of introduction times for future technologies [@meade2021modelling]

  - Network operators

  - Suppliers of network equipment

  - Regulatory authorities

## Diffusion

Large scale adoption studies

Add Figure

## Spatiality of the internet using supply-side measures

- Map internet infrastructure, *both*:

  - backbone (ADD REF), and

  - last mile (ADD REF, Singleton, Hannah)

- Economic effects:

  - Kolko

  - Nathan

## ???

@beardsell1999spatial used the employment in the computer industry to assess
its evolution

# Results

---

```{r results='asis', out.width = "60%"}

library(magick)
library(purrr)
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir

```

---

```{r results='asis', out.width = "60%"}

library(magick)
library(purrr)
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir

```

```{r eval=FALSE, fig.show='animate', ffmpeg.format='gif', dev='jpeg', interval=0.01}

hex.images <- list.files(path = path.images, pattern = "^hex", full.names = T)
for (i in hex.images){
  plot(i) # for example
}
```

## References {.allowframebreaks}