### 2023-02-16 paper/paper.Rmd
presentation

---
title: Title of submission to PLOS journal
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliation: University of Bristol and The Alan Turing Institute
    corresponding: e.tranos@bristol.ac.uk
address:
  - code: University of Bristol and The Alan Turing Institute
    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
abstract: |
  blah blah

author_summary: |
  blah blah
bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
header-includes:
- \usepackage{rotating, graphicx, dcolumn, booktabs, caption, xcolor}
editor_options: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
#library(rmarkdown)
```

# Introduction {#sec:1}

@beardsell1999spatial To characterize evolution of the computer
industry, we examine the distribution of relative employment across
cities in 1977 and how that distribution changes over time to 1992

There is no tendency of relative size distributions of urban computer
employment to collapse, go bimodal, or fully spread. Overall computers
exhibit some turbulence, with dramatic big winners and losers among
cities. In attracting or repelling an industry, urban heterogeneity is
important. Large, well educated cities near San Jose have much greater
chances of attracting high-tech employment and less of losing it

@bednarz2020pulled spatial innovation diffusion By using Bayesian
survival models with time-dependent data of wind turbine deployment and
firm foundation for 402 German regions between the years 1970 and 2015,
we show that the spatial evolution of the German wind energy industry
was more strongly influenced by local demand-- pull than local
supply--push processes.

The industry's initial locations are distributed relatively arbitrarily
and unpredictably, as their needs in terms of resources and skills are
diverse and distinct from the older existing industries (Boschma and
Lambooy, 1999). Consequently, emerging industries are characterized by
relatively high degrees of freedom in terms of location. In later
extensions of the concept, the assumption of the randomness of locations
was revised with greater importance assigned to regional conditions
(Boschma and Lambooy, 1999; Fornahl et al., 2012).

larger local technological niches (at least from a demand perspective)
are more likely to be formed in urban areas than in rural ones, which is
in line with the work of Ha¨gerstrand (1965a,b). However, size is not
the only regional characteristic that may matter in this context.
Regions also differ in accumulated experiences and the presence of tacit
knowledge with respect to products, as well as in actors' propensity
share this knowledge among producers (Martin et al., 2019)

@bento2015spatial The concept of absorptive capacity has also been
applied in the TIS context in terms of the con-struction of capacity
needed to transfer and implement a new technology in the receiving
country (cf.Van Alphen, 2011), with repercussions for the nature of
system functions

@das2022diffusion [not good paper] Invention, innovation and diffusion
are considered as the three pillars of the theory of technological
change and Schumpeter is considered as the father of the study of
diffusion

Epidemic approach: The approach emphasizes that the adoption of a
technology spreads like an infection among the potential adopters. By
taking a homogeneous group of potential adopters as an assumption, it
explains the adoption of any innovation.

To summarize the discussion, first, here the diffusion process is
selfperpetuating as initial use stimulates further use. Second, it
follows a disequilibrium path as the level of users is always lower than
the number of potential users (Stoneman 2002).

The criticisms are mainly based on the fact that although the approach
gives an idea of aggregate (industry or household) behaviour, it does
not focus on the individual's (firm or household) adoption process

Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.)
Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.)
Late Majority: next 34%, and the rest as (V.) Laggards

@morrill2020spatial Spatial diffusion is the process by which behavior
or characteristics of the landscape change as a result of what happens
elsewhere earlier. Spatial diffusion is the spread of the phenomenon,
over space and timed, from limited origins

Contagious diffusion=\> spatial autocorelation

expansion-type diffusion

the three geographic diffusion processes are (1) purely contagious where
distance or adjacency is the absolute controlling factor, (2) purely
hierarchical where size or urban position in the central place hierarchy
is the absolute controlling factor, or (3) where the location of change
is purely random. These are the pure or ideal forms of the diffusion
process; all real diffusion processes are the result of a combination of
these extremes.

The work by the communication sociologist Everett Rogers (1962, 1971,
1983) has emphasized the role of information, communication, formal and
informal media, opinion leaders and social networks, and economic and
psychological constraints on acceptance. Rogers's work stresses the
decision mechanism of the potential adopter; the work of Rogers
partitions the adopter's process of choosing to accept a new phenomenon
or trait into five stages: stage 1--the potential adopter gains
knowledge or awareness of an innovation; stage 2--persuasion is
exercised to adopt; stage 3--a decision is made to adopt; stage 4--the
decision to adopt is implemented; stage 5--the adopter confirms the
decision to adopt. In Rogers's stage theory there may be a substantial
time lag between when a potential adopter becomes aware of the new
characteristic and when a decision is actually made to adopt.

Hägerstrand considered diffusion to be a fundamental geographic process:
Whatever the phenomenon being diffused might be, one may consider it in
the context of a larger universal process of spatial diffusion.

@zook2022mapping Because new digital activities are rarely---if
ever---captured in official state data, researchers must rely on
information gathered from alternative sources. With this in mind, the
overarching aim of this article is outlining an approach for analyzing
the "new spaces and geographies" of digital phenomena and practices.

@leibowicz2016representing for energy There is no generally accepted
theory that explains diffusion rate heterogeneity across technologies,
but several factors are considered important. Greater unit scale and
larger market size contribute to slower diffusion. Requirements for
interrelated technologies or complex infrastructures also hinder the
diffusion process (Grubler, 2012).

Mobile phones benefited from early deployment in recreational boats and
automobiles, where the traditional competitor was not a viable option.
In the early stages of diffusion, performance is a more important driver
of adoption than cost competitiveness. Typically, significant cost
reductions only occur once the technology reaches a deployment level
capable of supporting standardization and mass production (Wilson,
2012).

Diffusion is a spatial (Hagerstrand, 1967) as well as a temporal
process, and historical evidence confirms that technologies diffuse at
different times, at different rates, and to different extents in
different places, and can be significantly influenced by policies
(Victor, 1993). An illustrative empirically based theory for technology
diffusion acrossmultiple regions is suggested by "Schmidt's Law"
(Grubler, 1990). With respect to a specific technology, it divides
regions into three groups: core, rim, and periphery. The core iswhere
the technology is invented, subject to experimentation, and first
deployed commercially. To develop the technology, the core must have
both sufficient innovative ability and economic motivation to make the
necessary resource commitment. As the technology spreads from the core
to the rim and eventually to the periphery, later adopters have the
advantage of benefiting from the experiences of earlier adopters via
knowledge spillover effects (Verdolini and Galeotti, 2011). At this
point, the technology has likely been refined and costs have likely
declined significantly. Schmidt's Law claims that, as a result,
diffusion proceeds at a faster rate in regions which adopt the
technology later. However, diffusion tends to be less pervasive in
regions where adoption is delayed, ultimately saturating at a lower
extent than in regions where adoption occurred earlier. This is often
attributed to a lack of proper institutions and infrastructures in later
adopters, markets whichmay not be as favorable to a technology, or the
availability of yet newer options that invite technological
leapfrogging.

Empirical evidence supports the validity of Schmidt's Law over a wide
range of technologies, time periods, and geographical contexts. A recent
meta-analysis of technology up-scaling found that diffusion accelerated
moving from the core to the rim and periphery for technologies as
diverse as natural gas power, oil refineries, and automobiles (Wilson,
2009). One historical example that conforms particularly well to
Schmidt's Law is the diffusion of coal power in Europe (Grubler, 2012).
England emerged as the core region for coal power because it had legal
and economic institutions that incentivized scientific pursuits,
domestic coal reserves, and a clear industrial motivation to replace
water power with coal. ...

Lastly, it can be globally optimal for innovative economies to deploy
advanced technologies more than what is locally optimal if this enables
faster diffusion in less advanced regions that could benefit from the
technology.

@wilson201281 Logistic growth describes an initial period of gradual
diffusion as a technology is introduced as a new commercial application,
moving then through a rapid, exponential growth phase, before slowing
and eventually saturating [@grubler1999dynamics]. The substitution of
incumbent technologies by new competitors leads to subsequent decline
and eventual obsolescence.

Early on in their lifecycle, new technologies are crude, imperfect, and
expensive (Rosenberg, 1994). New energy technologies are attractive for
their ability to perform a particular task or deliver a new or improved
energy service (Fouquet, 2010). This is often circumscribed by a
particular set of needs in a particular context: a market 'niche'.
End-users in niche markets are generally less sensitive to the effective
price of the energy service provided or have a higher willingness to pay
for its performance advantages (Fouquet, 2010). Thus initially,
performance dominates cost competitiveness (Wilson and Grubler, 2011).
Market niches afford some protection from competitive pressures,
allowing technologies to be tested and improved in applied settings,
reducing uncertainties with performance or market demand (Kemp et al.,
1998). Costs may only fall substantively after an extended period of
commercial experimentation, concurrent with the establishment of an
industrial base and characteristic moves towards standardisation and
mass production (Grubler, 1998). The influence of accumulating
production experience on costs is captured by the concept of learning.

@grubler1990rise Following the definition given by Rogers, 1983,
diffusion is the process by which an innovation is communicated through
certain communication channels over time among the members of a social
system. An innovation may be an idea, object or practice which is
perceived as new by an individual or another unit of adoption (e.g., an
organization) . Communication via communication channels refers to how
knowledge about an innovation is transferred between the members of a
social system. Communication takes time, and is therefore an important
aspect of innovation diffusion. In fact, all theoretical and empirical
diffusion studies agree that an innovation does not instantly diffuse
into a social system. Instead a typical time pattern of diffusion along
a S-shaped trajectory seems to be the rule. The S-shaped pattern of
diffusion appears to be a basic anthropologic phenomenon, as it is also
confirmed by a large number of studies· of primitive preindustrial
societies.

The scale of spatial innovation diffusion is similar\* to that of
temporal diffusion models: hierarchically decomposed.

Spatial innovation diffusion in Hagerstrand's model consists of three
stages. Stage 1: Local concentrations of initial acceptances (initial
agglomeration) . Stage 2: Radial dissemination outward from the initial
agglomeration, accompanied by the rise of secondary agglomerations.
Stage 3: The condensing and saturation stage, in which filling-in occurs
and diffusion eventually ceases.

Thus, three main characteristics define the spatial diffusion process:
(1) As in temporal diffusion models, an S-shaped pattern in the
cumulative level of adoption. (2) A "hierarchy effect" by which
diffusion proceeds from initial centers to secondary ones. Within the
context of urban systems (central places) this implies a hierarchical
spread of an innovation from larger to smaller centers, and from the
centers to the surrounding hinterland. (3) A "neighborhood effect" Le.,
diffusion proceeds outwards from innovation centers, first "hitting"
(Brown, 1981) nearby rather than far-away locations.

Later Hagerstrand conceptualized physical "barrier" effects like lakes
or uninhabited areas, which, in addition to distance, act as further
retarding effects on diffusion. These are formalized in the form of
"zero" or "half" contact multiplicators on the (distance decaying)
message flows.

With respect to the formalization of the communication flows Hagerstrand
defines a "mean information field" (MIF), in which the probability of
communication is a negative function of distance between individuals

@ding2010modeling model the spatial diffusion of mobile
telecommunications in China as well as its determinants. regional
socioeconomic characteristics play an important role in determining the
timing, speed, and level of mobile telecommunications diffusion in China

Rogers (1995) characterized early adopters as knowledgeable risk takers.
Griliches (1957) looked at the rate of return for early adopters and
characterized them as profit maximizers. The geographer Hagerstrand
(1967) was focusing on the character of early adopters in the
transmittal process that produces follow-on participants in the process.
Space was treated as a contiguity system with embedded characteristics
producing barrier elements to the diffusion process.

REVIEW ON Technological Diffusion

Perkins and Neumayer (2005) examined whether latecomers have a higher
diffusion rate of three technologies including digital telephone
mainline

@bento2018time explore What determines the duration of formative phases
for energy innovations in different markets? We are interested both in
initial markets (also: core, lead, first mover, early adopter) where
formative phases prepare technologies for mass commercialization, and in
follower markets (also: periphery, lag, late adopter) where accelerated
formative phases may benefit from diffusion and spillovers.

Hagerstrand (1968) was the first to demonstrate that diffusion typically
starts from innovative centers ('core´) and disseminates, through a
hierarchy of subcenters, to the periphery. This process is sequential
rather than simultaneous, and tends to accelerate from core to
periphery, where diffusion reaches a lower intensity than in the core
(Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of
spatial acceleration has been identified in marketing research and is
known as the "lead-lag effect" (Peres et al., 2010).

@perkins2005international analyze whether the rate at which new producer
technologies diffuse is significantly influenced by (1) latecomer
advantage and (2) engagement with the global economy via trade and
foreign investment

Indeed, precisely because of these latecomer advantages, developing
countries1 are believed to be well placed to catch up with developed
ones (Gerschenkron 1962; Abramovitz 1986).

Geographers, of course, were at the forefront of quantitative diffusion
research. Ha¨gerstrand's (1967) early work on understanding innovation
diffusion in Sweden using simulation modeling is particularly noteworthy
in this respect. However, already long before the ''cultural turn'' in
economic geography (Barnes 2001), geographers' interest in formally
modeling the diffusion of new innovations waned (Gregory 1985). Indeed,
much of the recent quantitative research into the diffusion of new
technology has been undertaken by economists, sociologists, and business
scholars.

Theoretical models disagree as to why firms adopt innovations at
different times. Epidemic models emphasize information (Griliches 1957).
Certain firms are hypothesized to adopt earlier because they come into
contact with, and learn from, adopters of the new technology before
others. Economic models, on the other hand, predominantly emphasize firm
heterogeneity (Ireland and Stoneman 1986). Firms adopt technologies at
different times because they differ with respect to various
organizational and environmental variables influencing the economic
returns from adoption (Blackman 1999).

Either way, the strong assumption is that developing countries can
acquire modern technology innovated in developed economies, often at a
fraction of the original research and development (R&D) costs, thereby
leapfrogging many decades of technological progress (Teece 2000).

The first involves a country's geographical location. Recent empirical
work suggests that diffusion is ''geographically localized'' (Globerman,
Kokko, and Sjo¨holm 2000; Keller 2002; Milner 2003) in that a technology
diffuses faster in a country where it is already more widely diffused in
neighboring countries. Underlying these regional effects are contagion
and contact with prior users or producers of technology.

@perkins2011internet Two main mechanisms are identified in the
literature to explain diffusion: (1) epidemic-type dynamics whereby
contact with previous adopters stimulates uptake as po tential adopters
learn about a new innovation; and (2) economic-type mechanisms whereby
potential users adopt a new innovation as it be comesmore profitable,
useful, or valuable, with uptake characteristically spreading as costs
be come lower, performance improves, or the po tential uses of the
innovation grow over time.

To this extent, our results contribute to a growing body of work that
has sought to caution against claims about the supposed novelty of the
Internet and the suggestion that it is somehow different.

@meade2021modelling There are numerous agents who would benefit from
accurate predictions of introduction times. These include: suppliers of
network equipment for production planning; suppliers of mobile phones
who need to have inventories of appropriately configured handsets;
international network planners who need to provide sufficient bandwidth
for communications between different countries; network operators and
regulatory authorities.

Classifying adopters of a technology along a time axis was pioneered by
Rogers (1962), who categorized adopters into five groups as innovators,
early adopters, early majority, late majority and laggards.

An a priori approach groups similar countries into segments, where
similarity is defined using economic, social, or political covariates.
Examples include Hofstede (2001) and Lee (1990). An a posteriori
analysis considers segments based on realized market behaviour, for
example see Sood et al. (2009). Some studies investigate changing
segment membership over time, see Cannon and Yaprak (2011).

@fritsch2015new

@fritsch2015new analyze the spatial diffusion of laser technology
research in West Germany from 1960, when this technology began, until
2005.

Hägerstrand (1952, 1967) was the first scholar to intensively analyze
spatial diffusion pro cesses. Based on the empirical observations of the
regional spread of process innovations, he hypothesized that new
technology is first implemented in the 'center' -- the large agglom
erations -- and then moves 'down' the spatial hierarchy, the last stop
being the 'periphery', i.e. remote and sparsely populated regions. There
are at least two factors that may explain this pattern: first, the
transfer of tacit knowledge via face-to-face contact and second, the
absorptive capacity of a region.

# Literature review {\#sec2} MAYBE SKIP THIS??

# Materials and Methods {#sec3}

@wilson201281 growth function description in p. 86

from R: Asym/(1+exp((xmid-input)/scal)) using the terms from Wilson k/(1
+ e\^(t_0 - t)/scal)

Wilson: k(1 + e\^-b(t - t_0)) So, scal = - 1/b

The curve is symmetric\* around t_0

**minimum R2 = 95%**, see also @grubler1990rise

The literature usually uses the saturation level as the asymptote. I am
using the total number of websites as we cannot compute a rate.

Moran's I as @ding2010modeling **TODO** for t_0, diffusion speed

## Multi-country analysis {#sec3.1}

## 3.2 Case study approach {#sec3.2}

# Digital technologies and spatial structure: a global view {#sec4}

# The impact of ICT on the US and the UK spatial structures {#sec5}

# Discussion and conclusions {#sec6}

contrary to results from future studies regarding social media
[@lengyel2020role], web technologies did not exclusively spread from a
central location.

@lengyel2020role We see that the OSN spread almost exclusively from the
original location (the capital Budapest, with an order of magnitude more
inhabitants than the next size town) to various parts of the country in
the early phase of the life-cycle. Later, diffusion became less
mono-centric and other towns also emerged as spreaders. Our findings
support the idea that spreading initially happens to large distances and
becomes more local over time. This is illustrated and discussed later in
Figure 2I.

# References {#references .unnumbered}