### 2023-02-16 presentations/gceg.Rmd

---
title: 'Digital economy in the UK: an evolutionary story'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: mybibfile.bib
biblio-style: "apalike"
css: style.css
---

```{r, echo=FALSE, results=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(rprojroot)
library(magick)
library(purrr)
library(kableExtra)

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size

# This is the project path
path <- find_rstudio_root_file()
```

## Contents
<br>
<div style="margin-left: 50px !important; text-align:left !important">
- Introduction
- Web data
- Spatial analysis
- S-shaped diffusion curves
- Conclusions
<br>
<br>
[etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/)
</div>

# Introduction

## Aims

<br>

- Map the **active** engagement with the *digital*
- Over time, **early stages** of the internet
- **Granular** and **multi-scale** spatial perspective

## Importance

- Understand how the adoption of new technologies evolves

- Guide policies for deployment of new technologies

- Predictions of introduction times for future technologies [@meade2021modelling]:

  - Network operators

  - Suppliers of network equipment

  - Regulatory authorities

## Technological diffusion

<br>

![](images/S-shaped-logistic-diffusion-curve_W640.jpg)

## *Spatial* diffusion processes

- As in temporal diffusion models, an **S-shaped** pattern in the
cumulative level of adoption

- A **hierarchy effect**: from main centers to secondary ones -- central places

- A **neighborhood effect**: diffusion proceeds outwards from innovation centers, 
first "hitting" nearby rather than far-away locations [@grubler1990rise]

@hagerstrand1965monte: from innovative centers (core) through a
hierarchy of subcenters, to the periphery

## Internet geographies

- Map internet infrastructure

  - backbone [@malecki2002economic; @tranos2013geography]

  - last mile [@riddlesden2014broadband; @budnitz2022working]

- Social media and user generated content

  - Here and now effect [@crampton2013beyond]

> Because new digital activities are rarely---if
ever---captured in official state data, researchers must rely on
information gathered from alternative sources [@zook2022mapping].

# Web data

## Web data: The Internet Archive
- The largest archive of webpages in the world
- 273 billion webpages from over 361 million websites, 15 petabytes of storage (1996 -)
- A web crawler starts with a list of URLs (a seed list) to crawl and downloads a copy of their content
- Using the hyperlinks included in the crawled URLs, new URLs are identified and crawled (snowball sampling)
- Time-stamp

## Web data: The Internet Archive
![](C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/figure/Picture1.png){width=70%}

## Web data: The Internet Archive
![](C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/figure/Picture2.png){width=70%}

## Our web data
- JISC UK Web Domain Dataset: all archived webpages from the .uk domain 1996-2012

- Curated by the British Library

<small>

- Tranos, E., and C. Stich. 2020. Individual internet usage and the availability of online content of local interest: A multilevel approach. *Computers, Environment and Urban Systems*, 79:101371.

- Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy in the UK: Regional productivity effects of early adoption. *Regional Studies*, 55:12, 1924-1938.

- Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from the ground up: a web data approach. *Environment and Planning B*, in press.

- Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to predict regional trade flows: data extraction, modelling, and validation, *Annals of the AAG*, in press.

</small>

## Our web data

- All .uk archived webpages which contain a UK postcode in the web text

- Circa 0.5 billion URLs with valid UK postcodes

<br>
<br>

<!-- <small></small> -->
20080509162138/http://www.website1.co.uk/contact_us IG8 8HD

## Data cleaning

- All the archived .uk webpages

- Archived during 1996-2012

- Commercial webpages (.co.uk)

- From webpages to websites:

    <small>- http://www.website1.co.uk/webpage1 and</small>

    <small>- http://www.website1.co.uk/webpage2 are part of the</small>

    <small>- http://www.website1.co.uk</small>

- 1 *vs.* multuple postcodes in a website

## Unique postcodes frequencies, 2000

```{r eval=TRUE, echo=FALSE, results='asis'}

# load 2000 co.uk
# setwd("C:/Users/TranosE/DataShare/archive/nuts")
path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"

all2000.duplicates <- data.table::fread(path.2000) #
# dim(all2000.duplicates) # 3336162
# it includes duplicated URLs: if one webpages includes multiple postcodes
# then it appears multiple times. This is ok for the nuts aggregation, but
# not for the frequencies
# This is only .co.uk

# one line for every host
all2000 <- unique(all2000.duplicates, by = c("host")) # 57897

# unique postcodes per website f table
f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
                                                              "(100,1e+03]", "(1e+03,1e+04]",
                                                              "(1e+04,1e+05]"))
levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")
kable(f.websites.pc,
      format = "html",
      digits = 3,
      format.args = list(big.mark = ",")) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```
<br>

- Websites with a large number of postcodes: e.g. directories, real estate websites

- Focus on websites with **one unique** postcode per year

## Directory website with a lot of postcodes
![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)

## Website with a unique postcode in London
![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)

# Analysis

## Mapping website density

---

```{r heatmaps, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

---

```{r heatmaps_log, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

## Spatial attributes

<br>

Two scales:

- Output areas (c. 200,000)

- Local authorities (c. 400)

---

<br>

**Neighborhood effect**: diffusion proceeds outwards from innovation centers, 
first "hitting" nearby rather than far-away locations [@grubler1990rise]

- Moran's I

- LISA maps

- Website density regressions

---

```{r moransi_oa, results='asis'}
path.in <- paste0(path, "/outputs/morani_oa.csv")
# read.csv(path.in) %>% dplyr::select(year, morani) %>% 
#   rename('Moran\'s I' = morani,
#          Year = year) %>% 
#   arrange(Year) %>% 
#   kable(digits = 3, 
#         label = "Moran\'s I per year, Output Areas") %>% 
#     kable_classic(full_width = F, html_font = "Cambria")

read.csv(path.in) %>% dplyr::select(year, morani) %>% 
  ggplot(aes(x=year, y=morani)) +
  geom_bar(stat = "identity") + 
  ylab('Moran\'s I') +
  theme_minimal() +
  ggtitle("Moran\'s I, number of websites per Output Areas") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

---

```{r moransi_la, results='asis'}
path.in <- paste0(path, "/outputs/morani_la.csv")

read.csv(path.in) %>% dplyr::select(year, morani) %>% 
  ggplot(aes(x=year, y=morani)) +
  geom_bar(stat = "identity") + 
  ylab('Moran\'s I') +
  theme_minimal() +
  ggtitle("Moran\'s I, number of websites per Local Authority") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

---

```{r lisa_oa, results='asis', out.width = "30%"}
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

---

```{r lisa_la, results='asis', out.width = "30%"}
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

## Website density regressions

<br>

$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$

<br>

<div style="margin-left: 50px !important; text-align:left !important">
$Place$:

- London, *or*

- Nearest city, *or*

- Nearest retail centre

$i$: Output Areas *or* Local Authorities
</div>

## Website density regressions

<br>

$\beta$ interpretation:

- The lower the $\beta$ is (or the larger the $|\beta|$ is)...

- ... the larger urban gravitation is for web adoption.

---

```{r b_London_oa, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/b_dist_london_oa.png")
knitr::include_graphics(path.image)
```

---

```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
path.image <- paste0(path, "/outputs/b_dist_london_la.png")
knitr::include_graphics(path.image)
```

---

```{r b_city_oa, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/b_dist_city_oa.png")
knitr::include_graphics(path.image)
```

---

```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/b_dist_city_la.png")
knitr::include_graphics(path.image)
```

---

```{r b_centre_oa, message=FALSE, out.height= "60%", out.width="60%"}
#fig.height=4, fig.width=4, 
path.image <- paste0(path, "/outputs/b_dist_centre_oa.png")
knitr::include_graphics(path.image)
```

---

```{r b_centre_la, message=FALSE, out.height= "60%", out.width="60%"}
path.image <- paste0(path, "/outputs/b_dist_centre_la.png")
knitr::include_graphics(path.image)
```

```{r r_squared, eval=F, results='asis'}
# This is for the RSquared of the different website density regressions.
# The RSquared does not drop much hence it is not worth reporting.

path.in1 <- paste0(path, "/outputs/r2_all_oa.csv")
path.in2 <- paste0(path, "/outputs/r2_all_la.csv")

read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
  ggplot(aes(x=year)) +
  geom_line(aes(y=London, color = "OA to London"), size=1) +
  geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
  geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +

  geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
  geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
  geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +

  scale_color_discrete(name = "Website density regressions") +
  scale_linetype_manual(name = "Website density regressions") +
  theme_minimal()

  scale_color_manual(values = c("red", "blue", "green"))

read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
  pivot_longer(names_to = "destinations",
               values_to = "Rsquared",
               cols = 2:7) %>% 
  arrange(destinations, year) %>% 
  ggplot(aes(x=year)) +
  geom_line(aes(y=Rsquared, group=destinations, colour=destinations)) +
  geom_textline(aes(y=Rsquared, group=destinations, colour=destinations, 
                    label=destinations),
                hjust = 1) +
  theme(legend.position = "none") +
  theme_minimal()

```

---

<br>

**Hierarchy effect**: from main centers to secondary ones -- central places

- Gini coefficient

---

```{r gini_oa, results='asis'}
path.in <- paste0(path, "/outputs/gini_oa.csv")

read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
  ggplot(aes(x=year, y=gini.a.n.)) +
  geom_bar(stat = "identity") + 
  ylab("Gini") +
  theme_minimal() +
  ggtitle("Gini coefficient, number of websites per Output Area") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

---

```{r gini_la, results='asis'}
path.in <- paste0(path, "/outputs/gini_la.csv")

read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
  ggplot(aes(x=year, y=gini.a.n.)) +
  geom_bar(stat = "identity") + 
  ylab("Gini") +
  theme_minimal() +
  ggtitle("Gini coefficient, number of websites per Local Auhtority") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

## Spatial attributes, a summary

- Spatial dependency relatively small and constant over time / scales

- At local scale, consistent hotspots over time

- More granular analysis reveals hotspots

- Almost perfect polarisation of web adoption in the early stages at a granular level

- More equally diffused at the Local Authority level

- Plateau overtime

- Distance effect: urban gravitation increases over time and then drops

- Consistent across scales and definitions of urban

# S-shaped diffusion curves

---

```{r UK S, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_uk.png")
knitr::include_graphics(path.image)
```

```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

library(magick)
las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
las_list <- lapply(las, image_read)

## join the images together
las_joined <- image_join(las_list)

## animate at 2 frames per second
las_animated <- image_animate(las_joined, fps = 2)

## view animated image
las_animated

## save to disk
image_write(image = las_animated,
            path = "las.gif")

  # map(image_read) %>% # reads each path file
  # image_join() %>% # joins image
  # image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

---

```{r Xmid_frequency_graph, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/xmid_f.png")
knitr::include_graphics(path.image)
```

---

```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/speed_pam.png")
knitr::include_graphics(path.image)
```

---

<!-- fast  -->

```{r Darlington, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000005.png")
knitr::include_graphics(path.image)
```

---

```{r Nottingham, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000018.png")
knitr::include_graphics(path.image)
```

---

```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000038.png")
knitr::include_graphics(path.image)
```

---

```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E09000001.png")
knitr::include_graphics(path.image)
```

---

<!-- slow -->

```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000002.png")
knitr::include_graphics(path.image)
```

---

```{r Leicester, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000016.png")
knitr::include_graphics(path.image)
```

---

```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000023.png")
knitr::include_graphics(path.image)
```

---

```{r Plymouth, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E06000026.png")
knitr::include_graphics(path.image)
```

---

```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E08000012.png")
knitr::include_graphics(path.image)
```

---

```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_E09000012.png")
knitr::include_graphics(path.image)
```

---

```{r Edinburgh, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/lad_S12000036.png")
knitr::include_graphics(path.image)
```

# Conclusions

## What have we learned?

- **Geography matters:** spatial dependency, urban gravitation

- Some indications of a hierarchical diffusion

- Granular analysis reveals patterns otherwise not visible

- Well-established theoretical approaches of diffusion survive even at a granular level

## What have we *yet* to learn?

- Explain the spatial patterns of fast/slow web adoption

  - Applied the same analysis fo OA

  - Survival regressions [@perkins2005international]

- Expand our definition of web to websites with # of postcodes > 1

## References {.allowframebreaks}