### 2023-03-20 src/hex_heatmaps.Rmd
lisa per firm

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(spdep)
library(tmap)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, msoa11cd, dointr, doterm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd

path.lookup <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oseast1m, osnrth1m) %>% 
  dplyr::rename(pc = pcds)

```

The internet archive data is saved on /hdd.

```{r}
n = 1 #number of unique postcodes

data.folder9610 <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder9610, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

data.path1112 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) 
df1112 <- read_csv(data.path1112)

df <- bind_rows(df9610, df1112) %>% 
  filter(V1.domain == n,
         year > 1995) %>% 
  left_join(lookup, by = "pc", suffix =c("",""))

rm(df9610, df1112)
```

## Boundary data

```{r}
# get MSOA 
# source: https://geoportal.statistics.gov.uk/datasets/middle-layer-super-output-areas-december-2011-boundaries-generalised-clipped-bgc-ew-v3/explore?location=52.849644%2C-2.489527%2C7.08

# path.geo <- paste0(path, "/data/raw/Middle_Layer_Super_Output_Areas_(December_2011)_Boundaries_Generalised_Clipped_(BGC)_EW_V3.geojson")
# msoa <- readOGR(path.geo)
# # UK BFC: la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson") 
# # source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe
# 
# # spatial transformations
# msoa <- spTransform(msoa, CRS("+init=epsg:4326"))

# get GB
path.gb <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
gb <- readOGR(path.gb) 
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
gb <- subset(gb, CTRY20NM != "Northern Ireland")
gb.df <- fortify(gb, region = "CTRY20NM")

# UK boundaries
# path.uk <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
# uk <- readOGR(path.uk)
# uk <- spTransform(uk, CRS("+init=epsg:4326"))
# uk.df <- fortify(uk, region = "CTRY20CD")
# 
# # find points not in GB
# help2 <- df %>% 
#   #slice_sample(n = 10000) %>% 
#   filter(!is.na(oseast1m), #6k rows with missing east/north for all years
#               year == t) %>%
#   st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() 
# 
# help2 <- df %>% 
#   filter(!is.na(oseast1m), #6k rows with missing east/north for all years
#               year == t) %>%
#   bind_cols(help2)
# 
# 
# points <- st_as_sf(help2, coords = c("X", "Y"), crs = 27700)
# gb_ <- st_as_sf(gb) #, crs = 27700)
# gb_ <- st_set_crs(gb_, 27700)
# 
# plot(gb_, col = "red")
# plot(points, pch = 19, cex = 1, add = T)
# 
# pip <- points %>%
#   st_join(gb_)
# 
# unique(pip$ctry14nm)
# missing <- pip %>% filter(is.na(ctry14nm)) %>% 
#   mutate(pc2dig = substr(pc, 1,2)) 
# unique(missing$pc2dig)
# 
# missing <- pip %>% filter(is.na(ctry14nm)) %>% 
#   filter(substr(pc, 1,2) !="BT")
# in 2010 58 points exl. NI have a postcode but not XY

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

```

## hex heatmaps

```{r}
# old code: counts per hexagon based on xy
#xy point

path.out <- paste0(path, "/outputs/hex", n)

for (t in 1996:2012){
  p <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == t,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(4326) %>%
    st_coordinates() %>%
    as_tibble() %>% 
    ggplot(aes(x=X,y=Y)) +

    stat_binhex(bins=150) + 
    geom_polygon(data = gb.df, aes(x=long, y = lat, group = group), #gb.df
                 fill=NA, color = "black", size = 0.1, alpha = 0.8)+
    scale_fill_gradientn(colours=c('light gray','blue'),
                         name='Frequency',na.value=NA) + #, trans = "log10") +
    coord_equal() +
    labs(x=NULL, y=NULL) +
    geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
    geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
    #theme(legend.key.size = unit(0.05, "cm")) + # legend.key.width=unit(1, "cm"), 
    ggmap::theme_nothing(legend = T) +
    theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
    #theme(legend.position=c(0.075,0.28))
    guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
    ggtitle(t)
  ggsave(p, filename = paste(path.out, t, ".png", sep = ""))
}

# geo_sf solution. same as lisa, but squeese hexagons
# gb %>% st_as_sf(crs = 27700) %>% 
# ggplot() + 
#     geom_sf(fill=NA, color = "black", size = 0.1, alpha = 0.8) + #geom_point(data = help.df, aes(x=X,y=Y))
#     stat_binhex(data = help.df, aes(x=X,y=Y),bins=150) + 
#     # geom_polygon(data = gb.df, aes(x=long, y = lat, group = group), #gb.df
#     #              fill=NA, color = "black", size = 0.1, alpha = 0.8)+
#     
#     #coord_equal() +
#     scale_fill_gradientn(colours=c('light gray','blue'),
#                          name='Frequency',na.value=NA) + #, trans = "log10") +
#     
#     labs(x=NULL, y=NULL) +
#     geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
#     geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
#     #theme(legend.key.size = unit(0.05, "cm")) + # legend.key.width=unit(1, "cm"), 
#     ggmap::theme_nothing(legend = T) +
#     theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
#     #theme(legend.position=c(0.075,0.28))
#     guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
#     ggtitle(t)
```
## hex heatmaps log transformed

```{r}
# old code: counts per hexagon based on xy
#xy point

path.out <- paste0(path, "/outputs/hex_log", n)

for (t in 1996:2012){
  p <- df %>%
    #slice_sample(n = 10000) %>% 
    filter(!is.na(oseast1m), #6k rows with missing east/north for all years
           year == t,
           substr(pc, 1,2) !="BT") %>% #drop Northern Ireland, for 2010 c. 4k obs.
    st_as_sf(coords = c("oseast1m", "osnrth1m"), crs = 27700) %>%
    st_transform(4326) %>%
    st_coordinates() %>%
    as_tibble() %>% 
    ggplot(aes(x=X,y=Y)) +

    #stat_binhex(bins=150) +
    stat_binhex(aes(fill=log(..count..)), bins=150) +
    geom_polygon(data = gb.df, aes(x=long, y = lat, group = group), #gb.df
                 fill=NA, color = "black", size = 0.1, alpha = 0.8)+
    scale_fill_gradientn(colours=c('light gray','blue'),
                         name='log(Frequency)',na.value=NA) + #, trans = "log10") +
    coord_equal() +
    labs(x=NULL, y=NULL) +
    geom_point(data = cities, aes(x=long, y=lat), colour = "black", size = .1) + 
    geom_text(size = 2, check_overlap = F, nudge_y = .2, data = cities, aes(x=long, y=lat, label = name)) +
    #theme(legend.key.size = unit(0.05, "cm")) + # legend.key.width=unit(1, "cm"), 
    ggmap::theme_nothing(legend = T) +
    theme(legend.title=element_text(size=5.5), legend.text=element_text(size=4)) +
    #theme(legend.position=c(0.075,0.28))
    guides(fill = guide_colorbar(barwidth = 0.5, barheight = 2.5)) +
    ggtitle(t)
  ggsave(p, filename = paste(path.out, t, ".png", sep = ""))
}
```