### 2023-10-27 presentations/web_diffusion.Rmd

---
title: 'Digital economy in the UK: an evolutionary story'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: bibliography.bib
biblio-style: "apalike"
css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, results=FALSE, include=FALSE}
library(knitr)
library(tidyverse)
library(rprojroot)
library(magick)
library(purrr)
library(kableExtra)

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size

# This is the project path
path <- find_rstudio_root_file()
```

## Contents

<br>

::: {style="margin-left: 50px !important; text-align:left !important"}
-   Introduction
-   Web data
-   Spatial analysis
-   S-shaped diffusion curves
-   Rank dynamics
-   Conclusions <br> <br>
    <!-- [etranos.info/post/gceg2022](https://etranos.info/post/gceg2022/) -->
:::

# Introduction

## Aims

<br>

-   Map the **active** engagement with the *digital*
-   Over time, **early stages** of the internet
-   **Granular** and **multi-scale** spatial perspective

## Aims

-   Diffusion of a new technology: the *web*
-   Geographers used to be interested in diffusion
-   @hagerstrand1968innovation
-   Passed the torch to economists and sociologists 
-   Why? Lack of granular data:

> Because new digital activities are rarely---if ever---captured in
> official state data, researchers must rely on information gathered
> from alternative sources [@zook2022mapping].

## Importance

-   Understand how the adoption of new technologies evolves

-   Guide policies for deployment of new technologies

-   Predictions of introduction times for future technologies
    [@meade2021modelling]:

    -   Network operators

    -   Suppliers of network equipment

    -   Regulatory authorities

## Technological diffusion

<br>

![](images/S-shaped-logistic-diffusion-curve_W640.jpg)

## *Spatial* diffusion processes

-   As in temporal diffusion models, an **S-shaped** pattern in the
    cumulative level of adoption

-   A **hierarchy effect**: from main centres to secondary ones --
    central places

-   A **neighborhood effect**: diffusion proceeds outwards from
    innovation centres, first "hitting" nearby rather than far-away
    locations [@grubler1990rise]

@hagerstrand1965monte: from innovative centres (core) through a
hierarchy of sub-centres, to the periphery

# Web data

## Long story short

<br>

-   Archived web data

-   Observe commercial websites 1996 - 2012 in the UK (*.co.uk*)

-   Geolocate:

    - to a unique postcode

    - to multiple postcodes

-------------------------------------------

```{r , eval=T, message=FALSE, out.width = "1200px"}
path.image <- paste0(path, "/presentations/images/past_publications.png")
knitr::include_graphics(path.image)
```

## Web data: The Internet Archive

-   The largest archive of webpages in the world
-   273 billion webpages from over 361 million websites, 15 petabytes of
    storage (1996 -)
-   A web crawler starts with a list of URLs (a seed list) to crawl and
    downloads a copy of their content
-   Using the hyperlinks included in the crawled URLs, new URLs are
    identified and crawled (snowball sampling)
-   Time-stamp

## Web data: The Internet Archive

![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture1.png){width="70%"}

## Web data: The Internet Archive

![](C:/Users/nw19521/OneDrive%20-%20University%20of%20Bristol/projects/archive/nuts/figure/Picture2.png){width="70%"}

## Our web data

-   JISC UK Web Domain Dataset: all archived webpages from the .uk
    domain 1996-2012

-   Curated by the British Library

<small>

-   Tranos, E., and C. Stich. 2020. Individual internet usage and the
    availability of online content of local interest: A multilevel
    approach. *Computers, Environment and Urban Systems*, 79:101371.

-   Tranos, E., T. Kitsos, and R. Ortega-Argilés. 2021. Digital economy
    in the UK: Regional productivity effects of early adoption.
    *Regional Studies*, 55:12, 1924-1938.

-   Stich, C., E. Tranos and M. Nathan. 2022. Modelling clusters from
    the ground up: a web data approach. *Environment and Planning B*, in
    press.

-   Tranos, E., A. C. Incera and G. Willis. 2022. Using the web to
    predict regional trade flows: data extraction, modelling, and
    validation, *Annals of the AAG*, in press.

</small>

## Our web data

-   All .uk archived webpages which contain a UK postcode in the web
    text

-   Circa 0.5 billion URLs with valid UK postcodes

<br> <br>

<!-- <small></small> -->

20080509162138/<http://www.website1.co.uk/contact_us> IG8 8HD

## Data cleaning

-   All the archived .uk webpages

-   Archived during 1996-2012

-   Commercial webpages (.co.uk)

-   From webpages to websites:

    <small>- <http://www.website1.co.uk/webpage1> and</small>

    <small>- <http://www.website1.co.uk/webpage2> are part of
    the</small>

    <small>- <http://www.website1.co.uk></small>

-   1 *vs.* multuple postcodes in a website

## Unique postcodes frequencies, 2000

```{r eval=TRUE, echo=FALSE, results='asis'}

# load 2000 co.uk
# setwd("C:/Users/TranosE/DataShare/archive/nuts")
path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"

all2000.duplicates <- data.table::fread(path.2000) #
# dim(all2000.duplicates) # 3336162
# it includes duplicated URLs: if one webpages includes multiple postcodes
# then it appears multiple times. This is ok for the nuts aggregation, but
# not for the frequencies
# This is only .co.uk

# one line for every host
all2000 <- unique(all2000.duplicates, by = c("host")) # 57897

# unique postcodes per website f table
f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
                                                              "(100,1e+03]", "(1e+03,1e+04]",
                                                              "(1e+04,1e+05]"))
levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")
kable(f.websites.pc,
      format = "html",
      digits = 3,
      format.args = list(big.mark = ",")) %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

<br>

-   Websites with a large number of postcodes: e.g. directories, real
    estate websites

-   Focus on websites with **one unique** postcode per year

## Directory website with a lot of postcodes

![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_directory_website_2010.png)

## Website with a unique postcode in London

![](C:/Users/nw19521/OneDrive - University of Bristol/projects/shoreditch/paper/images_screenshots/screenshot_2010.png)

# Web diffusion

## Mapping website density

------------------------------------------------------------------------

```{r heatmaps, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

<!-- ------------------------------------------------------------------------ -->

```{r eval = F, heatmaps_log, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

list.files(path = path.images, pattern = "^hex_log1", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

---

> **Neighbourhood effect**: diffusion proceeds outwards from innovation
> centers, first "hitting" nearby rather than far-away locations
> [@grubler1990rise]

-   Spatial dependency (Moran's I & LISA maps)

-   Website density regressions -- distance effect

-   Websites per firm in Local authorities (c. 400)

-   Websites in Output Areas (c. 200,000)

------------------------------------------------------------------------

```{r moransi_la, results='asis'}
path.in <- paste0(path, "/outputs/lisa/morani_la.csv")

read.csv(path.in) %>% dplyr::select(year, morani) %>% 
  ggplot(aes(x=year, y=morani)) +
  geom_bar(stat = "identity") + 
  ylab('Moran\'s I') +
  theme_minimal() +
  ggtitle("Moran\'s I, websites per firm a Local Authority") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

------------------------------------------------------------------------

```{r, eval=T, moransi_oa, results='asis'}
path.in <- paste0(path, "/outputs/lisa/morani_oa.csv")
# read.csv(path.in) %>% dplyr::select(year, morani) %>% 
#   rename('Moran\'s I' = morani,
#          Year = year) %>% 
#   arrange(Year) %>% 
#   kable(digits = 3, 
#         label = "Moran\'s I per year, Output Areas") %>% 
#     kable_classic(full_width = F, html_font = "Cambria")

read.csv(path.in) %>% dplyr::select(year, morani) %>% 
  ggplot(aes(x=year, y=morani)) +
  geom_bar(stat = "identity") + 
  ylab('Moran\'s I') +
  theme_minimal() +
  ggtitle("Moran\'s I, websites per Output Area") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

------------------------------------------------------------------------

```{r lisa_la, results='asis', out.width = "30%"}
path.images <- paste0(path, "/outputs/lisa/")

list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

------------------------------------------------------------------------

```{r lisa_oa, eval=T,results='asis', out.width = "30%"}
path.images <- paste0(path, "/outputs/lisa/")

list.files(path = path.images, pattern = "^lisa_level_pc1_oa", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

## Website density regressions

<br>

$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$

<br>

::: {style="margin-left: 50px !important; text-align:left !important"}
$Website\,Density_{i}$:

-   Websites per firm in a Local Authority $i$, *or*

-   Websites in an Output Area $i$

:::

## Website density regressions

<br>

$$Website\,Density_{i} = a + \beta Distance\,to\,Place_{i} + e_{i}$$

<br>

::: {style="margin-left: 50px !important; text-align:left !important"}
$Place$:

-   London, *or*

-   Nearest city, *or*

-   Nearest retail centre

:::

## Website density regressions

<br>

$\beta$ interpretation:

-   The lower the $\beta$ is (or the larger the $|\beta|$ is)...

-   ... the larger urban gravitation is for web adoption.

------------------------------------------------------------------------

```{r b_London_la, fig.height=5, fig.width=5, message=FALSE}
path.image <- paste0(path, "/outputs/lisa/b_dist_london_la.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r b_city_la, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_city_la.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r b_centre_la,  fig.height=5, fig.width=5,message=FALSE}
#, out.height= "60%", out.width="60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_centre_la.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r b_London_oa, eval=T, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_london_oa.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r b_centre_oa, eval = T,  fig.height=5, fig.width=5,message=FALSE}
#fig.height=4, fig.width=4, 
#, out.height= "60%", out.width="60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_city_oa.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r b_city_oa, eval = T, fig.height=5, fig.width=5, message=FALSE}
#, out.width = "60%"
path.image <- paste0(path, "/outputs/lisa/b_dist_centre_oa.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r r_squared, eval=T, results='asis'}
# This is for the RSquared of the different website density regressions.

path.in1 <- paste0(path, "/outputs/lisa/r2_all_oa.csv")
path.in2 <- paste0(path, "/outputs/lisa/r2_all_la.csv")

# read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
#   ggplot(aes(x=year)) +
#   geom_line(aes(y=London, color = "OA to London"), size=1) +
#   geom_line(aes(y=cities, color = "OA to nearest city"), size=1) +
#   geom_line(aes(y=retail, color = "OA to nearest retail centre"), size=1) +
# 
#   geom_line(aes(y=London.LA, color = "LA to London"), size=1, linetype = 2) +
#   geom_line(aes(y=cities.LA, color = "LA to nearest city"), size=1, linetype = 2) +
#   geom_line(aes(y=retail.LA, color = "LA to nearest retail centre"), size=1, linetype = 2) +
# 
#   scale_color_discrete(name = "R-squared for regression:") +
#   #scale_linetype_manual(name = "Website density regressions111") 
#   ylab("R-squared") + 
#   ggtitle("Website density regressions") +
#   theme_minimal() +
#   theme(axis.title.x=element_blank(),
#         plot.title = element_text(hjust = 0.5))

read.csv(path.in1) %>% left_join(read.csv(path.in2)) %>% 
  pivot_longer(!year, names_to = "regression", values_to = "Rsquared") %>% 
  mutate(LA = if_else(str_ends(regression, ".LA"), " to LA", "to OA"),
         regression = str_remove_all(regression, ".LA"),
         line = paste0(regression, " ", LA)) %>% 
  ggplot() +
  geom_line(aes(x=year, y = Rsquared, color = line),
            size = 1) +
  scale_color_discrete(name = "R-squared for regression:") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("R-squared") + 
  ggtitle("Website density regressions") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

#   scale_color_manual(values = c("red", "blue", "green"))

```

## Neighbourhood effect

- Spatial dependency
	- Relatively small and constant over time / scales
	- London hot spot early on
	- At local scale, consistent hotspots over time
	- Granular analysis reveals other hotspots

- Distance effect
	- Urban gravitation increases over time and then drops
	- Granular analysis: gravitation
	- Lost explanatory power over time

------------------------------------------------------------------------

<br>

> **Hierarchy effect**: from main centers to secondary ones -- central
places

-   Gini coefficient

------------------------------------------------------------------------

```{r gini_la, results='asis'}
path.in <- paste0(path, "/outputs/lisa/gini_la.csv")

read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
  ggplot(aes(x=year, y=gini.a.n.)) +
  geom_bar(stat = "identity") + 
  ylab("Gini") +
  theme_minimal() +
  ggtitle("Gini coefficient, websites per firm in Local Auhtority") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

------------------------------------------------------------------------

```{r gini_oa, eval=T, results='asis'}
path.in <- paste0(path, "/outputs/lisa/gini_oa.csv")

read.csv(path.in) %>% dplyr::select(year, gini.a.n.) %>% 
  ggplot(aes(x=year, y=gini.a.n.)) +
  geom_bar(stat = "identity") + 
  ylab("Gini") +
  theme_minimal() +
  ggtitle("Gini coefficient, websites per Output Area") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

## Hieararchy

-   Almost perfect polarisation of web adoption in the early stages at a
    granular level

-   Polarisation decreases over time

-   More equally diffused at the Local Authority level

# S-shaped diffusion curves

------------------------------------------------------------------------

```{r UK S, eval=T}
#, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"

path.image <- paste0(path, "/outputs/s_uk.png")
knitr::include_graphics(path.image)
```

```{r LA_S_gif, eval=F, results='asis', out.width = "60%"}
path.images <- paste0(path, "/outputs/")

library(magick)
las <- list.files(path = path.images, pattern = "^lad_", full.names = T) #%>% 
las_list <- lapply(las, image_read)

## join the images together
las_joined <- image_join(las_list)

## animate at 2 frames per second
las_animated <- image_animate(las_joined, fps = 2)

## view animated image
las_animated

## save to disk
image_write(image = las_animated,
            path = "las.gif")

  # map(image_read) %>% # reads each path file
  # image_join() %>% # joins image
  # image_animate(delay=50) #%>% # animates, can opt for number of loops
  #image_write("test.gif") # write to current dir
```

------------------------------------------------------------------------

```{r Xmid_frequency_graph, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/xmid_f.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r diffusion_speed_map, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/speed_pam.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

<!-- fast  -->

```{r Darlington, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000005.png")
knitr::include_graphics(path.image)
```

<!-- ------------------------------------------------------------------------ -->

```{r Nottingham, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000018.png")
knitr::include_graphics(path.image)
```

<!-- ------------------------------------------------------------------------ -->

```{r Reading, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000038.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r City_of_London, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000001.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

<!-- slow -->

```{r Middlesbrough, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000002.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r Leicester, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000016.png")
knitr::include_graphics(path.image)
```

<!-- ------------------------------------------------------------------------ -->

```{r Bristol, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000023.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r Plymouth, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E06000026.png")
knitr::include_graphics(path.image)
```

<!-- ------------------------------------------------------------------------ -->

```{r Liverpool, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E08000012.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r Hackney, eval=T, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_E09000012.png")
knitr::include_graphics(path.image)
```

<!-- ------------------------------------------------------------------------ -->

```{r Edinburgh, eval=F, fig.height=5, fig.width=5, message=FALSE, out.width = "60%"}
path.image <- paste0(path, "/outputs/s_lad_per_firm/lad_S12000036.png")
knitr::include_graphics(path.image)
```

## Diffusion speed

<br>

- Spatial heterogeneity
- Not clear pattern

# Rank dynamics: stability vs. volatility

------------------------------------------------------------------------

<br>

- Adoption heterogeneity 
- Perceptions of risk and economic returns from new technologies
- Early adopters vs. laggards, leapfrogging

------------------------------------------------------------------------

```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only05.png")
knitr::include_graphics(path.image)
```

------------------------------------------------------------------------

```{r , eval=T, fig.height=5, fig.width=5, message=FALSE}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only95.png")
knitr::include_graphics(path.image)
```

# Conclusions

----

<br>

-   **Geography matters:** spatial dependency, urban gravitation

-   Some indications of a hierarchical diffusion

-   Granular analysis reveals patterns otherwise not visible

-   Well-established theoretical approaches of diffusion survive even at
    a granular level

-   Stability and volatility: leapfrogging, early adopters dropping, but also maintaining

## References {.allowframebreaks}