### 2023-11-05 src/help_predict.Rmd
rf

---
title: "Untitled"
author: "Emmanouil Tranos"
date: "17/04/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(caret)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

```{r}
df.model <- df %>% complete(year, oa11cd, fill = list(n = 0)) %>% 
  arrange(oa11cd, year) %>% 
  left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
  left_join(dist.retail, by = c("oa11cd" = "oa11cd")) %>% 
  filter(!is.na(London)) %>% 
  dplyr::select(year, oa11cd, n, London, dist, dist.retail) %>% 
  left_join(classes.oa %>% dplyr::select(id, subgroup), by = c("oa11cd" = "id")) %>% 
  mutate(subgroup == as.factor(subgroup))

sapply(df.model, function(x) sum(is.na(x))) # mising NI
```

```{r}

train <- df.model %>% filter(year>2011)# | year==2009)
test  <- df.model %>% filter(year == 2012)

train_x <- train %>% dplyr::select(London, dist, dist.retail) #subgroup
train_x = scale(train_x)[,]
train_y = train %>% dplyr::select(n) %>% as.matrix()

test_x <- test %>% dplyr::select(London, dist, dist.retail) #subgroup
test_x = scale(test_x)[,]
test_y = test %>% dplyr::select(n) %>% as.matrix()

knnmodel = knnreg(train_x, train_y)

str(knnmodel)

pred_y = predict(knnmodel, data.frame(test_x))

print(data.frame(test_y, pred_y))

mse = mean((test_y - pred_y)^2)
mae = caret::MAE(test_y, pred_y)
rmse = caret::RMSE(test_y, pred_y)
#rsqr =

x = 1:length(test_y)

# plot(x, test_y, col = "red", type = "l", lwd=2,
#      main = "Boston housing test data prediction")
# lines(x, pred_y, col = "blue", lwd=2)
# legend("topright",  legend = c("original-medv", "predicted-medv"), 
#        fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
# grid()

test_y %>% as_tibble() %>% ggplot() +
  geom_line(aes(x = x, y = test_y), col = "red", lwd=.5) +
  geom_line(aes(x = x, y = pred_y), col = "blue", lwd=.5)

test_y %>% as_tibble() %>% ggplot() +
  geom_point(aes(x = test_y, y = pred_y))

```