### 2024-02-12 src/bua_RF.Rmd
bua

---
title: "Random Forests", BUAs
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

TODO:

- check missing bua
- check out
- regions
- growth

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
#library(rgdal)
#library(rgeos)
library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)
library(caret)
library(randomForest)
library(CAST)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>%  # for n == 1 
  ungroup()

# Partially complete panel as not all OAs have a website in at least one year
df <- df %>% filter(!is.na(oa11cd)) %>% 
  complete(oa11cd, year) %>% 
  replace(is.na(.), 0)

# temp for working on laptop
# path.in <- paste0(path, "/data/temp/help.oa.csv")
# df <- read_csv(path.in)

# https://geoportal.statistics.gov.uk/datasets/df834d6d654d4f4a9a211fdbbb8893bd_0/explore
# path.lookup.rg <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")
# lookup.rg <- read_csv(path.lookup.rg) %>% 
#   dplyr::select(-ObjectId)
```

## spatial data

```{r}

# Get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- st_read(path.geo)

# Spatial transformations
oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84

# Rename columns and keep only necessary columns
oa.ew <- oa.ew %>% dplyr::select(OA11CD, Shape__Area) %>% 
  rename(id = OA11CD,
         area = Shape__Area)
# oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
# colnames(oa.ew) <- c('id', 'area')

# Check the data structure
print(summary(oa.ew))

# Get OA for Scotland
path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- st_read(dsn = path.geo.sc, layer = "OutputArea2011_MHW")

# Spatial transformations
oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84

# Rename columns and keep only necessary columns
oa.sc <- oa.sc %>% dplyr::select(code, SHAPE_1_Ar) %>% 
  rename(id = code,
         area = SHAPE_1_Ar)
# oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
# colnames(oa.sc) <- c('id', 'area')

# Check the data structure
print(summary(oa.sc))

# Build OA for GB
oa.gb <- rbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# # Get GB
# path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
# gb <- st_read(path.gb)
# 
# # Spatial transformations
# gb <- st_transform(gb, 4326)  # EPSG code for WGS84
```

## Full panel

```{r}
#oa.sf <- st_as_sf(oa.uk)

oa.full <- replicate(17, oa.gb %>% as_tibble(), simplify = FALSE) %>%
  bind_rows(.id = 'id_') %>% 
  rename(year = id_) %>% 
  mutate(year = as.numeric(year) + 1995) %>% 
  dplyr::select(id, year)

df <- oa.full %>% left_join(df, by = c("id" = "oa11cd", "year" = "year")) %>% 
  replace(is.na(.), 0)

sapply(df, function(x) sum(is.na(x)))
rm(oa.full)
```

## OA to BUA

```{r}
# long to wide
df.w <- df %>% pivot_wider(names_from = year, values_from = n)

sf_use_s2(FALSE)
oa.gb.c <- st_centroid(oa.gb) %>%
  left_join(df.w)

# BUA
# source: https://geoportal.statistics.gov.uk/datasets/ad30b234308f4b02b4bb9b0f4766f7bb_0/explore
bua.path <- paste0(path, "/data/raw/BUA_2022_GB_-8042218937152150708.geojson")
bua <- st_read(bua.path)
bua <- st_transform(bua, 4326)  # EPSG code for WGS84

# OA fall within a BUA
df.within <- st_join(oa.gb.c, bua, join = st_intersects) %>% #, join = st_intersects
  as_tibble()

df.out <- df.within %>% filter(is.na(OBJECTID))
df.within <- df.within %>% filter(!is.na(OBJECTID))

# help <- st_is_within_distance(oa.gb.c %>% filter(id %in% df.out$id), bua, dist=500) %>% #, join = st_intersects
#   as_tibble()
df.out <- st_join(oa.gb.c %>% filter(id %in% df.out$id), bua, join = st_nearest_feature) %>% #, join = st_intersects
  as_tibble()

df <- bind_rows(df.within, df.out) %>% # 142953
  dplyr::select(BUA22CD, BUA22NM, 3:19) %>% 
  pivot_longer(!c(BUA22CD, BUA22NM), names_to = "year", values_to = "n") %>% 
  group_by(BUA22CD, year) %>% 
  summarise(n = sum(n)) %>% 
  ungroup() %>% 
  left_join(bua %>% as_tibble() %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
  relocate(BUA22NM, .after = BUA22CD) %>% 
  mutate(year = as.numeric(year))

# # WE ARE MISSING 42855 websites were in OA not in a BUA
# # The above might help but takes a long time
# df <- st_join(oa.gb.c, bua, join = st_intersects) %>% # 94265
#   as_tibble() %>% 
#   dplyr::select(BUA22CD, BUA22NM, 3:19) %>% 
#   pivot_longer(!c(BUA22CD, BUA22NM), names_to = "year", values_to = "n") %>% 
#   group_by(BUA22CD, year) %>% 
#   summarise(n = sum(n)) %>% 
#   ungroup() %>% 
#   left_join(bua %>% as_tibble() %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
#   relocate(BUA22NM, .after = BUA22CD) %>% 
#   mutate(year = as.numeric(year))

sapply(df, function(x) sum(is.na(x)))  
# df <- df %>% filter(is.na(BUA22CD))
# 42855 websites were in OA not in a BUA (3,536,635 - 3,493,780)
```

## Distances to cities

### City BUA

```{r}

city.names <- c("City of Westminster",
"Birmingham",
"Glasgow",
"Liverpool",
"Bristol",
"Manchester",
"Sheffield",
"Leeds",
"Edinburgh",
"Leicester")

#cities.bua <- df %>% filter(BUA22NM %in% city.names) %>% distinct(BUA22CD)
cities.bua <- bua %>% filter(BUA22NM %in% city.names)

cities.bua.c <- cities.bua %>% filter(BUA22NM %in% city.names) %>% 
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame() %>%
  #add_rownames(var = "id") %>% 
  bind_cols(cities.bua %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
  dplyr::select(BUA22CD, BUA22NM, X, Y)

bua.c <- bua %>% 
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame() %>%
  #add_rownames(var = "id") %>% 
  bind_cols(bua %>% dplyr::select(BUA22CD, BUA22NM)) %>% 
  dplyr::select(BUA22CD, BUA22NM, X, Y)
```

### Distance to cities

```{r}

dist.gb <- distm(cbind(bua.c$X, bua.c$Y), cbind(cities.bua.c$X, cities.bua.c$Y), fun=distHaversine) 
dist.gb <- round((dist.gb/1000),2) %>% 
  as_tibble()

names(dist.gb) <- cities.bua.c$BUA22NM #city.names

dist.gb <- dist.gb %>% 
  mutate(BUA22CD = bua.c$BUA22CD,
         BUA22NM = bua.c$BUA22NM) %>% 
  relocate(c(BUA22CD, BUA22NM))

# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
# dist$distMet <- apply(dist[,2:11], 1, min)

# min distance
dist.gb <- transform(dist.gb, dist = do.call(pmin, dist.gb[-(1:2)]))

# City name for minimum distance 
# +2L to accommodate for the first two columns: BUA22CD and BUA22NM
dist.gb <- dist.gb %>% mutate(dist.city.name = names(.)[max.col(.[3:12]*-1)+2L])

# Join with distance and area
df <- df %>% left_join(dist.gb %>% dplyr::select(-BUA22NM))

sapply(df, function(x) sum(is.na(x)))
```

## n for London and nearest city

```{R}
df <- df %>% mutate(dist.city.name = ifelse(dist.city.name == "City.of.Westminster",
                                             "City of Westminster", dist.city.name)) %>% 
  left_join(cities.bua.c %>%
              dplyr::select(BUA22CD, BUA22NM) %>% 
              rename(BUA22CD.city = BUA22CD),
            by = c("dist.city.name" = "BUA22NM")) %>% 
  left_join(df %>% dplyr::select(BUA22CD, year, n) %>% 
              rename(n.nearest.city = n),
            by = c("BUA22CD.city" = "BUA22CD", "year" = "year")) %>% 
  left_join(df %>% dplyr::select(BUA22NM, year, n) %>% 
              rename(n.London = n) %>% 
              filter(BUA22NM == "City of Westminster") %>% 
              rename(help = BUA22NM),
            by = c("year" = "year")) %>% 
  dplyr::select(-c(help, Leeds, Liverpool, Manchester, Sheffield, Leicester, Birmingham, 
                   Bristol, Edinburgh, Glasgow))
```

## Spatial and spatio-temporal lags

```{r}

bua.used <- df %>% distinct(BUA22CD) %>% as.vector()
sf_use_s2(FALSE)
bua.c <- bua %>% filter(BUA22CD %in% bua.used$BUA22CD) %>% 
  st_centroid()
kn <- knearneigh(bua.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(BUA22CD) %>% 
  mutate(n.lag = dplyr::lag(n, n=1, order_by=year),
         n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.London.lag = dplyr::lag(n.London, n=1, order_by = year),
         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year)) %>%
  ungroup()
```

## RF with CAST

```{r}

# Lookup: BUA to regions

# source: https://www.data.gov.uk/dataset/aca8408d-e422-4ec2-a22f-9bd8beb1d549/built-up-area-to-region-december-2022-lookup-in-great-britain
path.lookup <- paste0(path, "/data/raw/Built_Up_Area_to_Region_(December_2022)_Lookup_in_Great_Britain.csv")

# 142 BUA in more than one regions
lookup.region <- read_csv(path.lookup)

df <- df %>% left_join(lookup.region %>% 
                         filter(WHOLE_PART=="Whole") %>% 
                         dplyr::select(BUA22CD, RGN22CD, RGN22NM), 
                       by = c("BUA22CD" = "BUA22CD")) %>% 
  #filter(is.na(RGN22CD))                                      #142 BUA in more than one regions dropped
  relocate(c(RGN22NM, RGN22CD), .after=BUA22NM) %>% 
  ungroup()

# # North/South
# #lookup.region %>% distinct(RGN21NM, .keep_all = T)
# south <- c("London", "South West", "South West", "East of England")
# df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))

# For growth:
df <- df %>% group_by(BUA22CD) %>% 
  mutate(#n.lag = dplyr::lag(n),
         growth = log(n/n.lag),
                    abs.growth = n - n.lag) %>% 
  relocate(n.lag, .after = n)

# out.path <- paste0(path, "/data/temp/df_for_bua_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)

# rm(list=setdiff(ls(), "df"))

# # This is the project path
# path <- find_rstudio_root_file()

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(BUA22CD, BUA22NM, RGN22NM, RGN22CD, year, n, n.l.slag, n.nearest.city.lag, 
                n.London.lag, City.of.Westminster, dist,) %>% 
  rename(London = City.of.Westminster)
```

### unregister_dopar function

This is helpful when the parallel loop collapses.

```{r}
unregister_dopar <- function() {
  env <- foreach:::.foreachGlobals
  rm(list=ls(name=env), pos=env)
}
unregister_dopar()
```

### train the model in all data

```{r}
#RNGversion("4.0.2")

set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "BUA22CD",
                                timevar = "year", 
                                k = 10,
                                seed = 123)

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # It should be RHS variables + 2

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

detectCores()
cl <- makePSOCKcluster(6)
registerDoParallel(cl)

start.time <- Sys.time()
model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                     #London.v + city.v + retail.v + n.l.slag + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

stopCluster(cl)

end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken #7.3m

print(model.all)

# Random Forest 
# 
# 134528 samples
#      6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 113520, 105938, 105938, 113505, 113505, 105938, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE     
#   2     variance    81.30945  0.6421819  23.39696
#   2     extratrees  96.94266  0.5158416  26.31072
#   4     variance    78.74613  0.6597365  21.00403
#   4     extratrees  88.39690  0.5849950  25.41510
#   6     variance    79.57023  0.6514796  20.86435
#   6     extratrees  86.15949  0.6003589  25.47657
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 4, splitrule = variance and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's wensite density, t-1",
                                               ifelse(rowname == "n.London.lag", "London's wensite density, t-1",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
                                                             ifelse(rowname == "n.nearest.city.lag", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_BUA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

### train the model in a loop for all but one regions

```{r}
# Drop the 142 BUA in more than one regions
df <- df %>% filter(!is.na(RGN22CD))

#rm(list = setdiff(ls(), c("df", "indices")))

#folds
k <- 10 #

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 10) # It should be RHS variables + 2

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# detectCores()
cl <- makePSOCKcluster(4)
registerDoParallel(cl)

# train in every region
start.time <- Sys.time()

regions <- df %>% ungroup() %>% distinct(RGN22CD) 
regions <- as.vector(regions$RGN22CD)

for (i in regions){

  print(paste0("start loop for ", i))

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN22CD!=i), 
                                spacevar = "BUA22CD",
                                timevar = "year", 
                                k = k)

  # CV
  tc <- trainControl(method = "cv",
                   number = k,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                   data = df %>% filter(RGN22CD!=i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data

  # file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
  # save(model.all, file = file)  # file name = region not in the training data
  rm(model.all)

  print(paste0("end loop for ", i))

}

# stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken
```

### Resamples for plots

```{r}
# create a list of the above model outputs
filenames <- list.files(paste0(path, "/outputs/rf/models/oa_rf/regions"), pattern = "*.RData", full.names = T)

for (i in filenames){
  load(file = i)
  i <- gsub("/home/nw19521/projects/web.diffusion/outputs/rf/models/oa_rf/regions/|.RData", "", i)
  assign(i, model.all)  
}

 fit.model.all <- mget(ls(pattern = "^region_"))

# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# clear up
rm(list = ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# resamples
fit.model.all.res <- resamples(fit.model.all)
summary(fit.model.all.res)

# plot metrics

# TODO: fix the year labels

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(fit.model.all.res, scales=scales)

# TODO: fix the year labels
```

### Variable importance

```{r , include=TRUE, results= 'markup', message=FALSE}
lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
```

```{r}
help <- lapply(fit.model.all, function(x) varImp(x)) 
help <- lapply(help, function(x) x[[1]])
help <- do.call("rbind", help)

importance <- help %>% rownames_to_column() %>% 
  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
  rename(importance = Overall) %>% 
  relocate(importance, .after = year) %>% 
  dplyr::select(-rowname)

importance %>% ggplot() +
  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
  scale_color_discrete(name = "Features") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("Importance") + 
  ggtitle("Feature importance") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

```

### Test on one region

```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}

# for a reference point
df %>% group_by(year) %>%
  summarise(min = min(n), max=max(n),
            mean = mean(n), median = median(n)) %>%
  round(2) %>% kable()
```

```{r}
# train on all but one region, test on that region

df.predictions <- data.frame()
for (i in regions){
  #for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all", i)], 
                    df[df$RGN22CD==i,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN22CD==i)) #%>% arrange(year, ladcd)

    d <- d %>% dplyr::select(RGN22CD, RGN22NM, BUA22CD, year, n, 1)

    colnames(d)[6] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  #}
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN22CD, ".", trained.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth

# source: https://www.data.gov.uk/dataset/aca8408d-e422-4ec2-a22f-9bd8beb1d549/built-up-area-to-region-december-2022-lookup-in-great-britain
path.lookup <- paste0(path, "/data/raw/Built_Up_Area_to_Region_(December_2022)_Lookup_in_Great_Britain.csv")

# 142 BUA in more than one regions
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN22CD, RGN22NM) %>% 
  distinct()

path.out <- paste0(path, "/outputs/rf/test_regions_BUA.csv")

rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN22CD")) %>% 
  rename(Region = RGN22NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(Region, Rsquared) %>% 
  write_csv(path.out)
```