### 2024-06-11 src/level_lisa_oa_sf.Rmd

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(sf)
library(magick)
library(spdep)
library(kableExtra)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>%  # for n == 1 
  ungroup()

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## spatial data

```{r}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- st_read(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew %>% 
  dplyr::select(OA11CD, Shape__Area) %>% 
  rename(id = OA11CD,
         area = Shape__Area)

path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- st_read(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc %>% 
  dplyr::select(code, SHAPE_1_Ar) %>% 
  rename(id = code,
         area = SHAPE_1_Ar)

# NI
path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
  rename(id= SA2011,
         area = Hectares)

# Build OA for UK
oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
rm(oa.ew, oa.sc, oa.ni)

# get UK
path.uk <- "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/Countries_December_2023_Boundaries_UK_BUC/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"
uk <- st_read(path.uk)
# spatial transformations
uk <- st_transform(uk, 4326)  # EPSG code for WGS84
```

## lisa maps

```{r}
path.out <- paste0(path, "/outputs/lisa/oa_ggplot/lisa_level_pc", n, "_oa") # for unique PC
#path.out <- paste0(path, "/outputs/lisa/oa_ggplot/lisa_level_pc", m, "_oa")  # for multiple PC

morani <- data.frame()

#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(oa.uk)
listw <- nb2listw(nb, style = "W", zero.policy = T)

for (t in 1996:2012){
  test <- oa.uk
  test <- test %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
    mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
           year = ifelse(is.na(year), t, year))

  #spatial weight matrix
  #moved out of the loop

  #Moran's I
  globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
  a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
  morani <- rbind(a, morani)

  #LISA
  lmoran <- localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude)

  quadr <- attr(lmoran, "quadr")
  #quadr[lmoran[,5] > 0.05 | is.na(lmoran[,5]), ] <- NA
  quadr[p.adjust(lmoran[,5], method = "fdr") > 0.05 | is.na(lmoran[,5]), ] <- NA

  palette_new <- grDevices::colorRampPalette(
            colors = c("blue", "lightsalmon", "lightblue", "red", "white"))(5)

  cities <- maps::world.cities %>%
    filter(country.etc=="UK") %>% 
    arrange(pop) %>% tail(10) 
  tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

  ggplot(data = oa.uk, aes(fill = quadr$median)) +
    geom_sf(data = uk, fill = "white", linewidth = .5) +
    geom_sf(linewidth = 0) +
    geom_sf(data = tif_sf, fill = "black", size = .3) +
    #geom_sf_text(data = tif_sf, aes(label = name, fill = "black"), size = 4) +
    scale_fill_manual(values = palette_new, na.value="white", drop = F) +
    theme_minimal() +
    theme(panel.grid.major = element_line(colour = "transparent"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.title=element_blank()) +
          #legend.position="none") +
    guides(fill = guide_legend(reverse = TRUE)) +
    labs(title = t,
        subtitle = "LISA, websites per OA")

  filename = paste(path.out, t, ".png", sep = "")
  ggsave(filename, bg='#ffffff')
}
```

Moran I

```{r}
path.out <- paste0(path, "/outputs/lisa/oa_ggplot/morani_oa.csv")
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  #kable() %>% 
  write_csv(path.out)
```