### 2025-01-03 paper/submission1/anonymised.qmd

---
title: "A multi-scale story of the diffusion of a new technology: the web"
abstract: |
    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time. This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales. It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment

#insert libraries here
library(rprojroot)
library(tidyverse)
library(knitr)
library(kableExtra)
library(sf)
library(cowplot)
library(patchwork)

# This is the project path
path <- find_rstudio_root_file()
```

# Introduction {#sec-introduction}

<!-- geographers lack of interest for diffusion + motivation -->
Economic geographers were always interested in how new technologies and
innovations diffuse across space and time with the work of 
@hagerstrand1968innovation, which demonstrated the importance of communications 
as a mechanism for the diffusion of innovations, being the most impactful. 
It was followed by various empirical studies -- see for instance @ormrod1990 
and @iso2005. However, the torch of exploring and modelling such diffusion 
processes had been passed to other disciplines such as economics, business 
studies and sociology well before the 'cultural turn' of economic geography
[@perkins2005international]. A potential explanation of the lack of
geographical studies exploring the diffusion of new and, more specifically for 
this paper, digital technologies across both space and time can be attributed to 
the scarcity of relevant and granular enough data, a problem also highlighted 
by @iso2005 and @kemeny2011international. As @zook2022mapping pointed, digital 
activities are hardly ever captured in official data.

<!-- contribution -->
This paper offers such a contribution: an economic geography study revealing
the importance of spatial mechanisms for the diffusion of an intangible, 
digital technology that is the Web in the UK at a high level of spatial 
granularity during the 1996-2012 period.
<!-- how a new technology that is the Web diffused over space and time  -->
It does so by employing a novel source of big data and machine 
learning algorithms to model the *active* engagement with Web during that 
period and tests how different spatial and temporal mechanisms shaped its 
diffusion. This paper also exemplifies how the combination of data sources 
and state-of-the-art research methods which escape the traditional social 
science domain can offer new lenses to geographical research regarding the 
understanding of technological diffusion.

<!-- motivation: stakeholders -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffuse over space and time and use this knowledge to make predictions
about the diffusion of related future technologies. As per
@leibowicz2016representing, historical studies agree that technologies diffuse 
differently in terms of times, rates, and geographies and can be driven
by related policies [@victor1993]. @meade2021modelling highlight
that a variety of actors have a direct interest in gaining such
knowledge including network equipment suppliers, network operators,
regulatory and local authorities. These processes and their effects vary
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places [@leibowicz2016representing]. Despite the spatial heterogeneity
of such diffusion mechanisms and the policy relevance, there is a scarcity 
of studies analysing the diffusion of new digital technologies at a detailed
geographical level because of the lack of detailed enough data and adequate 
research methods.

<!-- motivation: regional development -->
This gap becomes even more prominent if we consider the importance of 
technology and its adoption in economic geography and macroeconomic growth 
theories. Simply put, differences in technological adoption and sophistication 
may lead to uneven development as new technologies can increase productivity 
and allow firms to introduce new products [@solow1957technical; 
@aghion1990model; @kemeny2011international]. Linked to the subject of this 
paper, the Web -- and the active engagement with this technology that is 
creating and maintaining instead of just browsing a website -- can be 
considered as the underpinning technology for the transformative process 
identified by @capello2024nexus as the *digital service economy*. Importantly, 
the focus here is not the Web in general, but instead commercial websites.
After all, diffusion together with invention and innovation are considered the 
pillars of technological change [@das2022diffusion]. Findings from previous 
studies illustrated that early engagement with digital technologies such as 
the Web led to longstanding positive productivity gains for these regions that 
hosted early adopters [@tranosuk]. Hence, obtaining a robust understanding of 
the spatial diffusion mechanisms is crucial in mapping future development 
trajectories.

<!-- motivation: intangible/digital -->
Another line of motivation has to do with the intangible nature of the Web 
and the debate about the footloose nature of digital technologies, which
contrary to several earlier general purpose technologies are 'light in weight' 
[@coyle1999weightless]. Following the seminal work of @gaspar1998information, 
which demonstrated that technological improvements in telecommunications lead 
to increased demand for face-to-face communications and, consequently, 
enhance the importance of cities, empirical studies from economic geography 
and urban economics explored whether digital technologies substitute or 
complement agglomeration externalities and spatial structure 
[@kolko_death_2000; @ioannides2008effect; @tranos2021ubiquitous].
With the exception of @sinai2004geography, who focused, among other things, 
on web content and demonstrated that more such content is available in larger
markets, this strand of literature has been mostly concerned with the
spatiality of the Internet's hard and tangible infrastructure
and shied away from exploring the spatial dimensions of one of the 
most distinct software layers of the Internet infrastructure, the Web.

<!-- diffusion and adoption -->
Technological diffusion, which is by definition an aggregated process,
can be discussed in parallel with individual adoption mechanisms. On the
one hand, @rogers2010diffusion identifies early adopter of new
technologies as 'knowledgeable risk takers' and @griliches1957 as
'profit maximisers' [@ding2010modeling]. Such individual agents are
rewarded because of their attitude towards new technologies and
innovations. On the other hand, @perkins2011internet attribute diffusion
to two processes: (i) epidemic-like mechanisms, which are governed by
distance, proximity and social interactions, and (ii) economic
mechanisms as new innovations are adopted by users as they become more
profitable, valueable and useful.
<!-- The criticisms are mainly based on the fact that although the approach -->
<!-- gives an idea of aggregate (industry or household) behaviour, it does -->
<!-- not focus on the individual's (firm or household) adoption process -->

<!-- early engagement with the web -->
This paper focuses on the diffusion of the Web as new technology during
the 1996-2012 period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
Internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
9 per cent of UK's households had access to the Internet in 1998
[@ons2018], the web included mostly static webpages, there were no
social media and web browsing involved exclusively a desktop PCs
as there were no smartphones [@tranosuk]. Hence, it is fair to say that
the study period captures the very early stages of the diffusion of a
new technology that is the Web until its maturity. The former is a key 
point in the lifecycle of a new technology. Firstly, during this period new 
technologies are expensive, crude and imperfect [@rosenberg1994exploring; 
@wilson201281]. A simple comparison between Web 1.0 and Web 3.0 applications 
clearly illustrates this [@tranos2020social]: while a static website 
facilitates one-way information dissemination, a platform like `github` 
enables cooperation between users and the creation of new information, 
meaning, and even knowledge [@faraj2016special; @barassi2012does]. 
During this period the performance of a new technology is the main 
attraction and not the cost to access and use it [@wilson2011lessons]. 
There is a broader theoretical discussion in the literature about 
the early adoption motives. As summarised by @perkins2005international, 
on one hand, epidemic models highlight the role of interpersonal 
contacts as a way for new technologies to diffuse. On the other hand, 
economic models underline the importance of heterogeneity. Different firms 
have different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories [@neffke2011regions; @simmie2014new]. This
argument is even more relevant when the focus is on digital technologies 
because of the commonly found lag between investment and economic returns as 
reflected in the Solow paradox [@acemoglu2014return; 
@brynjolfsson2018artificial].

[^1]: <https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft>

<!-- data -->
The data used here depict the active engagement with the Web as they contain 
geolocated and time-stamped counts of commercial websites from the UK. Instead 
of adopting metrics reflecting the passive engagement with the Internet such as 
Internet adoption and speeds [e.g. @blank2018local; @destefano2022fuzzy], this 
paper observes yearly counts of geolocated websites, which reflect the creation 
and maintenance of websites. It needs to be highlighted here that although 
there are various types of underpinning web technologies -- namely markup, 
stylesheet, client- and server-side coding languages as well as different
web browsers and other related programming language -- the focus here
is on the diffusion of the Web in its totality. For a more detailed and 
cross-country technological diffusion study, see @PAPAGIANNIDIS2015308.

<!-- description of the methods and results -->
To analyse these data, a diverse set of methods are employed. Firstly, 
cumulative adoption curves are estimated for the UK as a whole and for the 
Local Authority Districts (LAD), illustrating, for the first time, such 
S-shaped diffusion patterns for local areas. This system level analysis also 
reveals insights about the volatility of the adoption of the Web. Then, 
exploratory spatial data analysis offers some first indications of the 
importance of the different spatial diffusion mechanisms in shaping the 
diffusion of the Web in the UK. Importantly, the analysis adopts two distinct 
geographical scales: the LAD and the most detailed scale of UK census areas, 
the Output Areas (OA). Lastly, a novel methodological framework based on 
Machine Learning (ML) algorithms is developed to test the 
role of such mechanisms at these two scales. The results reveal, for the first 
time, the importance of different spatial diffusion mechanisms in the diffusion 
of an intangible, digital technology at very granular geographical scales.

<!-- contents -->
The structure of the paper goes as follows. @sec-litreview reviews the key
concepts in the diffusion literature as well as relevant empirical studies. 
Then, @sec-datamethods describes the data and methods used in this paper. 
The results of the different stages of the analysis are presented in 
@sec-results. The discussion of the results and the conclusions of the paper
are offered in @sec-conclusions.

<!-- @grubler1990rise Later Hagerstrand conceptualized physical "barrier" -->
<!-- effects like lakes or uninhabited areas, which, in addition to distance, -->
<!-- act as further retarding effects on diffusion. These are formalized in -->
<!-- the form of "zero" or "half" contact multiplicators on the (distance -->
<!-- decaying) message flows. -->

<!-- @grubler1990rise With respect to the formalization of the communication -->
<!-- flows Hagerstrand defines a "mean information field" (MIF), in which the -->
<!-- probability of communication is a negative function of distance between -->
<!-- individuals -->

# Literature review {#sec-litreview}

<!-- key concepts -- copied from help.Rmd -->
Geographical diffusion is a synthesis of different processes. On the one
hand, a purely spatial or, in other words, contagious processes can be 
identified. Adjacency and, more broadly, distance are the key drivers of
diffusion. This perspective draws similarities with epidemics:
innovations just like pathogens spread because of contagion and,
consequently, proximity and exposure [@hivner2003facilitating]. On the
other hand, there is a hierarchical processes. According to this, instead of 
a horizontal distance-based diffusion mechanism, the top-down hierarchy
of urban systems shapes technological diffusion. In reality, the synthesis 
of these two processes better reflects how new technologies diffuse over 
space [@morrill2020spatial].

These ideas were firstly introduced by Torsten Hägerstrand and his
thesis entitled 'Innovation Diffusion as a Spatial Process'
[@hagerstrand1968innovation]. Hägerstrand was the first to identify
diffusion as a geographical process. The starting point was the idea
that diffusion is based on passing information through social networks,
which themselves tend to be defined by geography. Hence, he identified
a 'neighbourhood' mechanism of how information, and consequently,
innovation diffuse. He used agricultural innovations to test and model
his ideas using Monte Carlo simulations. Hägerstrand also incorporated
the role of hierarchy and how some phenomena maybe firstly adopted in
larger cities and then diffuse to second tier ones. This is a sequential
instead of a simultaneous process, which resembles the 'lead-lag'
spatial acceleration effect in market research [@bento2018time;
@PERES201091]. Hägerstrand is more widely known though for highlighting
the role time plays in the diffusion of innovations: an early-pioneering
period, a middle fast accelerating period and a final saturation one
[@morrill2020spatial].

The temporal dimension was further explored by Everett Rogers and his
seminal work on 'Diffusion of Innovations' [@rogers2010diffusion].
Rogers being a sociologist focused not on the diffusion of
innovations over space and time, but instead on the adoption of new
technologies and innovations by individuals and the individual
mechanisms that drive the decisions behind adoption. He identified five
groups of individuals regarding their adoption speed: innovators, early
adopters, early majority, late majority and laggards. The key mechanism of
diffusion and adoption is communication and how knowledge is transferred 
within a social system. All above approaches agree that diffusion and the
cumulative adoption of a new technology are represented by an S-shaped 
line plot [@grubler1990rise].

Schmidt's Law empirically illustrates a similar pattern. *Core* and
usually highly agglomerated regions is where new technologies are
invented and commercially deployed [@grubler1990rise].
This is where the first adopters tend to be based. Then, technologies
spread to the *rim* and eventually to the *periphery*. Although adoption
pace might be higher when new technologies finally arrive to the
periphery, the saturation levels there may never reach the ones in the
core because of the lack of infrastructure or other necessary
institutions [@leibowicz2016representing].

@grubler1990rise effectively summarises the three key spatial diffusion 
mechanisms: (i) the cumulative level of adoption follows an S-shaped
pattern just like purely temporal models; (ii) diffusion is shaped by a
hierarchical mechanism in a form of a centrifugal force: from core to
periphery; and (iii) diffusion is also shaped by distance and proximity. 
These are the three mechanisms that the empirical analysis in @sec-results 
investigates.

<!-- previous studies -->
The remaining of this section reviews empirical studies which analysed
the diffusion on new technologies over space and time. Although the
spatial dimension is present in most of the following studies, the level
of spatial detail is always more coarse than the one adopted in this
paper. @beardsell1999spatial studied the evolution of the computer
industry in 317 US metro areas during the 1977-1992 period using
employment data. Their analysis indicated that the relative size
distribution holds for urban computer employment and also urban
heterogeneity is essential in explaining this distribution. In a recent
study, @bednarz2020pulled focused on wind turbines and modelled their
spatial diffusion across 402 German regions during 1970-2015. Their
key finding is that local demand than local supply was the main driving
factor. @haller2011determinants employed firm panel data for Ireland
during 2003-2005 and illustrated that when firms are located in a
region or industry with high share of firms having a website, they are
more likely to have a website too.

At a global scale @perkins2005international explored whether the
diffusion rate of new technologies is driven by a latecomer advantage
and the engagement with the global economy via foreign direct
investments and trade. Their results illustrate that indeed latecomers
and developing countries experience diffusion of new technologies more
rapidly than early adopters and developed countries. At the same scale,
@perkins2011internet explored whether the adoption of previous
communication technologies that is mail, telegrams and telephones were
shaped by similar socioeconomic factors as the Internet. Their results
indicated common patterns regarding the drivers behind the adoption of
different communication technologies.

Turning to studies that share more technological and scalar similarities
with this paper, @ding2010modeling modelled the spatial diffusion of
mobile telecommunications across regions in China. Their analysis
indicated that socioeconomic characteristics are important determinants
of the timing, speed and the level of mobile diffusion within China.
Using data from a Hungarian online social network, @lengyel2020role
analysed its adoption and the churn at a very granular spatial level.
Their results were in agreement with the early theoretical and empirical
contributions reviewed here: assortativity, urban scaling and distance
are the key drivers of spatial diffusion. @bakher2013diffusion compared 
different the growth curves as a means to model diffusion of online 
shopping in Australia during the 1998-2009 period. At a global scale
@PAPAGIANNIDIS2015308 modelled the diffusion of different web
technologies and practices across countries. Using similar but less extensive 
data as the one used here, their analysis illustrated how the diffusion 
of different web technologies and practices follow an S-shaped pattern 
as well as the different diffusion rates of the different technologies 
and practices.

All in all, this paper tests whether the three distinct spatial diffusion 
mechanisms shaped the diffusion of the Web in the UK at very granular
spatial scales from the Web's early stages onwards. The next section discusses
the empirical framework.

<!-- @fritsch2015new analyze the spatial diffusion of laser technology -->
<!-- research in West Germany from 1960, when this technology began, until -->
<!-- 2005. -->
<!-- in one region  -->
<!-- @leibowicz2016representing for energy There is no generally accepted -->
<!-- theory that explains diffusion rate heterogeneity across technologies, -->
<!-- but several factors are considered important. Greater unit scale and -->
<!-- larger market size contribute to slower diffusion. Requirements for -->
<!-- interrelated technologies or complex infrastructures also hinder the -->
<!-- diffusion process (Grubler, 2012). -->

<!-- Mobile phones benefited from early deployment in recreational boats and -->
<!-- automobiles, where the traditional competitor was not a viable option. -->
<!-- In the early stages of diffusion, performance is a more important driver -->
<!-- of adoption than cost competitiveness. Typically, significant cost -->
<!-- reductions only occur once the technology reaches a deployment level -->
<!-- capable of supporting standardization and mass production (Wilson, -->
<!-- 2012). -->

<!-- @leibowicz2016representing Empirical evidence supports the validity of -->
<!-- Schmidt's Law over a wide range of technologies, time periods, and -->
<!-- geographical contexts. A recent meta-analysis of technology up-scaling -->
<!-- found that diffusion accelerated moving from the core to the rim and -->
<!-- periphery for technologies as diverse as natural gas power, oil -->
<!-- refineries, and automobiles (Wilson, 2009). One historical example that -->
<!-- conforms particularly well to Schmidt's Law is the diffusion of coal -->
<!-- power in Europe (Grubler, 2012). England emerged as the core region for -->
<!-- coal power because it had legal and economic institutions that -->
<!-- incentivized scientific pursuits, domestic coal reserves, and a clear -->

<!-- industrial motivation to replace water power with coal. ... -->

<!-- country level, not sure if needed. -->

<!-- @bento2018time explore What determines the duration of formative phases -->
<!-- for energy innovations in different markets? We are interested both in -->
<!-- initial markets (also: core, lead, first mover, early adopter) where -->
<!-- formative phases prepare technologies for mass commercialization, and in -->
<!-- follower markets (also: periphery, lag, late adopter) where accelerated -->
<!-- formative phases may benefit from diffusion and spillovers. -->

# Data and Methods {#sec-datamethods}

<!-- main metric -->
To capture the diffusion of the Web, a website density metric is developed 
for two different geographical scales: the Local Authority Districts (LAD) 
and the Output Areas (OA). The former is an administrative unit and there 
are 374 LAD in the UK. The latter is the smallest census-based geographical 
unit and there are c. 230,000 of them in the UK. This methodological choice 
will allow the mapping of the diffusion of web technologies and the assessment 
of the diffusion mechanisms at these two very different spatial scales.

<!-- web data -->
The counts of websites at these scales are calculated using data from the 
Internet Archive[^2] and, specifically, the JISC UK Web Domain Dataset 
[@ukwebarchive]. The Internet Archive is one of the most complete and oldest 
archive of webpages in the world operating since 1996 
[@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which discovers 
webpages by following the hyperlinks of every webpage it archives.
This dataset, which is curated by the British Library, contains all
the archived webpages from the UK country code top level domain (ccTLD - .uk) 
from the 1996–2012 period. In essence, this is a long list of 2.5 billion URLs 
of archived webpages including also the archival timestamp.

[^2]: [https://archive.org/](See%20https://archive.org/).

Instead of using the whole .uk ccTLD, this paper focuses on its commercial 
subset, the .co.uk second level domain (SLD). This choices decrease the 
heterogeneity of the web data as such commercial websites have specific 
aims: they are used to diffuse information, support online transactions 
and share opinions [@THELWALL2000441; @blazquez2018big]. 
Although a UK company can adopt a generic TLD such as .com and these 
cases escape the data used here, such omissions should not affect the 
validity of the results given the popularity of the .uk ccTLD [@tranosuk]: 
UK consumers prefer to visit a .uk website when they are searching for 
products or services [@hope]; and anecdotal evidence indicates that during 
the first half of 2000, three .co.uk domains were registered every minute 
[@oecd_coms]. Importantly, previous studies illustrated that .co.uk is the 
most popular UK SLD [@tranosuk].

The text from these webpages was scanned using a regular expression (regex) to
identify strings of text which resemble UK postcodes and one fifth of them 
included a mention to a postcode [@BL2013geo]. This information allows the 
geolocation of the data and the creation of the LAD and OA counts.

The data cleaning process includes an aggregation step, through which
the archived webpages are aggregated to the parent websites. This website 
reconstruction process lend itself to the construction of *website* instead 
of *webpage* density metrics. Websites tend to represent specific organisations 
or entities, and, arguably, are more meaningful observational units than 
webpages, which can ignore the upstream dependency of the website they belong 
to. Based on the following example, all three webpages are part of the 
same website ([http://www.website.co.uk](http://www.website.co.uk)). Only 
the overall website and not the nested webpages are considered for the counts 
as otherwise the density metrics would have been biased towards large, place 
specific, websites.

  - [http://www.website.co.uk/webpage_a](http://www.website.co.uk/webpage_a) B15 2TT

  - [http://www.website.co.uk/webpage_b](http://www.website.co.uk/webpage_b) BS8 1TH

  - [http://www.website.co.uk/webpage_c](http://www.website.co.uk/webpage_c) B15 2TT

\noindent What is challenging is that this aggregation approach,
which has been used elsewhere [@tranosuk; @shoreditch] may lead to websites with 
multiple postcodes. As per the above example, 
[www.website.co.uk](http://www.website.co.uk) includes two unique postcodes: 
B15 2TT and BS8 1SS. The distribution of the number of postcodes per website for 
2000 is presented in Table \ref{f2000}, which illustrates a wide range. 
At the left end of the distribution, there are websites anchored to a unique 
location (72% of all the reconstructed websites in 2000), which may represent a 
small company with a single trading location. At the right end, there are 
websites with thousands of different postcodes. Considering the time period of 
the analysis, such cases can represent directories which used to be popular 
in the pre-search engines early times of the commercial Internet as well as real 
estate websites [@tranosuk].

The analysis is firstly conducted on websites, which only contain one unique 
postcode. As a robustness check, the analysis is replicated for an extended 
subset of websites, which include up to 10 unique postcodes to capture 
commercial websites which point to multiple locations. They are geolocated 
by equally attaching them to these locations. This extended sample includes 94 
percent of all the archived websites in 2000.

```{r eval=T, echo=FALSE, results='asis'}

# load 2000 co.uk
# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
path.2000 <- paste0(path, "/data/temp/all2000couk.csv")

all2000.duplicates <- data.table::fread(path.2000) #
# dim(all2000.duplicates) # 3336162
# it includes duplicated URLs: if one webpages includes multiple postcodes
# then it appears multiple times. This is ok for the nuts aggregation, but
# not for the frequencies
# This is only .co.uk

# one line for every host
all2000 <- unique(all2000.duplicates, by = c("host")) # 57897

# unique postcodes per website f table
f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
                                                              "(100,1e+03]", "(1e+03,1e+04]",
                                                              "(1e+04,1e+05]"))
levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")

f.websites.pc <- f.websites.pc %>% 
  dplyr::select(-cumfreq) %>% 
  rename(Postcodes = level,
         F = freq,
         'F (%)' = perc,
         'Cummulative F' = cumperc)

kable(f.websites.pc,
      format = "latex",
      digits = 3,
      booktabs = T,
      format.args = list(big.mark = ","),
      caption = "Number of unique postcodes per .co.uk website, 2000.\\label{f2000}") %>%
  kableExtra::footnote(general = "Tranos et al. 2021",
                       general_title = "Source: ",
                       footnote_as_chunk = T)  
   # kable_classic(full_width = F, html_font = "Cambria")
```

Another data cleaning step deals with some extreme outliers. Figure 
\ref{correct} plots the website counts for the top 1000 postcodes. Some obvious 
outliers can be observed for the 2002-2006 period for a handful of postcodes, 
which can be attributed to link farms [@BL2013links]. The website counts for 
these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF), which 
in 2004 or 2005 had more than 1000 websites pointing to them, were replaced 
with predicted values based on a simple panel regression model with postcode 
fixed effects and yearly dummy variables. These postcodes refer to a residential 
area with a small park in South London, residential areas outside Warwick and 
Manchester, a rural area in the south border of England and Wales and small 
business parks in Crew and Derby. Hence, it is difficult to believe that these
postcodes genuinely hosted such a large amounts of websites, which were later 
extensively declined. To put the magnitude of the data imputation into 
perspective, this process affected 6 out of the 557,808 postcodes included in 
the data. As the @sec-results illustrates, this led to a small increase of the 
model predictive power.

```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
path.image1 <- paste0(path, "/outputs/pc_year_1.png")
path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)

par(mar=rep(0,4)) # no margins

p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
```

To create the website density metric, the yearly website counts at the LAD level 
are standardised by the number of firms in LAD to avoid biases associated with 
LAD hosting a large number of firms. Given that there is no such readily 
available statistic for the OA, the actual OA level counts are used. However, 
because of the the consistent spatial definition of OA (they host 40-250 
households),[^4] website counts in OA are interpreted as a density metric too.

[^4]: [https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies](https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies)

<!-- system-level analysis -->
These website density metrics are used for the three different stages of the 
analysis. Firstly, a system-level analysis explores how the diffusion of web 
technologies in the UK fits with the well-established S-curve of cumulative 
adoption. To do so, the following logistic function (Equation \ref{eq:s}) is 
estimated for the whole of the UK and for each LAD separately:

\begin{align}
y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
\end{align}

\noindent $k$ is the asymptote or, in other words, the saturation level, $b$ 
the overall growth rate, and $t_{0}$ the *inflection point* of maximum growth 
at $k/2$, where the logistic function is symmetrical [@wilson201281]. 
The $t_0$ of each LAD is compared against the $t_0$ of the UK to delineate 
whether a LAD reached that point faster or slower than the country average. 
Importantly, an accuracy criterion was imposed and only LAD with $R^2 > 0.9$ 
were included in this analysis. To estimate Equation \ref{eq:s} a self-starting 
logistic growth model was employed using the `nls` and `SSlogis` functions in 
`R`. This logistic growth model depicts the initial period of slow diffusion of 
a new technology, which is followed by the fast, exponential growth stage to
then eventually slow down and saturate [@wilson201281; @grubler1999dynamics].

<!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->
<!-- The literature usually uses the saturation level as the asymptote. I am -->
<!-- using the total number of websites as we cannot compute a rate. -->

The system-level analysis also focuses on the volatility of the web adoption to 
depict places with high concentration of early adopters and, equally, 
latecomers. To do so, the change of the ranking of the UK LAD over time is 
plotted and discussed. Both the S-curve and the volatility analysis focus only 
on LAD as the very large number of OA would have made such analysis difficult 
to visualise and interpret.

<!-- ESDA -->
Next, exploratory spatial data analysis depicts whether the two main 
drivers of spatial diffusion -- namely neighbourhood and hierarchy -- underpin 
the diffusion of the Web in the UK. To capture the former and following 
@ding2010modeling the Moran's I and the Local Index of Spatial Association 
(LISA) are estimated for the website density. To address the hierarchy mechanism, 
the Gini coefficient -- an established metric of inequality -- is 
calculated. All the above are computed and plotted longitudinally both for the 
LAD and the OA.

<!-- RF -->
Lastly, a modelling framework is developed to test the above diffusion 
mechanisms. The overarching aim is to build a model that can test how well
these mechanisms can predict website density:

\begin{align}
Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
\end{align}

\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest (RF).
This is a popular ML algorithm for both regression and classification
problems [@biau2012analysis]. It was introduced by @breiman2001random and has 
become a go-to data science tool. RF can effectively handle skewed distributions 
and outliers, model non-linear relationships, require minimal hyperparameter 
tuning, exhibit low sensitivity to these parameters, and have relatively short 
training times [@Caruana2008; @liaw2002classification; @yan2020using]. These 
attributes match well with the website density data characteristics including 
skewness especially for the OA. Also, the large data size (c. 230k data points 
for each of the 17 years) calls for fast training times. Importantly, RF 
predictions tend to be more accurate than those from single regression trees and 
outperform Ordinary Least Squares in out-of-sample predictions, even with 
moderate-sized training data and a small number of predictors 
[@mullainathan2017machine; @athey2019machine; @sulaiman2011intelligent; 
@pourebrahim2019trip; @biau2012analysis].

RF is a tree-based ensemble learning algorithm [@breiman2001random]. It begins 
by generating random samples of the training data, which are then used to grow 
regression trees to predict the dependent variable. Data points and predictors 
are randomly sampled for the different trees. The trees are trained in parallel 
using their own bootstrapped samples of the training data. A crucial feature of 
RF is their ability not to overfit, meaning they can generalize well to unseen 
test data. While each tree may overfit individually, the ensemble of trees does 
not because the errors of individual trees are averaged, reducing the overall 
variance and preventing overfitting [@last2002improving]. For regression 
problems, RF predictions are made by averaging the predictions of all decision 
trees.

RF have been widely employed to address regression research problems. 
@pourebrahim2019trip combined a spatial interaction modelling framework with ML 
algorithms including RF to predict commuting flows in New York City. 
@sinha2019assessing advocated for adopting spatial ensemble learning approaches, 
such as RF, to model spatial data with high autocorrelation and heterogeneity. 
@creditspatial predicted employment density in Los Angeles using spatially 
explicit RF. @guns2014recommending used RF to build a recommendation system for 
research collaborations. @ren2019predicting trained RF to predict the 
socio-economic status of cities using various online and mobility predictors. 
@tranos2023using utilised hyperlinks data and RF to make out-of-sample 
predictions of interregional trade. @zhou2023geography employed such a 
framework to assess whether key predictors of obesity differ across English 
cities.

# Results {#sec-results}

## System-level analysis

<!-- S-shaped -->
```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Grwoth curve, UK"}
path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
knitr::include_graphics(path.image)
```

Figure \ref{s_uk} plots the cumulative adoption of the Web in total in the UK 
during the 1996-2012 period. Not surprisingly, the cumulative adoption is well 
represented by a curve that resembles the figure S. The vertical 
line around year 2003 illustrates the point where the modelled cumulative 
adoption is equal to 50% of the maximum. This $t_0$ *inflection* point signals 
the maximum adoption speed and is used here to determine whether a UK LAD 
reached that inflection point earlier or later than the UK average. 
Specifically, the S curve and the inflection point are estimated for every LAD 
individually and then compared against the UK average. LAD that reached their 
inflection point earlier than the country average are labelled as *fast* and 
the rest as *slow*. Figure \ref{s_map} maps this pattern. There is a clear high 
concentration of fast adopting LAD in the South and East of England as expected. 
However, there is also such a concentration in Wales, Scotland and the North 
West of England. At a more detailed level, there are some expected 
examples of LAD with relevant industrial backgrounds delineated as fast: the 
City of London, a world-renowned cluster of finance industries [@cook2007role], 
and Reading, a town with high-tech service industries in proximity to London 
and its main airport, Heathrow [@england2005polynet]. However, there are also 
LAD which were expected to appear as fast -- e.g. Hackney in central London and 
Bristol, a well-established creative cluster [@oatley1999cultural; 
@bassett2002cultural] -- but were delineated as slow. Nevertheless, there is 
obvious spatial clustering of fast and slow LAD. Importantly, out of the ten 
fastest LAD, nine were located in the South East of England and London and one 
in Cambridge (see Table \ref{table.s.lads} in the Appendix). The above 
observations remain almost unchanged when the analysis includes websites 
with up to 10 unique postcodes -- see Figures \ref{s_uk10} and \ref{s_map10} in 
the Appendix -- advocating towards the robustness of the analysis.

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD adoption rates"}
path.image <- paste0(path, "/outputs/s/speed_map.png")
knitr::include_graphics(path.image)
```

<!-- discussion -->
The S-curves of the cumulative adoption of Web in the UK in total and in the 
UK LAD demonstrate a pattern well aligned with all previous studies discussed 
in @sec-litreview. Importantly, the above findings are in line with previous 
studies exploring the diffusion of web technologies [@PAPAGIANNIDIS2015308], 
social media [@lengyel2020role] and online shopping [@bakher2013diffusion]. 
Similar to @beardsell1999spatial, who analysed the spatial evolution of 
computer industry employment across 317 US metro areas, spatial heterogeneity
is evident in the diffusion speed when the scale of analysis is as detailed
as the LAD. Importantly, no previous studies estimated such cumulative 
adoption curves for areal units as small as the UK LADS.

<!-- rankings -->
The next step assesses the stability and volatility of the LAD in terms of 
their adoption of web technologies. As highlighted in the literature 
[@risk_perceptions], different agents have different perceptions about and 
levels of acceptance of the risks and the potential economic returns associated 
with the adoption of new technologies -- see for instance the seminal work of
@venkatesh2000theoretical. To reveal such aggregated patterns, Figure \ref{rank} 
plots the ranking of UK LAD based on website density. To decrease noise, the 
average ranking of 1996-1998 and 2010-2012 is plotted instead of the individual 
years. The colours depict the 5th and 95th percentile of the absolute 
difference of the LAD ranking between 1996-1998 and 2010-2012. While there are 
quite a few obvious cases of LAD that maintained their position between the 
beginning and the end of the study period both at the top or at the bottom of 
the hierarchy, there are also quite a few LAD that changed drastically 
their position. Some of these LAD enjoyed a process that at the first instance 
looks like *leapfrogging* since they managed to jump at the top of the hierarchy 
despite their slow start. There is extensive literature regarding the potential 
benefits of technological leapfrogging. The underpinning argument is that 
latecomers can adopt and benefit from new technologies that have been developed 
elsewhere without incurring the hefty initial R&D costs [@teece2008firm].
Although the leapfrogging literature does not pay much attention to cities and 
regions [@yu2018sustainability], previous research highlighted the long term 
and sustained productivity benefit of the early adopters of web technologies 
in the UK [@tranosuk]. So, although the LAD system is volatile, it is not clear 
whether the LAD with high concentration of late adopters will gain any 
latecomer-related benefits in a way similar to countries experiencing 
technological leapfrogging. Still, the opportunity to catch up that 
@perkins2005international illustrated at the country level, is also visible at 
the much more detailed LAD scale. As previously, this is consistent with 
including websites with up to 10 postcodes (see Figure \ref{rank10}).

```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of wed diffusion (LAD)"}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
knitr::include_graphics(path.image)
```

## Exploratory analysis

<!-- Neighbourhood effect -->
Figures \ref{morani} and \ref{lisa} offer a first insight into whether a 
neighbourhood mechanism underpins the diffusion of the Web in the UK as they 
present the Moran's I and the LISA maps of website density respectively for LAD 
and OA. Starting from the former, spatial autocorrelation was higher in the 
beginning of the study period and then after 2000 dropped slightly and 
stabilised around 0.2. This reflects the early concentration of high website 
density around London, which over time diffused as high website density clusters 
can be seen in other parts of the country away from London (Figure \ref{lisa}). 
An almost reverse pattern can be observed for OA. At the beginning of the study 
period Moran's I was around 0.5 and it plateaued after 2000 around 0.2. 
Because of the very small size of OA, at the early stages of the diffusion 
of web technologies their adoption was spatially scattered. This is reflected in 
the lack of any significant clusters in 1996 (Figure \ref{lisa}). Eventually, as 
the adoption rate increased, more such clusters of high website density were 
formed and this is reflected both in the Moran's I and the LISA maps in Figures 
\ref{morani} and \ref{lisa}. A similar pattern is observed for the extended 
dataset (see Figures\ref{morani10} and \ref{lisa10}). In this case, the 
magnitude of spatial autocorrelation is slightly higher illustrating a stronger 
spatial clustering mechanism. All in all, the exploratory spatial data analysis 
advocates towards an underpinning neighbourhood mechanism and the different 
scales of analysis illustrate how it evolved differently over time. 
<!-- discussion -->
Interestingly, the magnitude of Moran's I is similar to the ones revealed
by @ding2010modeling about mobile phone adoption in China albeit the much more 
detailed spatial scale of this paper.

```{r, morani, echo=FALSE, message=FALSE, fig.cap="\\label{morani}Website density Moran's I"}

path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa.csv")

la.m <- read_csv(path.in.la) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'LAD')
oa.m <- read_csv(path.in.oa) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'OA')

bind_rows(la.m, oa.m) %>% 
  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Moran\'s I') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa}\\centering Website density LISA maps", out.height="80%"}

path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la1996.png")
path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2000.png")
path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2012.png")
path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa1996.png")
path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2000.png")
path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2012.png")
path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)  
p3 <- ggdraw() + draw_image(path.image3)  
p4 <- ggdraw() + draw_image(path.image4) 
p5 <- ggdraw() + draw_image(path.image5)
p6 <- ggdraw() + draw_image(path.image6)  
p7 <- ggdraw() + draw_image(path.image7)

par(mar=rep(0,4)) # no margins

library(patchwork)
# (p1 | p4)/(p2 | p5)/(p3 | p6)/ p7 +
#   #plot_layout(widths = 1)
#     plot_layout(widths = c(1, -0.5))

# p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 / p7 +
#   plot_layout(ncol = 3, widths = c(1, -0.85, 1), heights = c(1,1,1,2))

p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))

# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
```

To illustrate whether a hierarchical process also underpins the diffusion of web 
technologies, the Gini coefficient is calculated yearly both for LAD and OA. As 
a metric of inequality, the Gini coefficient demonstrates whether website 
density is concentrated in a small number of LAD or OA, or whether it is more 
equally spread across the country. Both scales of analysis in Figure \ref{gini} 
illustrate the same picture. At the beginning of the commercial Internet, 
website density was extremely unequal, or, in other words, only a few places had 
websites associated with them. Inequality dropped and plateaued after 2000 for 
both scales. This is illustrative of a hierarchical diffusion mechanism that 
led over time to a more equal spread. Interestingly, the year 2000 is again a 
period of change for this diffusion mechanism as it was for the neighbourhood 
process. There is a substantial difference between the Gini coefficient 
magnitude for LAD and OA, but this is expected as the very small size of OA 
equates to a lot of polygons without any websites pointing to them -- for 
example residential OA.

```{r, gini, echo=FALSE, message=FALSE, fig.cap="\\label{gini}Website density Gini coefficient"}

path.in.la <- paste0(path, "/outputs/gini/gini_la.csv")
path.in.oa <- paste0(path, "/outputs/gini/gini_oa.csv")

la.gini <- read_csv(path.in.la) %>% 
  mutate(spatial.unit = 'LAD')
oa.gini <- read_csv(path.in.oa) %>% 
  mutate(spatial.unit = 'OA')

bind_rows(la.gini, oa.gini) %>%
  rename(Gini = 'gini(a$n)') %>% 
  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Gini') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

## Modelling outputs

The next section incorporates the above mechanisms of the spatial
diffusion of the Web into a modelling framework. The aim is to use variables 
depicting these spatial processes in order to predict the diffusion of the Web 
as a new technology in the UK over space and time and across different scales. 
Specifically, four different models are estimated. Firstly, all the data points 
for the OA and LAD are utilised in order to build two RF models and assess their 
capacity to predict the adoption of the Web at the two different scales. 
These two models will reveal the predictive capacity of the diffusion mechanisms 
and also demonstrate how the importance of such variables changes across different 
scales. The next two sets of models are again trained at these two scales: 
OA and LAD. However, instead of using all the data points, the OA and
the LAD from one of the twelve UK regions are held out for the model training.
Then, the trained model is used to predict website density for the OA or the LAD 
of the held-out region. This process takes place recursively for all twelve UK
regions. The regional differences in the predictive capacity of the different
samples will reveal how dissimilar these spatial process are across regions 
and, importantly, at different scales.

It needs to be highlighted here that the cross validation (CV) for all models
is spatially and temporally sensitive. Instead of using
10 random and space- and time-agnostic folds,the `CAST` package is employed 
which allows holding back data points from specific years and spatial units 
and use them for testing in order to estimate the model performance 
[@meyer2018improving].

The models include variables that capture the three processes that the relevant 
literature and the descriptive analysis highlighted. Namely, the models capture: 
(i) a hierarchical mechanism with diffusion running from main centres to 
secondary ones, (ii) a neighborhood mechanism according to which diffusion 
first hits nearby locations, and (iii) the rather canonical pattern of 
diffusion over time as reflected in the S-shaped pattern in the cumulative level 
of adoption.

To capture the hierarchical mechanism the models include as predictors a one
year lag of website density in London, the largest city in the UK, a
one year lag of the website density in the nearest city and the same for
the nearest retail centre. Due to the small sizes of the retail centres,
the latter is only relevant for and included in the OA-level models. In 
addition, the models include the distance to London, the nearest city and the 
nearest retail centre. The underlying logic is that the level of website
adoption in a spatial unit depends on the level of the adoption in
places further up in the urban hierarchy the previous year. The inclusion of
the distance variables incorporates spatial structure into the hierarchy 
argument. To depict the neighbourhood mechanism, the website density of the 
neighbouring spatial units in the previous year is employed. This is calculated 
using rook continuity. Again, the underpinning rationale is that the level of 
web adoption within a spatial unit depends on the level of web adoption in the 
neighbouring spatial units the year before. This represents the 'hitting nearby 
locations first' argument. Therefore, the spatial and temporal lag of the 
website density in LAD and OA is calculated. Lastly, the temporal pattern of 
the cumulative adoption is captured by a time trend variable. Hence, all four 
models will follow the following generic form (Eq. \ref{model}):

```{=tex}
\begin{align} \label{model}
Website\,Density_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}
```

To assess the predictive power of the model, three broadly utilised
metrics are employed: the coefficient of determination ($R^2$), mean
absolute error (MAE) and root mean square error (RMSE):

```{=tex}
\begin{align}
R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
\end{align}
```
```{=tex}
\begin{align}
MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
\end{align}
```
```{=tex}
\begin{align}
RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
\end{align}
```

\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which consists of
$N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$ predicted value
for the dependent variable and $\overline{y_{k}}$ is the average value
of $y$. The last two metrics are expressed in the same units as the
dependent variable -- websites per firm for the LAD modes and the number
of websites for the OA models -- while the first one is the coefficient
of determination between the observed and the predicted values of
website adoption. Regarding $MAE$, it is the absolute difference between
the observed and the predicted website adoption. While $MAE$ does not
penalise for large errors, $RSME$ does, as it is proportional to the
squared difference between the observed and the predicted trade flows.
Hence larger errors weigh more for $RMSE$ [@pontius2008components].

@tbl-model-metrics presents the model performance for the first set of models, 
for which all data points are employed for training and testing via CV. The 
first one is trained and tested on 374 LAD and the second on 
232,296 OA, both for the 16 year period (1997-2012). The results are remarkably 
good considering that the are the outcome of space and time sensitive 
CV, so the the model does not suffer from overfitting. At the LAD level the 
model predicts 85% of the variation of website density. Both error metrics 
indicate that the model error is $(\frac{1}{20}, \frac{1}{30})$ of a website per 
firm. At the OA the $R^2$ drops down to 32%. Considering its granularity, this 
is still a remarkable performance. To contextualise it, the model results in a 
$MAE$ equal to one website for areas small enough to host less than 140 
households.
^[According to the Office for National Statistics, 80% of OA in England and 
Wales host 110-139 households,
[www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).]
Because of the small size of the spatial units, the distribution is highly skewed
and a significant part of them is not linked to any websites. In 1997 only 1% of 
the UK OA were associated with at least one website. This should not come as a 
surprise as this was the very beginning of the commercial Internet and any 
activities with a digital footprint were concentrated in a handful of areas. 
This was clearly illustrated in Figure \ref{lisa}. At the end of the study 
period almost half of the UK OA were not associated with a website. Again, 
given the granularity of the data this should not come as a surprise.

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.028 | 0.850    | 0.019 |
| Output Areas      | 3.284 | 0.320    | 1.034 |

: Model metrics {#tbl-model-metrics}

The results in @tbl-model-metrics are robust against different specification. 
As noted earlier, the data imputation process resulted in a modest improvement 
of the predictive capacity of the model. As per
@tbl-model-metrics-no-correction all accuracy metrics are only slightly worst 
when the data imputation is not applied. As expected, the improvement is 
higher for the more detailed scale of analysis. Then, @tbl-model-metrics-10 
presents the model results for the extended dataset, which includes websites 
with up to 10 unique postcodes. While the results for the LAD are slightly 
less accurate, the $R^{2}$ of the OA model is more than double in comparison 
to the base results in @tbl-model-metrics. All in all, the modelling exercise
revealed how well can the three diffusion mechanisms predict 
the diffusion of the Web as a new technology. Importantly, this predictive
power is highly robust against against different scales and delineations
of the diffusion of the web.

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.032 | 0.810    | 0.019 |
| Output Areas      | 5.000 | 0.205    | 1.047 |

: Model metrics without data imputation {#tbl-model-metrics-no-correction}

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.188  | 0.77    | 0.122 |
| Output Areas      | 22.467 | 0.478   | 5.681 |

: Model metrics for the extended dataset {#tbl-model-metrics-10}

Figure \ref{var.imp} plots the importance of the different predictors. When the 
focus is on LAD, the website density in the nearest city, in London and in the 
neighbouring LAD the year before are the most important predictors. They are 
followed by the yearly trend, while the spatial configuration as reflected
in distances to London or the nearest city only plays a minor role. This can be 
attributed to the relatively coarse spatial scale of LAD. Nevertheless, all 
previously discussed spatial processes are at play in the diffusion of web 
technologies at the LAD level: the first two predictors depict the hierarchical 
mechanism, the spatial and temporal lag of website density the neighbourhood 
mechanism, and the yearly trend the time-sensitive cumulative adoption pattern.

When the much more granular scale of OA is adopted, the picture is almost 
reversed. The two most important predictors are the distance to London and to 
the nearest city followed by the spatial and temporal lag of website density 
and the distance to the nearest retail centre. They still depict a hierarchical 
mechanism, but proximity to the different population centres is more important 
than their lagged web densities in predicting website diffusion. The 
neighbourhood mechanism is still strong but less important at this scale. What 
is interesting is the almost negligible role of the yearly trend and London's 
website density. While the former probably illustrates the large heterogeneity 
in how the Web has been adopted at this very fine scale, the latter highlights 
that the importance of past web adoption rates in large population centres is 
surpassed by proximity to them and spatial configuration at this scale.

The above pattern of variable importance is largely in accordance with the 
models estimated for the extended dataset. A comparison with Figure 
\ref{var.imp10} illustrates that the main difference is the importance of the 
neighbourhood mechanism for the OA as the space and time lag of the website 
density is by far the most important predictor when the extended dataset is 
employed. But still, the variable importance patterns for the two datasets are 
not dissimilar.

```{r varimpLA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.LAD}Variable importance, LAD"}
path.image <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
knitr::include_graphics(path.image)
```

```{r varimpOA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.OA}Variable importance, OA"}
path.image <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.png")
knitr::include_graphics(path.image)
```

```{r varimp, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp}Variable importance", out.height="60%"}
path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")

importance.oa <- read_csv(path.importance.oa) %>% 
  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
  mutate(variable = str_to_sentence(variable))
importance.lad <- read_csv(path.importance.lad) %>% 
  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
  mutate(variable = str_to_sentence(variable))

importance.oa %>% 
  rename(OA = importance) %>% 
  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
  dplyr::select(-variable.y, -id) %>% 
  rename(variable = variable.x) %>% 
  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
  mutate(order = ifelse(variable == "Distance to london", 1,
                        ifelse(variable == "London's wensite density, t-1", 2,
                               ifelse(variable == "Distance to the nearest city", 3,
                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
                                             ifelse(variable == "Distance to the nearest retail centre", 5,
                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
         fill = spatial.unit)) +
  geom_col(position = 'dodge') +
  ylab('') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit: ', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous('Variable importance') +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7))
```

```{r regions}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(caption= "Regional differences\\label{table.regions}", 
        digits=3,
        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
```

Table \ref{table.regions} presents the results of the recursive hold out models
and allows to (i) highlight the regional heterogeneity of the spatial 
processes behind the diffusion of the Web, and (ii) compare this heterogeneity
across the two different scales of analysis. It illustrates some striking 
similarities, but also a few significant differences. At the top of the 
similarity with the rest of the country we see West Midlands and Yorkshire
and the Humber for both scales. The web density for the LAD and OA of these 
regions can be very well predicted by models trained on data for the LAD 
and OA of the rest of the country. In other words, the spatial diffusion 
mechanisms for these regions represent quite well the UK's average. This
should not come as as surprise given their urban structure: they both contain
a large urban centre (Birmingham and Leeds), smaller commuter towns as well 
as more rural settlements. The bottom of the similarity ranking is populated 
by Scotland, Northern Ireland, East of England and London for both scales. 
The diffusion mechanisms for these regions diverge significantly from the 
country's average and this could be attributed to rurality and remoteness 
or to high levels of urbanisation and sophistication of economic activities 
for the case of London and for East of England, which contains the Cambridge 
area.

Still, there are three regions the spatial diffusion mechanisms of which 
differ a lot between the two scales. North West and South West are very
close to the UK's average at the more granular OA scale, while the opposite
applies for South East. Interestingly, the first two regions are well-known 
tourist destinations (Lake District and Cornwall-Devon) and the exploratory
analysis revealed clusters of high web density only visible at the OA scale 
(Figure \ref{lisa}. Then, the South East is very well integrated with the 
Greater London, hosts the broader Oxford area -- a known hub of research 
excellence -- and Figure \ref{lisa} revealed high web density clusters at 
the LAD level.

# Conclusions {#sec-conclusions}

This paper assessed the importance of spatial mechanisms in the diffusion 
of a new technology -- the Web -- at a high level of spatial granularity.
Specifically, it assessed the capacity of the following three spatial 
diffusion mechanisms in predicting website density in the UK over space and 
time: (i) the hierarchical mechanism according to which diffusion runs 
from main urban centres to secondary ones, (ii) the neighborhood mechanism 
which highlights that diffusion first hits nearby locations, and (iii) the 
rather canonical pattern of diffusion over time as reflected in the S-shaped 
pattern of the cumulative level of adoption. This is the first time that 
the importance of granular spatial mechanisms for the diffusion of an
intangible, digital technology is exposed. The time period of the study is 
also important as it captures the very early stages of this technology until 
its maturity. Importantly, the empirical framework and the data employed here 
allows to observe the active engagement with the Web that is creating and 
maintaining a commercial website instead of the more passive action of 
browsing webpages and accessing the Internet.

The analysis firstly illustrated how well the S-pattern of cumulative adoption 
fits at the local scale of LAD. It highlighted the heterogeneity of adoption 
rates between LAD -- with both more and less expected examples -- as well as 
the spatial clustering of LAD with similar adoption rates. The exploratory 
analysis also illustrated instances of remarkable stability as well as 
volatility of the web diffusion system that  is indicative of leapfrogging. 
Previous research had already indicated the long-term productivity gains of 
the early adoption of such technologies [@tranosuk]. Then, the paper developed 
a robust and novel ML modelling framework to predict website diffusion over 
space and time using the three spatial diffusion mechanisms based on space- 
and time-sensitive CV without suffering from overfitting. The model revealed 
the remarkably good predictive capacity of these mechanisms at both scales. 
For the first time, the power of these mechanisms in predicting the diffusion 
of a new technology was empirically tested, importantly, at very fine scales.

The modelling framework also revealed that the importance of these spatial 
mechanisms differ across different scales. While spatial structure as 
reflected in distances to different urban centres is highly important at the 
OA scale, it hardly plays any role at the more coarse LAD scale. And while 
the almost canonical pattern of cumulative adoption is a key predictor at the 
LAD level, it only plays a small role at the detailed OA level. Then, the 
analysis revealed another level of spatial heterogeneity. While for some UK 
regions these spatial diffusion mechanisms are very similar to the rest of 
the country, other regions diverge significantly from the country average.
Importantly, the above results are robust against different specifications 
and subsets of the data.

The paper offers a number of different contributions. Firstly, it assesses, 
for the first time, the role of these spatial diffusion mechanisms at 
such granular spatial scales. While previous studies had 
highlighted these mechanisms [@hagerstrand1968innovation; @rogers2010diffusion; 
@grubler1990rise], their role had never been tested before at such granular 
scale and, importantly, for a digital technology as weightless as the Web. 
The findings talk directly to this literature  [e.g. @beardsell1999spatial; 
@haller2011determinants; @lengyel2020role]. More broadly, the paper offers 
a robust methodological framework, which utilises novel data and modern ML 
algorithms and techniques to model the diffusion of a new technology across 
space and time. It directly addresses the data problem which prevented economic
geographers in engaging  with diffusion studies [@iso2005; 
@kemeny2011international; @zook2022mapping].
Thus, the paper seeks to renew economic geography's interest in the diffusion 
of new technologies. Current data science advances including the availability 
of granular enough data and adequate methods can enable economic 
geographers to reveal interesting spatial patterns of diffusion of new 
technologies as this paper has demonstrated. This is important as (i) 
various stakeholders are interested in predicting localised demand for new 
technologies for planning purposes [@leibowicz2016representing; 
@meade2021modelling], and (ii) the adoption of such digital technologies 
is associated with economic gains [@solow1957technical; @aghion1990model; 
@kemeny2011international; @tranosuk; @capello2024nexus]. In addition, while 
most related studies focused on tangible technologies, this paper sheds 
light on the diffusion of a truly weightless digital technology that is the 
Web. Thus, it adds much more granular findings to the limited related studies
[@sinai2004geography; @haller2011determinants]. Importantly, although the 
geographical focus of this paper is the UK, the empirical framework developed 
here is easily transferable to other geographical contexts.

<!-- Future research can further investigate other -->
<!-- determinants of the diffusion of the Web. For instance, the archived web data -->
<!-- can reveal important information about the industrial base of different places -->
<!-- as well as the industries which were faster in adopting the Web. -->

<!-- reset the figure, table and section numbering -->
\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

# Appendix {#appendix}

<!-- #sec-appendix -->
<!-- .appendix -->

\counterwithin{figure}{section}
\counterwithin{table}{section}

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk10}Grwoth curve, UK; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
knitr::include_graphics(path.image)
```

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map10}LAD adoption rates; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/s/speed_map_10.png")
knitr::include_graphics(path.image)
```

```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank10}Dynamics of wed diffusion (LAD); up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av_10.png")
knitr::include_graphics(path.image)
```

```{r, morani10, echo=FALSE, message=FALSE, fig.cap="\\label{morani10}Website density Moran's I; up to 10 postcodes per website"}

path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_10.csv")
path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa_10.csv")

la.m <- read_csv(path.in.la) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'LAD')
oa.m <- read_csv(path.in.oa) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'OA')

bind_rows(la.m, oa.m) %>% 
  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Moran\'s I') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa10}\\centering Website density LISA maps; up to 10 postcodes per website", out.height="80%"}

path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la1996.png")
path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2000.png")
path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2012.png")
path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa1996.png")
path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2000.png")
path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2012.png")
path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)  
p3 <- ggdraw() + draw_image(path.image3)  
p4 <- ggdraw() + draw_image(path.image4) 
p5 <- ggdraw() + draw_image(path.image5)
p6 <- ggdraw() + draw_image(path.image6)  
p7 <- ggdraw() + draw_image(path.image7)

par(mar=rep(0,4)) # no margins

p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))

# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)

```

```{r, gini10, echo=FALSE, message=FALSE, fig.cap="\\label{gini10}Website density Gini coefficient; up to 10 postcodes per website"}

path.in.la <- paste0(path, "/outputs/gini/gini_la_10.csv")
path.in.oa <- paste0(path, "/outputs/gini/gini_oa_10.csv")

la.gini <- read_csv(path.in.la) %>% 
  mutate(spatial.unit = 'LAD')
oa.gini <- read_csv(path.in.oa) %>% 
  mutate(spatial.unit = 'OA')

bind_rows(la.gini, oa.gini) %>%
  rename(Gini = 'gini(a$n)') %>% 
  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Gini') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r varimp10, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp10}Variable importance; up to 10 postcodes per website", out.height="60%"}
path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected_10.csv")
path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected_10.csv")

importance.oa <- read_csv(path.importance.oa) %>% 
  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
  mutate(variable = str_to_sentence(variable))
importance.lad <- read_csv(path.importance.lad) %>% 
  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
  mutate(variable = str_to_sentence(variable))

importance.oa %>% 
  rename(OA = importance) %>% 
  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
  dplyr::select(-variable.y, -id) %>% 
  rename(variable = variable.x) %>% 
  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
  mutate(order = ifelse(variable == "Distance to london", 1,
                        ifelse(variable == "London's wensite density, t-1", 2,
                               ifelse(variable == "Distance to the nearest city", 3,
                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
                                             ifelse(variable == "Distance to the nearest retail centre", 5,
                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
         fill = spatial.unit)) +
  geom_col(position = 'dodge') +
  ylab('') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit: ', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous('Variable importance') +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7))
```

```{r regions_no_imp}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(caption= "Regional differences; up to 10 postcodes per website\\label{table.regions.no.imp}", 
        digits=3,
        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
```

```{r, echo=FALSE, message = F, error=FALSE}

path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- st_read(path.geo, quiet = T)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- st_transform(la, 4326)

# regions
path.regions <- paste0(path, "/data/raw/LAD21_RGN21_EN_LU.csv")
regions <- read_csv(path.regions)

s.la.path <- paste0(path, "/data/temp/s_la_per_firm.csv")

read.csv(s.la.path) %>% 
  left_join(la %>% as_tibble() %>% select(LAD21CD, LAD21NM),
            by = c("ladcd" = "LAD21CD")) %>% 
  arrange(estimate) %>%
  left_join(regions %>% select(LAD21CD, RGN21NM), by = c("ladcd" = "LAD21CD")) %>% 
  select(LAD21NM, RGN21NM, estimate, std.error, r2, fast) %>% 
  rename(LAD = LAD21NM,
         Region = RGN21NM,
         "t_0 estimate" = "estimate",
         "R-squared" = "r2",
         "Std. error" = "std.error",
         "Diffusion speed" = "fast") %>% 
  kable(caption= "S-curve estiamtes for LAD; up to 10 postcodes per website\\label{table.s.lads}", 
        digits=3,
        align = c("l", "l", "c", "c", "c", "c"),
        col.names = c("LAD", "Region", "$t_0$ estimate", "Std. error", "$R^2$", "Diffusion speed"))
```

# References {#references .unnumbered}