### 2025-01-21 src/ranks_oa_corrected.Rmd

---
title: "Ranks, OA corrected"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
# library(sf)
# library(spdep)
# library(knitr)
# library(REAT)
# library(tidygeocoder)
# library(geosphere)
# library(broom)
# library(doParallel)
# library(raster)
library(plm)
#library(lmtest)
# library(caret)
# library(randomForest)
# library(CAST)
library(ggbump)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

## Problem postcodes

```{r}
n = 1 #number of unique postcodes. 
m = 11

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) %>% 
  filter(
      #V1.domain < m,             # for multiple pc
      V1.domain == n,             # for n == 1
      year > 1995)

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n)             # for n == 1

df.long <- bind_rows(df9610, df1112) %>% 
  filter(year > 1995 & year < 2013) %>% 
  group_by(pc, year) %>% 
  summarise(n = n()) %>%
  ungroup() %>% 
  #arrange(desc(n)) %>% 
  pivot_wider(names_from = year, values_from = n, names_prefix = "n_") %>% 
  dplyr::select(pc, n_1996, n_1999, n_2000, n_2001, n_2002, n_2003, n_2004, n_2005, n_2006, n_2007, n_2008, n_2009, n_2010, n_2011, n_2012) %>%
  arrange(desc(n_2004), desc(n_2005)) %>% 
  slice_head(n=1000) %>% 
  pivot_longer(!pc, names_to = "year", values_to = "n") %>% 
  mutate(year = as.integer(sub("n_", "", year, fixed = TRUE))) %>% 
  replace(is.na(.), 0) %>% 
  mutate(outlier = ifelse(n > 1000 & (year==2004 | year == 2005), pc, ""))

problem.pcs <- df.long %>% filter(year==2004 | year == 2005, 
                   n > 1000) %>% 
  distinct(pc)

# n > 1500: "M28 2SL"  "SE24 9HP" "CV8 2ED"  "CW1 6GL"   
# n > 1000: SE24 9HP, CV8 2ED, GL16 7YA, CW1 6GL, M28 2SL, DE21 7BF
```

## Corrected

The above exploratory analysis indicated that there is a huge increase in 2004-2006
in the above postcodes. I turn them to NAs and then use a regression to impute these
gaps.

```{r}
df.corrected <- bind_rows(df9610, df1112) %>% 
  filter(year > 1995 & year < 2013) %>% 
  group_by(pc, year) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(n = ifelse(pc %in% as.character(problem.pcs$pc) & year > 2001 & year <2007, NA, n))

df.corrected <- pdata.frame(df.corrected)
predictions <- round(predict(plm(n ~ as.factor(year), effect = "individual", 
                                 model = "within", data = df.corrected), newdata = df.corrected), 2)

df.corrected$n[is.na(df.corrected$n)] <- predictions[is.na(df.corrected$n)]
```

## OA df

```{r}
df <- df.corrected %>% 
  mutate(year = as.integer(as.character(year))) %>% 
  filter(year > 1995 & year < 2013) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  summarise(n = sum(n)) %>% # **********ATTENTION*********  
  ungroup() %>% 
  complete(oa11cd, year) %>% 
  filter(!is.na(oa11cd)) %>% # drop c. 10 - 200 per year, which are not assigned to a LAD
  replace(is.na(.), 0) 
```

## Average ranks websites per firm t and 2012 only: 05 and 95

Very awkward figure as the ranks are compresses because of the low ranges. For
instance, there are so many 0s, so all these OAs are together in the bottom.

```{r}

t <- 2000

out.path <- paste0(path, "/outputs/ranks/oa", t,"_2012_only0595_av.png")

df %>%

  # left_join(oa.uk %>% as_tibble() %>% dplyr::select(id),
  #           by = c("ladcd" = "id")) %>%

  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
  filter(year > 1997 & year < 2011) %>% 
  group_by(year) %>% 
  #mutate(rank = rank(desc(n))) %>% # this should be the same as min_rank()
  mutate(rank = min_rank(desc(n))) %>% 
  ungroup() %>% 
  filter(year==1998 | year == 2010) %>% 
  arrange(oa11cd, year) %>% 
  group_by(oa11cd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>% #filter(abs(diff_1st) > 300) #arrange(diff_1st)

  #ggplot(aes(abs(diff_1st))) + geom_histogram()

  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .95, na.rm = T))
  # q.95 = 136884, q.9 = 136884, q.1 = 10469, q.05 = 3915. **abs()** # for dense_rank()
  # q.95 = 302, q.9 = 302, q.1 = 287, q.05 = 277. **abs()** # for min_rank()
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 277 | abs(diff_1st) > 302, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 277, "05",
                    ifelse(abs(diff_1st) > 302, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, rank, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  # scale_y_reverse(labels = c(30000, 60000, 90000, 120000, 161000)) +
  scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

# ggsave(out.path)
```

Plot n

```{r}

t <- 2000

out.path <- paste0(path, "/outputs/ranks/oa_n_", t,"_2012_only0595_av.png")

df %>%

  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>%

  filter(year==1998 | year == 2010) %>% 
  group_by(year) %>% 
  mutate(n.st = n / max(n)) %>%

  group_by(oa11cd) %>% 
  mutate(diff_1st = -(n.st-dplyr::lag(n.st))) %>%

  # ggplot(aes(abs(diff_1st))) + geom_histogram()

  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
  # q.95 = 0.01221996, q.9 = 0.006924644, q.1 = 0, q.05 = 0.
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 0 | abs(diff_1st) > 0.01221996, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 0, "05",
                    ifelse(abs(diff_1st) > 0.01221996, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, n.st, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  #scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

ggsave(out.path)
```

## For specific LA

```{r}

t <- 2000
LAD <- "E09000002"

#out.path <- paste0(path, "/outputs/ranks/oa", t,"_2012_only0595_av.png")

# Lookup: OA to regions
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Lower_layer_Super_Output_Area_to_Middle_layer_Super_Output_Area_to_Local_Authority_District_(December_2011)_Lookup_in_England_and_Wales.csv")  
lookup.lad <- read_csv(path.lookup) %>% dplyr::select(OA11CD, LAD11CD, LAD11NM)

df %>% left_join(lookup.lad, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(LAD11NM, LAD11CD), .after=year) %>% 
  filter(LAD11CD == LAD) %>% 
  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>% 
  filter(year > 1997 & year < 2011) %>% 
  group_by(year) %>% 
  #mutate(rank = rank(desc(n))) %>% # this should be the same as min_rank()
  mutate(rank = min_rank(desc(n))) %>% 
  ungroup() %>% 
  filter(year==1998 | year == 2010) %>% 
  arrange(oa11cd, year) %>% 
  group_by(oa11cd) %>% 
  #mutate(diff_1st = -diff(rank))
  mutate(diff_1st = -(rank-dplyr::lag(rank))) %>% #filter(abs(diff_1st) > 300) #arrange(diff_1st)

  #ggplot(aes(abs(diff_1st))) + geom_histogram()

  #ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
  # q.95 = 9.2, q.05 = 0. **abs()** # for min_rank()
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 0 | abs(diff_1st) > 9.2, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 0, "05",
                    ifelse(abs(diff_1st) > 9.2, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, rank, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  # scale_y_reverse(labels = c(30000, 60000, 90000, 120000, 161000)) +
  scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

# ggsave(out.path)
```

Plot n

```{r}

t <- 2000
LAD <- "E09000023"

out.path <- paste0(path, "/outputs/ranks/oa_n_", t,"_2012_only0595_av.png")

df %>% left_join(lookup.lad, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(LAD11NM, LAD11CD), .after=year) %>% 
  filter(LAD11CD == LAD) %>% 
  mutate(period = ifelse(year < 1999, 1, 
                         ifelse(year > 2009, 3, 2))) %>%
  group_by(oa11cd, period) %>%
  mutate(n = ifelse(period != 2, mean(n), n)) %>%

  filter(year==1998 | year == 2010) %>% 
  group_by(year) %>% 
  mutate(n.st = n / max(n)) %>%

  group_by(oa11cd) %>% 
  mutate(diff_1st = -(n.st-dplyr::lag(n.st))) %>%

  # ggplot(aes(abs(diff_1st))) + geom_histogram()

  # ungroup() %>% summarise(q = quantile(abs(diff_1st), .05, na.rm = T))
  # q.95 = 0.01221996, q.9 = 0.006924644, q.1 = 0, q.05 = 0.
  # ungroup() %>% summarise(max = max(abs(diff_1st), na.rm = T),
  #                         min = min(abs(diff_1st), na.rm = T))

  mutate(alpha = ifelse(abs(diff_1st) <= 0 | abs(diff_1st) > 0.01221996, 1, 0.5)) %>% 
  mutate(alpha = ifelse(is.na(diff_1st), 0.5, alpha)) %>%

  mutate(q = ifelse(abs(diff_1st) <= 0, "05",
                    ifelse(abs(diff_1st) > 0.01221996, "95", "m"))) %>% 
  mutate(q = ifelse(is.na(diff_1st), "m", q)) %>%

  ggplot(aes(year, n.st, group = oa11cd, alpha=alpha)) +
  geom_bump(aes(colour = q)) + #colour = "red"
  scale_color_manual(values = c("darkorchid3", "orange", "grey70"),
                     breaks = c("05", "95"),
                     labels = c("Stability","Volatility")) + 
  labs(color = "oa11cd") +
  #scale_y_reverse() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  #theme(legend.position="none") +
  scale_x_continuous(NULL, 
                     #limits = c(1998, 2010), 
                     breaks=c(1998, 2010),
                     labels = c("1996-98","2010-12")) +
  guides(alpha = "none")

#ggsave(out.path)
```