### 2025-01-21 paper/submission2/tranos2025.qmd

---
title: "A multi-scale story of the diffusion of a new technology: the web"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol
        department: 
        address: 
        city: 
        state: State
        postal-code: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
    Despite the broad consensus that differences in technological adoption and sophistication can lead to uneven trajectories of local development as new technologies can increase productivity and allow firms to introduce new products and services, economic geographers have shied away from exploring the diffusion of digital technologies across space and time. This paper fills this gap by exposing the roles of different spatial mechanisms in the diffusion of an intangible, digital technology at very granular geographical scales. It does so by employing a novel source of data and machine learning algorithms to model the active engagement with Web during 1996-2012.
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment

#insert libraries here
library(rprojroot)
library(tidyverse)
library(knitr)
library(kableExtra)
library(sf)
library(cowplot)
library(patchwork)

# This is the project path
path <- find_rstudio_root_file()
```

# Introduction {#sec-introduction}

<!-- ADD SOMETHING ABOUT MULTIPLE SCALES -->

<!-- geographers lack of interest for diffusion + motivation -->
Economic geographers were always interested in how new technologies and
innovations diffuse across space and time with the work of
@hagerstrand1968innovation, which demonstrated the importance of
communications as a mechanism for the diffusion of innovations, being
the most impactful. It was followed by various empirical studies -- see
for instance @ormrod1990 and @iso2005. However, the torch of exploring
and modelling such diffusion processes had been passed to other
disciplines such as economics, business studies and sociology well
before the 'cultural turn' of economic geography
[@perkins2005international; @ding2010modeling].
<!-- NEW spatial granularity --> As a consequence, relevant studies
did not investigate diffusion mechanisms at granular geographical scales to 
provide local policy insights, for instance, regarding infrastructural d
evelopment and local technology policies. 
A potential explanation of the lack of geographical studies exploring the 
diffusion of new and, more specifically for this paper, digital technologies 
across both space and time can be attributed to the scarcity of relevant and 
granular enough data, a problem also highlighted by @iso2005 and
@kemeny2011international. As @zook2022mapping pointed, digital
activities are hardly ever captured in official data.

<!-- contribution -->
This paper offers such a contribution: an economic geography study
revealing the importance of spatial mechanisms for the diffusion of an
intangible, digital technology that is the Web in the UK at a high level
of spatial granularity 
<!-- NEW TO JUSTIFY DATA TIME PERIOD-->
from the very early days until its maturity (1996-2012).
<!-- NEW DATA CLARIFICATION -->
It specifically focuses on the diffusion of commercial websites.
It does so by employing a novel source of big data and machine learning
algorithms to model the *active* engagement with Web -- that is creating and 
maintaining instead of just browsing a website -- during that period
and tests how different spatial and temporal mechanisms shaped its
diffusion. This paper also exemplifies how the combination of data
sources and state-of-the-art research methods which escape the
traditional social science domain can offer new lenses to geographical
research regarding the understanding of technological diffusion.

<!-- motivation: stakeholders -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffuse over space and time and use this knowledge to make predictions
about the diffusion of related future technologies. As per
@leibowicz2016representing, historical studies agree that technologies
diffuse differently in terms of times, rates, and geographies and can be
driven by related policies [@victor1993]. @meade2021modelling highlight
that a variety of actors have a direct interest in gaining such
knowledge including network equipment suppliers, network operators,
regulatory and local authorities. These processes and their effects vary
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places [@leibowicz2016representing]. Despite the spatial heterogeneity
of such diffusion mechanisms and the policy relevance, there is a
scarcity of studies analysing the diffusion of new digital technologies
at a detailed geographical level because of the lack of detailed enough
data and adequate research methods.

<!-- motivation: regional development -->
This gap becomes even more prominent if we consider the importance of
technology and its adoption in economic geography and macroeconomic
growth theories. Simply put, differences in technological adoption and
sophistication may lead to uneven development as new technologies can
increase productivity and allow firms to introduce new products
[@solow1957technical; @aghion1990model; @kemeny2011international].
Linked to the subject of this paper, the Web and, more specifically, 
commercial websites can be considered as the underpinning
technology for the transformative process identified by
@capello2024nexus as the *digital service economy*. Importantly, the
focus here is not the Web in general, but instead commercial websites.
After all, diffusion together with invention and innovation are
considered the pillars of technological change [@das2022diffusion].
Findings from previous studies illustrated that early engagement with
digital technologies such as the Web led to longstanding positive
productivity gains for these regions that hosted early adopters
[@tranosuk]. Hence, obtaining a robust understanding of the spatial
diffusion mechanisms is crucial in mapping future development trajectories.

<!-- NEW motivation: intangible/digital -->
Another line of motivation has to do with the very nature of the Web as a 
digital technology and how it evolved over time. The digital geographies field 
has debunked premature claims about the Internet's placeless-ness by mapping 
various of its facets and technological layers [@tranos2013death], but the lack 
of relevant long time-series data prevented researchers from observing digital
phenomena overtime. Most of these studies focused on its hard infrastructure 
[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
Exceptions include early work from @zook2000web, @zook2000web and 
@moss1997tracking, which utilised domain name registration data and use it to
map and illustrate the spatial footprint of the web content. Pivotal as these 
contributions have been in understanding the economic geographies of the 
Internet, they were based on cross-sectional, often spatially detailed, mapping 
and on shallow time comparisons (e.g. 1993-1998) and did not observe the 
evolution and diffusion of the Web until its maturity. As we know from the 
relevant literature, different technologies diffuse at different rates and 
there is no theory that can explain these differences 
[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
Web among firms can offer useful insights.

<!-- early engagement with the web -->
This paper focuses on the diffusion of the Web as new technology during
the 1996-2012 period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
Internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
9 per cent of UK's households had access to the Internet in 1998
[@ons2018], the web included mostly static webpages, there were no
social media and web browsing involved exclusively a desktop PCs as
there were no smartphones [@tranosuk]. Hence, it is fair to say that the
study period captures the very early stages of the diffusion of a new
technology that is the Web until its maturity. The former is a key point
in the lifecycle of a new technology. Firstly, during this period new
technologies are expensive, crude and imperfect
[@rosenberg1994exploring; @wilson201281]. A simple comparison between
Web 1.0 and Web 3.0 applications clearly illustrates this
[@tranos2020social]: while a static website facilitates one-way
information dissemination, a platform like `github` enables cooperation
between users and the creation of new information, meaning, and even
knowledge [@faraj2016special; @barassi2012does]. During this period the
performance of a new technology is the main attraction and not the cost
to access and use it [@wilson2011lessons]. There is a broader
theoretical discussion in the literature about the early adoption
motives. As summarised by @perkins2005international, on one hand,
epidemic models highlight the role of interpersonal contacts as a way
for new technologies to diffuse. On the other hand, economic models
underline the importance of heterogeneity. Different firms have
different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories [@neffke2011regions; @simmie2014new]. This
argument is even more relevant when the focus is on digital technologies
because of the commonly found lag between investment and economic
returns as reflected in the Solow paradox [@acemoglu2014return;
@brynjolfsson2018artificial].

[^1]: <https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft>

<!-- data -->
The data used here depict the active engagement with the Web as they
contain geolocated and time-stamped counts of commercial websites from
the UK. Instead of adopting metrics reflecting the passive engagement
with the Internet such as Internet adoption and speeds [e.g.
@blank2018local; @destefano2022fuzzy], this paper observes yearly counts
of geolocated, commercial websites, which reflect the creation and 
maintenance of such websites. It needs to be highlighted here that although 
there are various types of underpinning web technologies -- namely markup,
stylesheet, client- and server-side coding languages as well as
different web browsers and other related programming language -- the
focus here is on the diffusion of the Web as an overarching technology
instead of more specific web technologies. For a such a detailed and 
cross-country technological diffusion study, see @PAPAGIANNIDIS2015308.

<!-- description of the methods and results -->
<!-- AMMEND I AFTER CHANGE THE RESULTS -->
To analyse these data, a diverse set of methods are employed. Firstly,
cumulative adoption curves are estimated for the UK as a whole and for
the Local Authority Districts (LAD), illustrating, for the first time,
such S-shaped diffusion patterns for local areas. This system level
analysis also reveals insights about the volatility of the adoption of
the Web. Then, exploratory spatial data analysis offers some first
indications of the importance of the different spatial diffusion
mechanisms in shaping the diffusion of the Web in the UK. Importantly,
the analysis adopts two distinct geographical scales: the LAD and the
most detailed scale of UK census areas, the Output Areas (OA). Lastly, a
novel methodological framework based on Machine Learning (ML) algorithms
is developed to test the role of such mechanisms at these two scales.
The results reveal, for the first time, the importance of different
spatial diffusion mechanisms in the diffusion of an intangible, digital
technology at very granular geographical scales.

<!-- contents -->
<!-- AMMEND I AFTER CHANGE THE RESULTS -->
The structure of the paper goes as follows. @sec-litreview reviews the
key concepts in the diffusion literature as well as relevant empirical
studies. Then, @sec-datamethods describes the data and methods used in
this paper. The results of the different stages of the analysis are
presented in @sec-results. The discussion of the results and the
conclusions of the paper are offered in @sec-conclusions.

# The mechanisms that shape the diffusion of a new technology {#sec-litreview}

<!-- The fundamentals of technology diffusion -->
The diffusion of new technologies and innovations is driven by three
different mechanisms [@grubler1990rise; @morrill2020spatial]: (i)
distance and proximity, (ii) urban hierarchies in the form of a
centrifugal forces as new technologies diffuse from core to periphery,
and (iii) over time the cumulative level of adoption follows an S-shaped
pattern.
<!-- neighbourhood -->
@hagerstrand1968innovation was the first to identify diffusion as a
geographical process. The starting point was the idea that diffusion is
based on passing information through social networks, which themselves
tend to be geographically bounded. Hence, he identified a
'neighbourhood' mechanism of how information, and consequently,
innovations and new technologies diffuse. This perspective draws
similarities with epidemics: new technologies and innovations, just like
pathogens, spread because of contagion and, consequently, proximity and
exposure [@hivner2003facilitating].

<!-- hierarchy -->
@hagerstrand1968innovation also emphasised the role of the top-down
hierarchy of urban systems and highlighted that new technologies and
innovations are firstly adopted in larger cities and then diffuse to
second tier ones. This is a sequential instead of a simultaneous
process, which resembles the 'lead-lag' spatial acceleration effect in
market research [@bento2018time; @PERES201091]. Schmidt's Law
empirically illustrates a similar pattern [@grubler1990rise]. Core and
highly agglomerated regions is where new technologies are invented and
commercially deployed. This is where the first adopters tend to be
based. Then, technologies spread to the rim and eventually to the
periphery. Although adoption pace might be higher when new technologies
finally arrive to the periphery, the saturation levels there may never
reach the ones in the core because of the lack of infrastructure or
other necessary institutions [@leibowicz2016representing].

<!-- time -->
Diffusion is naturally a dynamic process and Hägerstrand recognised the
role time plays in the diffusion of innovations and new technologies: an
early-pioneering period, a middle fast accelerating period and a final
saturation one [@morrill2020spatial]. This was further explored by
@rogers2010diffusion, who identified five types of adopters of new
technologies and innovations: innovators, early adopters, early
majority, late majority and laggards. Hence, plotting the cumulative
adoption of a new technology over time often leads to an S-shaped line
plot [@grubler1990rise].

Although the above processes have been studied in innovation studies,
the distancing of economic geographers from research questions about the
diffusion of new technologies and innovations resulted to a knowledge gap
regarding the explanatory power of these mechanisms at granular, local
scales [@ding2010modeling]. As discussed in @sec-introduction, such 
spatially explicit knowledge can directly inform local technology
and infrastructure policies. Indeed, while the spatial dimension is
present in the following studies, the geographical level of analysis is
often too coarse to ascertain the role of the above mechanisms at the
local scale. @beardsell1999spatial studied the evolution of the computer
industry in 317 US metropolitan areas during the 1977-1992 period using
employment data. Their analysis indicated that the relative size
distribution holds for urban computer employment and also urban
heterogeneity is essential in explaining this distribution. In a recent
study, @bednarz2020pulled focused on wind turbines and modelled their
spatial diffusion across 402 German regions during 1970-2015. Their key
finding is that local demand than local supply was the main driving
factor. @haller2011determinants employed firm panel data for Ireland
during 2003-2005 and illustrated that when firms are located in a region
or industry with high share of firms having a website, they are more
likely to have a website too. At a global scale @perkins2011internet
explored whether the adoption of previous communication technologies
that is mail, telegrams and telephones were shaped by similar
socioeconomic factors as the Internet. Their results indicated common
patterns regarding the drivers behind the adoption of different
communication technologies. @feldman2015rknowledge focused on inventions
and used patent data to map how the the rDNA technology diffused among
inventors across 366 US metropolitan areas. @ding2010modeling modelled
the spatial diffusion of mobile telecommunications across the highest
level of administrative division in China (29 regions in total). Their
analysis indicated that socioeconomic characteristics are important
determinants of the timing, speed and the level of mobile diffusion
within China. @bakher2013diffusion compared different the growth curves
as a means to model the diffusion of online shopping in Australia during
the 1998-2009 period, but they did not consider any spatial mechanisms.
At a global scale @PAPAGIANNIDIS2015308 modelled the diffusion of
different web technologies and practices across countries. Using similar
but less extensive data as the one used here, their analysis illustrated
how the diffusion of different web technologies and practices follows an
S-shaped pattern as well as the different diffusion rates of the
different technologies and practices. The most detailed consideration of
the spatial footprint of the diffusion of a new technology comes from
@lengyel2020role, who analysed the adoption and churn of a Hungarian
online social network. Their results were in agreement with the early
theoretical and empirical contributions reviewed here: assortativity,
urban scaling and distance are the key drivers of spatial diffusion.
Based on the above, the main empirical hypothesis states:

> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
and the the S-shaped pattern of the cummulative level of adoption shaped
the diffusion of the commercial Web in the UK even at granular geographical 
scales.

<!-- EARLY/LATE adoption-->
Understanding the diffusion of a new technology in multiple, detailed spatial
scales is of interest to economic geography as potential local economic 
benefits may occur because of concentration of early adopters or through 
technological leapfrogging. On the one hand, early adopters are often rewarded 
because of their attitudes towards new technologies and innovations. 
@rogers2010diffusion identifies them as 'knowledgeable risk takers' and 
@griliches1957 as 'profit maximisers' [@ding2010modeling]. On the other hand,
latecomers can adopt and benefit from new technologies that have been developed
elsewhere without incurring the hefty initial R&D costs [@teece2008firm]. At a 
global scale @perkins2005international explored whether the diffusion rate of 
new technologies is driven by a latecomer advantage and the engagement with the 
global economy via foreign direct investments and trade. Their results 
illustrate that indeed latecomers and developing countries experience diffusion 
of new technologies more rapidly than early adopters and developed countries. 
At a regional scale, previous research highlighted the long term and 
sustained productivity benefit of the early adopters of web technologies in 
the UK [@tranosuk]. Although the early adoption and leapfrogging literature has 
not paid much attention on cities and regions [@yu2018sustainability], the 
potential local economic benefits justify the need to investigate the
dynamics of the web diffusion in different scales in the UK. Hence, the second 
empirical hypothesis states:

> H~2~: The diffusion of the commecrial Web is shaped by both early adoption 
and late comer effects.

These are the two hypotheses are empirically tested in @sec-results using the 
data and methods described in the next section.

<!-- Technological diffusion, which is by definition an aggregated process, -->
<!-- can be discussed in parallel with individual adoption mechanisms. On the -->
<!-- one hand,  -->
<!-- On the other hand, @perkins2011internet attribute diffusion -->
<!-- to two processes: (i) epidemic-like mechanisms, which are governed by -->
<!-- distance, proximity and social interactions, and (ii) economic -->
<!-- mechanisms as new innovations are adopted by users as they become more -->
<!-- profitable, valueable and useful. -->

# Data and Methods {#sec-datamethods}

<!-- main data -->
To capture the diffusion of the commercial Web, a website density metric is
developed for two different geographical scales that are discussed later in 
this section. The main data are counts of commercial websites that are 
calculated using data from the Internet Archive[^2] and, specifically, the 
JISC UK Web Domain Dataset [@ukwebarchive]. The Internet Archive is one of the 
most complete and oldest archive of webpages in the world operating since
1996 [@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage it
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK country code top level
domain (ccTLD - .uk) from the 1996–2012 period. In essence, this is a
long list of 2.5 billion URLs of archived webpages including also the
archival timestamp.

[^2]: [https://archive.org/](See%20https://archive.org/).

Instead of using the whole .uk ccTLD, this paper focuses on its
commercial subset, the .co.uk second level domain (SLD). This choice
is aligned with the topic of the paper -- the diffusion of the Web for 
commercial functions -- and decreases the heterogeneity of the web data as such 
commercial websites have specific aims: they are used to diffuse information, 
support online transactions and share opinions [@THELWALL2000441; 
@blazquez2018big]. Although a UK company can adopt a generic TLD such as .com 
and these cases escape the data used here, such omissions should not affect the
validity of the results given the popularity of the .uk ccTLD
[@tranosuk]: UK consumers prefer to visit a .uk website when they are
searching for products or services [@hope]; and anecdotal evidence
indicates that during the first half of 2000, three .co.uk domains were
registered every minute [@oecd_coms]. Importantly, previous studies
illustrated that .co.uk is the most popular UK SLD [@tranosuk]. Henceforward 
any reference to websites will refer to commercial -- .co.uk -- websites.

The text from these webpages was scanned using a regular expression
(regex) to identify strings of text which resemble UK postcodes and one
fifth of them included a mention to a postcode [@BL2013geo]. This
information allows the geolocation of the data and the creation of the
LAD and OA counts. To prevent false positives, the resulted postcodes
were matched against official postcode lookup files [@ons].

The data cleaning process includes an aggregation step, through which
the archived webpages are aggregated to the parent websites. This
website reconstruction process lend itself to the construction of
*website* instead of *webpage* density metrics. Websites tend to
represent specific organisations or entities, and, arguably, are more
meaningful observational units than webpages, which can ignore the
upstream dependency of the website they belong to. Based on the
following example, all three webpages are part of the same website
(<http://www.website.co.uk>). Only the overall website and not the
nested webpages are considered for the counts as otherwise the density
metrics would have been biased towards large, place specific, websites.
It is important to note here that websites commonly contain place references, 
with previous research estimating this share to be around 70% 
[@hill2009georeferencing].

-   <http://www.website.co.uk/webpage_a> B15 2TT

-   <http://www.website.co.uk/webpage_b> BS8 1TH

-   <http://www.website.co.uk/webpage_c> B15 2TT

\noindent What is challenging is that this aggregation approach, which
has been used elsewhere [@tranosuk; @shoreditch] may lead to websites
with multiple postcodes. As per the above example,
[www.website.co.uk](http://www.website.co.uk) includes two unique
postcodes: B15 2TT and BS8 1SS. The distribution of the number of
postcodes per website for 2000 is presented in Table \ref{f2000}, which
illustrates a wide range. At the left end of the distribution, there are
websites anchored to a unique location (72% of all the reconstructed
websites in 2000), which may represent a small company with a single
trading location. At the right end, there are websites with thousands of
different postcodes. Considering the time period of the analysis, such
cases can represent directories which used to be popular in the
pre-search engines early times of the commercial Internet as well as
real estate websites [@tranosuk]. Previous research which used the same 
methodology for a small subset of these data illustrates the robustness of the 
above approach: its results matched in terms of accuracy past rounds of 
research based on extensive qualitative work, interviews and observational 
studies for the same study area [@shoreditch].

The analysis is firstly conducted on websites, which only contain one
unique postcode. As a robustness check, the analysis is replicated for
an extended subset of websites, which include up to 10 unique postcodes
to capture commercial websites which point to multiple locations. They
are geolocated by equally attaching them to these locations. This
extended sample includes 94 percent of all the archived websites in
2000.

```{r eval=T, echo=FALSE, results='asis'}

# load 2000 co.uk
# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
path.2000 <- paste0(path, "/data/temp/all2000couk.csv")

all2000.duplicates <- data.table::fread(path.2000) #
# dim(all2000.duplicates) # 3336162
# it includes duplicated URLs: if one webpages includes multiple postcodes
# then it appears multiple times. This is ok for the nuts aggregation, but
# not for the frequencies
# This is only .co.uk

# one line for every host
all2000 <- unique(all2000.duplicates, by = c("host")) # 57897

# unique postcodes per website f table
f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
                                                              "(100,1e+03]", "(1e+03,1e+04]",
                                                              "(1e+04,1e+05]"))
levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")

f.websites.pc <- f.websites.pc %>% 
  dplyr::select(-cumfreq) %>% 
  rename(Postcodes = level,
         F = freq,
         'F (%)' = perc,
         'Cummulative F' = cumperc)

kable(f.websites.pc,
      format = "latex",
      digits = 3,
      booktabs = T,
      format.args = list(big.mark = ","),
      caption = "Number of unique postcodes per .co.uk website, 2000.\\label{f2000}") %>%
  kableExtra::footnote(general = "Tranos et al. 2021",
                       general_title = "Source: ",
                       footnote_as_chunk = T)  
   # kable_classic(full_width = F, html_font = "Cambria")
```

Another data cleaning step deals with some extreme outliers. Figure
\ref{correct} plots the website counts for the top 1000 postcodes. Some
obvious outliers can be observed for the 2002-2006 period for a handful
of postcodes, which can be attributed to link farms [@BL2013links]. The
website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1000
websites pointing to them, were replaced with predicted values based on
a simple panel regression model with postcode fixed effects and yearly
dummy variables. These postcodes refer to a residential area with a
small park in South London, residential areas outside Warwick and
Manchester, a rural area in the south border of England and Wales and
small business parks in Crew and Derby. Hence, it is difficult to
believe that these postcodes genuinely hosted such a large amounts of
websites, which were later extensively declined. To put the magnitude of
the data imputation into perspective, this process affected 6 out of the
557,808 postcodes included in the data. As the @sec-results illustrates,
this led to a small increase of the model predictive power.

```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
path.image1 <- paste0(path, "/outputs/pc_year_1.png")
path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)

par(mar=rep(0,4)) # no margins

p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
```

<!-- Multi-scale -->
To empirically test the two hypotheses stated in @sec-litreview, τhe website 
density metric was calculated for two different scales the Local Authority
Districts (LAD) and the Output Areas (OA). The former is an administrative unit 
and there are 374 LAD in the UK. The latter is the smallest census-based 
geographical unit and there are c. 230,000 of them in the UK. To create the 
website density metric, the yearly website counts at the LAD level are 
standardised by the number of firms in LAD to avoid biases associated with LAD 
hosting a large number of firms. Given that there is no such readily available 
statistic for the OA, the actual OA level counts are used. However, because of 
the the consistent spatial definition of OA (they host 40-250 households),[^3] 
website counts in OA are interpreted as a density metric too.

[^3]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>

<!-- system-level analysis -->
These website density metrics are used for the three different stages of
the analysis. Firstly, a system-level analysis explores how the
diffusion of web technologies in the UK fits with the well-established
S-curve of cumulative adoption. To do so, the following logistic
function (Equation \ref{eq:s}) is estimated for the whole of the UK and
for each LAD separately:

\begin{align}
y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
\end{align}

\noindent $k$ is the asymptote or, in other words, the saturation level,
$b$ the overall growth rate, and $t_{0}$ the *inflection point* of
maximum growth at $k/2$, where the logistic function is symmetrical
[@wilson201281]. The $t_0$ of each LAD is compared against the $t_0$ of
the UK to delineate whether a LAD reached that point faster or slower
than the country average. Importantly, an accuracy criterion was imposed
and only LAD with $R^2 > 0.9$ were included in this analysis. To
estimate Equation \ref{eq:s} a self-starting logistic growth model was
employed using the `nls` and `SSlogis` functions in `R`. This logistic
growth model depicts the initial period of slow diffusion of a new
technology, which is followed by the fast, exponential growth stage to
then eventually slow down and saturate [@wilson201281;
@grubler1999dynamics].

<!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->

<!-- The literature usually uses the saturation level as the asymptote. I am -->

<!-- using the total number of websites as we cannot compute a rate. -->

The system-level analysis also focuses on the volatility of the web
adoption to depict places with high concentration of early adopters and,
equally, latecomers. To do so, the change of the ranking of the UK LAD
over time is plotted and discussed. Both the S-curve and the volatility
analysis focus only on LAD as the very large number of OA would have
made such analysis difficult to visualise and interpret.

<!-- ESDA -->
Next, exploratory spatial data analysis depicts whether the two main
drivers of spatial diffusion -- namely neighbourhood and hierarchy --
underpin the diffusion of the Web in the UK. To capture the former and
following @ding2010modeling the Moran's I and the Local Index of Spatial
Association (LISA) are estimated for the website density. To address the
hierarchy mechanism, the Gini coefficient -- an established metric of
inequality -- is calculated. All the above are computed and plotted
longitudinally both for the LAD and the OA.

<!-- RF -->
Lastly, a modelling framework is developed to test the above diffusion
mechanisms. The overarching aim is to build a model that can test how
well these mechanisms can predict website density:

\begin{align}
Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S-curve_{t}\label{eq:rf.generic}
\end{align}

\noindent Equation \ref{eq:rf.generic} is estimated using Random Forest
(RF). This is a popular ML algorithm for both regression and
classification problems [@biau2012analysis]. It was introduced by
@breiman2001random and has become a go-to data science tool. RF can
effectively handle skewed distributions and outliers, model non-linear
relationships, require minimal hyperparameter tuning, exhibit low
sensitivity to these parameters, and have relatively short training
times [@Caruana2008; @liaw2002classification; @yan2020using]. These
attributes match well with the website density data characteristics
including skewness especially for the OA. Also, the large data size (c.
230k data points for each of the 17 years) calls for fast training
times. Importantly, RF predictions tend to be more accurate than those
from single regression trees and outperform Ordinary Least Squares in
out-of-sample predictions, even with moderate-sized training data and a
small number of predictors [@mullainathan2017machine; @athey2019machine;
@sulaiman2011intelligent; @pourebrahim2019trip; @biau2012analysis].

RF is a tree-based ensemble learning algorithm [@breiman2001random]. It
begins by generating random samples of the training data, which are then
used to grow regression trees to predict the dependent variable. Data
points and predictors are randomly sampled for the different trees. The
trees are trained in parallel using their own bootstrapped samples of
the training data. A crucial feature of RF is their ability not to
overfit, meaning they can generalize well to unseen test data. While
each tree may overfit individually, the ensemble of trees does not
because the errors of individual trees are averaged, reducing the
overall variance and preventing overfitting [@last2002improving]. For
regression problems, RF predictions are made by averaging the
predictions of all decision trees.

RF have been widely employed to address regression research problems.
@pourebrahim2019trip combined a spatial interaction modelling framework
with ML algorithms including RF to predict commuting flows in New York
City. @sinha2019assessing advocated for adopting spatial ensemble
learning approaches, such as RF, to model spatial data with high
autocorrelation and heterogeneity. @creditspatial predicted employment
density in Los Angeles using spatially explicit RF.
@guns2014recommending used RF to build a recommendation system for
research collaborations. @ren2019predicting trained RF to predict the
socio-economic status of cities using various online and mobility
predictors. @tranos2023using utilised hyperlinks data and RF to make
out-of-sample predictions of interregional trade. @zhou2023geography
employed such a framework to assess whether key predictors of obesity
differ across English cities.

# Results {#sec-results}

## System-level analysis

<!-- S-shaped -->

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Grwoth curve, UK"}
path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
knitr::include_graphics(path.image)
```

Figure \ref{s_uk} plots the cumulative adoption of the Web in total in
the UK during the 1996-2012 period. Not surprisingly, the cumulative
adoption is well represented by a curve that resembles the figure S. The
vertical line around year 2003 illustrates the point where the modelled
cumulative adoption is equal to 50% of the maximum. This $t_0$
*inflection* point signals the maximum adoption speed and is used here
to determine whether a UK LAD reached that inflection point earlier or
later than the UK average. Specifically, the S curve and the inflection
point are estimated for every LAD individually and then compared against
the UK average. LAD that reached their inflection point earlier than the
country average are labelled as *fast* and the rest as *slow*. Figure
\ref{s_map} maps this pattern. There is a clear high concentration of
fast adopting LAD in the South and East of England as expected. However,
there is also such a concentration in Wales, Scotland and the North West
of England. At a more detailed level, there are some expected examples
of LAD with relevant industrial backgrounds delineated as fast: the City
of London, a world-renowned cluster of finance industries
[@cook2007role], and Reading, a town with high-tech service industries
in proximity to London and its main airport, Heathrow
[@england2005polynet]. However, there are also LAD which were expected
to appear as fast -- e.g. Hackney in central London and Bristol, a
well-established creative cluster [@oatley1999cultural;
@bassett2002cultural] -- but were delineated as slow. Nevertheless,
there is obvious spatial clustering of fast and slow LAD. Importantly,
out of the ten fastest LAD, nine were located in the South East of
England and London and one in Cambridge (see Table \ref{table.s.lads} in
the Appendix). The above observations remain almost unchanged when the
analysis includes websites with up to 10 unique postcodes -- see Figures
\ref{s_uk10} and \ref{s_map10} in the Appendix -- advocating towards the
robustness of the analysis.

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD adoption rates"}
path.image <- paste0(path, "/outputs/s/speed_map.png")
knitr::include_graphics(path.image)
```

<!-- discussion -->
The S-curves of the cumulative adoption of Web in the UK in total and in
the UK LAD demonstrate a pattern well aligned with all previous studies
discussed in @sec-litreview. Importantly, the above findings are in line
with previous studies exploring the diffusion of web technologies
[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
shopping [@bakher2013diffusion]. Similar to @beardsell1999spatial, who
analysed the spatial evolution of computer industry employment across
317 US metro areas, spatial heterogeneity is evident in the diffusion
speed when the scale of analysis is as detailed as the LAD. Importantly,
no previous studies estimated such cumulative adoption curves for areal
units as small as the UK LADS.

<!-- rankings -->
The next step assesses the stability and volatility of the LAD in terms
of their adoption of web technologies. As highlighted in the literature
[@risk_perceptions], different agents have different perceptions about
and levels of acceptance of the risks and the potential economic returns
associated with the adoption of new technologies -- see for instance the
seminal work of @venkatesh2000theoretical. To reveal such aggregated
patterns, Figure \ref{rank} plots the ranking of UK LAD based on website
density. To decrease noise, the average ranking of 1996-1998 and
2010-2012 is plotted instead of the individual years. The colours depict
the 5th and 95th percentile of the absolute difference of the LAD
ranking between 1996-1998 and 2010-2012. While there are quite a few
obvious cases of LAD that maintained their position between the
beginning and the end of the study period both at the top or at the
bottom of the hierarchy, there are also quite a few LAD that changed
drastically their position. Some of these LAD enjoyed a process that at
the first instance looks like *leapfrogging* since they managed to jump
at the top of the hierarchy despite their slow start. There is extensive
literature regarding the potential benefits of technological
leapfrogging. The underpinning argument is that latecomers can adopt and
benefit from new technologies that have been developed elsewhere without
incurring the hefty initial R&D costs [@teece2008firm]. Although the
leapfrogging literature does not pay much attention to cities and
regions [@yu2018sustainability], previous research highlighted the long
term and sustained productivity benefit of the early adopters of web
technologies in the UK [@tranosuk]. So, although the LAD system is
volatile, it is not clear whether the LAD with high concentration of
late adopters will gain any latecomer-related benefits in a way similar
to countries experiencing technological leapfrogging. Still, the
opportunity to catch up that @perkins2005international illustrated at
the country level, is also visible at the much more detailed LAD scale.
As previously, this is consistent with including websites with up to 10
postcodes (see Figure \ref{rank10}).

```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of wed diffusion (LAD)"}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
knitr::include_graphics(path.image)
```

## Exploratory analysis

<!-- Neighbourhood effect -->
Figures \ref{morani} and \ref{lisa} offer a first insight into whether a
neighbourhood mechanism underpins the diffusion of the Web in the UK as
they present the Moran's I and the LISA maps of website density
respectively for LAD and OA. Starting from the former, spatial
autocorrelation was higher in the beginning of the study period and then
after 2000 dropped slightly and stabilised around 0.2. This reflects the
early concentration of high website density around London, which over
time diffused as high website density clusters can be seen in other
parts of the country away from London (Figure \ref{lisa}). An almost
reverse pattern can be observed for OA. At the beginning of the study
period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
Because of the very small size of OA, at the early stages of the
diffusion of web technologies their adoption was spatially scattered.
This is reflected in the lack of any significant clusters in 1996
(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
such clusters of high website density were formed and this is reflected
both in the Moran's I and the LISA maps in Figures \ref{morani} and
\ref{lisa}. A similar pattern is observed for the extended dataset (see
Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
spatial autocorrelation is slightly higher illustrating a stronger
spatial clustering mechanism. All in all, the exploratory spatial data
analysis advocates towards an underpinning neighbourhood mechanism and
the different scales of analysis illustrate how it evolved differently
over time. <!-- discussion --> Interestingly, the magnitude of Moran's I
is similar to the ones revealed by @ding2010modeling about mobile phone
adoption in China albeit the much more detailed spatial scale of this
paper.

```{r, morani, echo=FALSE, message=FALSE, fig.cap="\\label{morani}Website density Moran's I"}

path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa.csv")

la.m <- read_csv(path.in.la) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'LAD')
oa.m <- read_csv(path.in.oa) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'OA')

bind_rows(la.m, oa.m) %>% 
  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Moran\'s I') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa}\\centering Website density LISA maps", out.height="80%"}

path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la1996.png")
path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2000.png")
path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2012.png")
path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa1996.png")
path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2000.png")
path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2012.png")
path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)  
p3 <- ggdraw() + draw_image(path.image3)  
p4 <- ggdraw() + draw_image(path.image4) 
p5 <- ggdraw() + draw_image(path.image5)
p6 <- ggdraw() + draw_image(path.image6)  
p7 <- ggdraw() + draw_image(path.image7)

par(mar=rep(0,4)) # no margins

library(patchwork)
# (p1 | p4)/(p2 | p5)/(p3 | p6)/ p7 +
#   #plot_layout(widths = 1)
#     plot_layout(widths = c(1, -0.5))

# p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 / p7 +
#   plot_layout(ncol = 3, widths = c(1, -0.85, 1), heights = c(1,1,1,2))

p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))

# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
```

To illustrate whether a hierarchical process also underpins the
diffusion of web technologies, the Gini coefficient is calculated yearly
both for LAD and OA. As a metric of inequality, the Gini coefficient
demonstrates whether website density is concentrated in a small number
of LAD or OA, or whether it is more equally spread across the country.
Both scales of analysis in Figure \ref{gini} illustrate the same
picture. At the beginning of the commercial Internet, website density
was extremely unequal, or, in other words, only a few places had
websites associated with them. Inequality dropped and plateaued after
2000 for both scales. This is illustrative of a hierarchical diffusion
mechanism that led over time to a more equal spread. Interestingly, the
year 2000 is again a period of change for this diffusion mechanism as it
was for the neighbourhood process. There is a substantial difference
between the Gini coefficient magnitude for LAD and OA, but this is
expected as the very small size of OA equates to a lot of polygons
without any websites pointing to them -- for example residential OA.

```{r, gini, echo=FALSE, message=FALSE, fig.cap="\\label{gini}Website density Gini coefficient"}

path.in.la <- paste0(path, "/outputs/gini/gini_la.csv")
path.in.oa <- paste0(path, "/outputs/gini/gini_oa.csv")

la.gini <- read_csv(path.in.la) %>% 
  mutate(spatial.unit = 'LAD')
oa.gini <- read_csv(path.in.oa) %>% 
  mutate(spatial.unit = 'OA')

bind_rows(la.gini, oa.gini) %>%
  rename(Gini = 'gini(a$n)') %>% 
  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Gini') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

## Modelling outputs

The next section incorporates the above mechanisms of the spatial
diffusion of the Web into a modelling framework. The aim is to use
variables depicting these spatial processes in order to predict the
diffusion of the Web as a new technology in the UK over space and time
and across different scales. Specifically, four different models are
estimated. Firstly, all the data points for the OA and LAD are utilised
in order to build two RF models and assess their capacity to predict the
adoption of the Web at the two different scales. These two models will
reveal the predictive capacity of the diffusion mechanisms and also
demonstrate how the importance of such variables changes across
different scales. The next two sets of models are again trained at these
two scales: OA and LAD. However, instead of using all the data points,
the OA and the LAD from one of the twelve UK regions are held out for
the model training. Then, the trained model is used to predict website
density for the OA or the LAD of the held-out region. This process takes
place recursively for all twelve UK regions. The regional differences in
the predictive capacity of the different samples will reveal how
dissimilar these spatial process are across regions and, importantly, at
different scales.

It needs to be highlighted here that the cross validation (CV) for all
models is spatially and temporally sensitive. Instead of using 10 random
and space- and time-agnostic folds,the `CAST` package is employed which
allows holding back data points from specific years and spatial units
and use them for testing in order to estimate the model performance
[@meyer2018improving].

The models include variables that capture the three processes that the
relevant literature and the descriptive analysis highlighted. Namely,
the models capture: (i) a hierarchical mechanism with diffusion running
from main centres to secondary ones, (ii) a neighborhood mechanism
according to which diffusion first hits nearby locations, and (iii) the
rather canonical pattern of diffusion over time as reflected in the
S-shaped pattern in the cumulative level of adoption.

To capture the hierarchical mechanism the models include as predictors a
one year lag of website density in London, the largest city in the UK, a
one year lag of the website density in the nearest city and the same for
the nearest retail centre. Due to the small sizes of the retail centres,
the latter is only relevant for and included in the OA-level models. In
addition, the models include the distance to London, the nearest city
and the nearest retail centre. The underlying logic is that the level of
website adoption in a spatial unit depends on the level of the adoption
in places further up in the urban hierarchy the previous year. The
inclusion of the distance variables incorporates spatial structure into
the hierarchy argument. To depict the neighbourhood mechanism, the
website density of the neighbouring spatial units in the previous year
is employed. This is calculated using rook continuity. Again, the
underpinning rationale is that the level of web adoption within a
spatial unit depends on the level of web adoption in the neighbouring
spatial units the year before. This represents the 'hitting nearby
locations first' argument. Therefore, the spatial and temporal lag of
the website density in LAD and OA is calculated. Lastly, the temporal
pattern of the cumulative adoption is captured by a time trend variable.
Hence, all four models will follow the following generic form (Eq.
\ref{model}):

\begin{align} \label{model}
Website\,Density_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}

To assess the predictive power of the model, three broadly utilised
metrics are employed: the coefficient of determination ($R^2$), mean
absolute error (MAE) and root mean square error (RMSE):

\begin{align}
R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
\end{align}
\begin{align}
MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
\end{align}
\begin{align}
RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
\end{align}

\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which
consists of $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$
predicted value for the dependent variable and $\overline{y_{k}}$ is the
average value of $y$. The last two metrics are expressed in the same
units as the dependent variable -- websites per firm for the LAD modes
and the number of websites for the OA models -- while the first one is
the coefficient of determination between the observed and the predicted
values of website adoption. Regarding $MAE$, it is the absolute
difference between the observed and the predicted website adoption.
While $MAE$ does not penalise for large errors, $RSME$ does, as it is
proportional to the squared difference between the observed and the
predicted trade flows. Hence larger errors weigh more for $RMSE$
[@pontius2008components].

@tbl-model-metrics presents the model performance for the first set of
models, for which all data points are employed for training and testing
via CV. The first one is trained and tested on 374 LAD and the second on
232,296 OA, both for the 16 year period (1997-2012). The results are
remarkably good considering that the are the outcome of space and time
sensitive CV, so the the model does not suffer from overfitting. At the
LAD level the model predicts 85% of the variation of website density.
Both error metrics indicate that the model error is
$(\frac{1}{20}, \frac{1}{30})$ of a website per firm. At the OA the
$R^2$ drops down to 32%. Considering its granularity, this is still a
remarkable performance. To contextualise it, the model results in a
$MAE$ equal to one website for areas small enough to host less than 140
households. [^4] Because of the small size of the spatial units, the
distribution is highly skewed and a significant part of them is not
linked to any websites. In 1997 only 1% of the UK OA were associated
with at least one website. This should not come as a surprise as this
was the very beginning of the commercial Internet and any activities
with a digital footprint were concentrated in a handful of areas. This
was clearly illustrated in Figure \ref{lisa}. At the end of the study
period almost half of the UK OA were not associated with a website.
Again, given the granularity of the data this should not come as a
surprise.

[^4]: According to the Office for National Statistics, 80% of OA in
    England and Wales host 110-139 households,
    [www.ons.gov.uk](https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas).

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.028  | 0.850   | 0.019 |
| Output Areas      | 3.284  | 0.320   | 1.034 |

: Model metrics {#tbl-model-metrics}

The results in @tbl-model-metrics are robust against different
specification. As noted earlier, the data imputation process resulted in
a modest improvement of the predictive capacity of the model. As per
@tbl-model-metrics-no-correction all accuracy metrics are only slightly
worst when the data imputation is not applied. As expected, the
improvement is higher for the more detailed scale of analysis. Then,
@tbl-model-metrics-10 presents the model results for the extended
dataset, which includes websites with up to 10 unique postcodes. While
the results for the LAD are slightly less accurate, the $R^{2}$ of the
OA model is more than double in comparison to the base results in
@tbl-model-metrics. All in all, the modelling exercise revealed how well
can the three diffusion mechanisms predict the diffusion of the Web as a
new technology. Importantly, this predictive power is highly robust
against against different scales and delineations of the diffusion of
the web.

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.032  | 0.810   | 0.019 |
| Output Areas      | 5.000  | 0.205   | 1.047 |

: Model metrics without data imputation
{#tbl-model-metrics-no-correction}

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.188  | 0.77    | 0.122 |
| Output Areas      | 22.467 | 0.478   | 5.681 |

: Model metrics for the extended dataset {#tbl-model-metrics-10}

Figure \ref{var.imp} plots the importance of the different predictors.
When the focus is on LAD, the website density in the nearest city, in
London and in the neighbouring LAD the year before are the most
important predictors. They are followed by the yearly trend, while the
spatial configuration as reflected in distances to London or the nearest
city only plays a minor role. This can be attributed to the relatively
coarse spatial scale of LAD. Nevertheless, all previously discussed
spatial processes are at play in the diffusion of web technologies at
the LAD level: the first two predictors depict the hierarchical
mechanism, the spatial and temporal lag of website density the
neighbourhood mechanism, and the yearly trend the time-sensitive
cumulative adoption pattern.

When the much more granular scale of OA is adopted, the picture is
almost reversed. The two most important predictors are the distance to
London and to the nearest city followed by the spatial and temporal lag
of website density and the distance to the nearest retail centre. They
still depict a hierarchical mechanism, but proximity to the different
population centres is more important than their lagged web densities in
predicting website diffusion. The neighbourhood mechanism is still
strong but less important at this scale. What is interesting is the
almost negligible role of the yearly trend and London's website density.
While the former probably illustrates the large heterogeneity in how the
Web has been adopted at this very fine scale, the latter highlights that
the importance of past web adoption rates in large population centres is
surpassed by proximity to them and spatial configuration at this scale.

The above pattern of variable importance is largely in accordance with
the models estimated for the extended dataset. A comparison with Figure
\ref{var.imp10} illustrates that the main difference is the importance
of the neighbourhood mechanism for the OA as the space and time lag of
the website density is by far the most important predictor when the
extended dataset is employed. But still, the variable importance
patterns for the two datasets are not dissimilar.

```{r varimpLA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.LAD}Variable importance, LAD"}
path.image <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.png")
knitr::include_graphics(path.image)
```

```{r varimpOA, eval = F, echo=FALSE, message = F, fig.cap="\\label{var.imp.OA}Variable importance, OA"}
path.image <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.png")
knitr::include_graphics(path.image)
```

```{r varimp, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp}Variable importance", out.height="60%"}
path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")

importance.oa <- read_csv(path.importance.oa) %>% 
  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
  mutate(variable = str_to_sentence(variable))
importance.lad <- read_csv(path.importance.lad) %>% 
  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
  mutate(variable = str_to_sentence(variable))

importance.oa %>% 
  rename(OA = importance) %>% 
  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
  dplyr::select(-variable.y, -id) %>% 
  rename(variable = variable.x) %>% 
  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
  mutate(order = ifelse(variable == "Distance to london", 1,
                        ifelse(variable == "London's wensite density, t-1", 2,
                               ifelse(variable == "Distance to the nearest city", 3,
                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
                                             ifelse(variable == "Distance to the nearest retail centre", 5,
                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
         fill = spatial.unit)) +
  geom_col(position = 'dodge') +
  ylab('') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit: ', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous('Variable importance') +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7))
```

```{r regions}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(caption= "Regional differences\\label{table.regions}", 
        digits=3,
        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
```

Table \ref{table.regions} presents the results of the recursive hold out
models and allows to (i) highlight the regional heterogeneity of the
spatial processes behind the diffusion of the Web, and (ii) compare this
heterogeneity across the two different scales of analysis. It
illustrates some striking similarities, but also a few significant
differences. At the top of the similarity with the rest of the country
we see West Midlands and Yorkshire and the Humber for both scales. The
web density for the LAD and OA of these regions can be very well
predicted by models trained on data for the LAD and OA of the rest of
the country. In other words, the spatial diffusion mechanisms for these
regions represent quite well the UK's average. This should not come as
as surprise given their urban structure: they both contain a large urban
centre (Birmingham and Leeds), smaller commuter towns as well as more
rural settlements. The bottom of the similarity ranking is populated by
Scotland, Northern Ireland, East of England and London for both scales.
The diffusion mechanisms for these regions diverge significantly from
the country's average and this could be attributed to rurality and
remoteness or to high levels of urbanisation and sophistication of
economic activities for the case of London and for East of England,
which contains the Cambridge area.

Still, there are three regions the spatial diffusion mechanisms of which
differ a lot between the two scales. North West and South West are very
close to the UK's average at the more granular OA scale, while the
opposite applies for South East. Interestingly, the first two regions
are well-known tourist destinations (Lake District and Cornwall-Devon)
and the exploratory analysis revealed clusters of high web density only
visible at the OA scale (Figure \ref{lisa}. Then, the South East is very
well integrated with the Greater London, hosts the broader Oxford area
-- a known hub of research excellence -- and Figure \ref{lisa} revealed
high web density clusters at the LAD level.

# Conclusions {#sec-conclusions}

<!-- ADD MORE ABOUT LOCAL POLICIES: E.G. USE MY RESULTS TO MODEL WHERE INFRASTRUCTURE OR GSM ANTENNAS SHOULD BE LOCALTED -->

This paper assessed the importance of spatial mechanisms in the
diffusion of a new technology -- the Web -- at a high level of spatial
granularity. Specifically, it assessed the capacity of the following
three spatial diffusion mechanisms in predicting website density in the
UK over space and time: (i) the hierarchical mechanism according to
which diffusion runs from main urban centres to secondary ones, (ii) the
neighborhood mechanism which highlights that diffusion first hits nearby
locations, and (iii) the rather canonical pattern of diffusion over time
as reflected in the S-shaped pattern of the cumulative level of
adoption. This is the first time that the importance of granular spatial
mechanisms for the diffusion of an intangible, digital technology is
exposed. The time period of the study is also important as it captures
the very early stages of this technology until its maturity.
Importantly, the empirical framework and the data employed here allows
to observe the active engagement with the Web that is creating and
maintaining a commercial website instead of the more passive action of
browsing webpages and accessing the Internet.

The analysis firstly illustrated how well the S-pattern of cumulative
adoption fits at the local scale of LAD. It highlighted the
heterogeneity of adoption rates between LAD -- with both more and less
expected examples -- as well as the spatial clustering of LAD with
similar adoption rates. The exploratory analysis also illustrated
instances of remarkable stability as well as volatility of the web
diffusion system that is indicative of leapfrogging. Previous research
had already indicated the long-term productivity gains of the early
adoption of such technologies [@tranosuk]. Then, the paper developed a
robust and novel ML modelling framework to predict website diffusion
over space and time using the three spatial diffusion mechanisms based
on space- and time-sensitive CV without suffering from overfitting. The
model revealed the remarkably good predictive capacity of these
mechanisms at both scales. For the first time, the power of these
mechanisms in predicting the diffusion of a new technology was
empirically tested, importantly, at very fine scales.

The modelling framework also revealed that the importance of these
spatial mechanisms differ across different scales. While spatial
structure as reflected in distances to different urban centres is highly
important at the OA scale, it hardly plays any role at the more coarse
LAD scale. And while the almost canonical pattern of cumulative adoption
is a key predictor at the LAD level, it only plays a small role at the
detailed OA level. Then, the analysis revealed another level of spatial
heterogeneity. While for some UK regions these spatial diffusion
mechanisms are very similar to the rest of the country, other regions
diverge significantly from the country average. Importantly, the above
results are robust against different specifications and subsets of the
data.

The paper offers a number of different contributions. Firstly, it
assesses, for the first time, the role of these spatial diffusion
mechanisms at such granular spatial scales. While previous studies had
highlighted these mechanisms [@hagerstrand1968innovation;
@rogers2010diffusion; @grubler1990rise], their role had never been
tested before at such granular scale and, importantly, for a digital
technology as weightless as the Web. The findings talk directly to this
literature [e.g. @beardsell1999spatial; @haller2011determinants;
@lengyel2020role]. More broadly, the paper offers a robust
methodological framework, which utilises novel data and modern ML
algorithms and techniques to model the diffusion of a new technology
across space and time. It directly addresses the data problem which
prevented economic geographers in engaging with diffusion studies
[@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
seeks to renew economic geography's interest in the diffusion of new
technologies. Current data science advances including the availability
of granular enough data and adequate methods can enable economic
geographers to reveal interesting spatial patterns of diffusion of new
technologies as this paper has demonstrated. This is important as (i)
various stakeholders are interested in predicting localised demand for
new technologies for planning purposes [@leibowicz2016representing;
@meade2021modelling], and (ii) the adoption of such digital technologies
is associated with economic gains [@solow1957technical;
@aghion1990model; @kemeny2011international; @tranosuk;
@capello2024nexus]. In addition, while most related studies focused on
tangible technologies, this paper sheds light on the diffusion of a
truly weightless digital technology that is the Web. Thus, it adds much
more granular findings to the limited related studies
[@sinai2004geography; @haller2011determinants]. Importantly, although
the geographical focus of this paper is the UK, the empirical framework
developed here is easily transferable to other geographical contexts.

<!-- Future research can further investigate other -->

<!-- determinants of the diffusion of the Web. For instance, the archived web data -->

<!-- can reveal important information about the industrial base of different places -->

<!-- as well as the industries which were faster in adopting the Web. -->

<!-- reset the figure, table and section numbering -->

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

# Appendix {#appendix}

<!-- #sec-appendix -->

<!-- .appendix -->

\counterwithin{figure}{section}
\counterwithin{table}{section}

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk10}Grwoth curve, UK; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
knitr::include_graphics(path.image)
```

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map10}LAD adoption rates; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/s/speed_map_10.png")
knitr::include_graphics(path.image)
```

```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank10}Dynamics of wed diffusion (LAD); up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av_10.png")
knitr::include_graphics(path.image)
```

```{r, morani10, echo=FALSE, message=FALSE, fig.cap="\\label{morani10}Website density Moran's I; up to 10 postcodes per website"}

path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_10.csv")
path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa_10.csv")

la.m <- read_csv(path.in.la) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'LAD')
oa.m <- read_csv(path.in.oa) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'OA')

bind_rows(la.m, oa.m) %>% 
  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Moran\'s I') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa10}\\centering Website density LISA maps; up to 10 postcodes per website", out.height="80%"}

path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la1996.png")
path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2000.png")
path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2012.png")
path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa1996.png")
path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2000.png")
path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2012.png")
path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)  
p3 <- ggdraw() + draw_image(path.image3)  
p4 <- ggdraw() + draw_image(path.image4) 
p5 <- ggdraw() + draw_image(path.image5)
p6 <- ggdraw() + draw_image(path.image6)  
p7 <- ggdraw() + draw_image(path.image7)

par(mar=rep(0,4)) # no margins

p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))

# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)

```

```{r, gini10, echo=FALSE, message=FALSE, fig.cap="\\label{gini10}Website density Gini coefficient; up to 10 postcodes per website"}

path.in.la <- paste0(path, "/outputs/gini/gini_la_10.csv")
path.in.oa <- paste0(path, "/outputs/gini/gini_oa_10.csv")

la.gini <- read_csv(path.in.la) %>% 
  mutate(spatial.unit = 'LAD')
oa.gini <- read_csv(path.in.oa) %>% 
  mutate(spatial.unit = 'OA')

bind_rows(la.gini, oa.gini) %>%
  rename(Gini = 'gini(a$n)') %>% 
  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Gini') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r varimp10, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp10}Variable importance; up to 10 postcodes per website", out.height="60%"}
path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected_10.csv")
path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected_10.csv")

importance.oa <- read_csv(path.importance.oa) %>% 
  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
  mutate(variable = str_to_sentence(variable))
importance.lad <- read_csv(path.importance.lad) %>% 
  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
  mutate(variable = str_to_sentence(variable))

importance.oa %>% 
  rename(OA = importance) %>% 
  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
  dplyr::select(-variable.y, -id) %>% 
  rename(variable = variable.x) %>% 
  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
  mutate(order = ifelse(variable == "Distance to london", 1,
                        ifelse(variable == "London's wensite density, t-1", 2,
                               ifelse(variable == "Distance to the nearest city", 3,
                                      ifelse(variable == "Nearest city's wensite density, t-1", 4,
                                             ifelse(variable == "Distance to the nearest retail centre", 5,
                                                    ifelse(variable == "Nearest retail centre's wensite density, t-1", 6,
                                                           ifelse(variable == "Spatial and temporal lag of wensite density", 7, 8)))))))) %>% 
  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
         fill = spatial.unit)) +
  geom_col(position = 'dodge') +
  ylab('') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit: ', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous('Variable importance') +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7))
```

```{r regions_no_imp}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(caption= "Regional differences; up to 10 postcodes per website\\label{table.regions.no.imp}", 
        digits=3,
        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
```

```{r, echo=FALSE, message = F, error=FALSE}

path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- st_read(path.geo, quiet = T)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- st_transform(la, 4326)

# regions
path.regions <- paste0(path, "/data/raw/LAD21_RGN21_EN_LU.csv")
regions <- read_csv(path.regions)

s.la.path <- paste0(path, "/data/temp/s_la_per_firm.csv")

read.csv(s.la.path) %>% 
  left_join(la %>% as_tibble() %>% select(LAD21CD, LAD21NM),
            by = c("ladcd" = "LAD21CD")) %>% 
  arrange(estimate) %>%
  left_join(regions %>% select(LAD21CD, RGN21NM), by = c("ladcd" = "LAD21CD")) %>% 
  select(LAD21NM, RGN21NM, estimate, std.error, r2, fast) %>% 
  rename(LAD = LAD21NM,
         Region = RGN21NM,
         "t_0 estimate" = "estimate",
         "R-squared" = "r2",
         "Std. error" = "std.error",
         "Diffusion speed" = "fast") %>% 
  kable(caption= "S-curve estiamtes for LAD; up to 10 postcodes per website\\label{table.s.lads}", 
        digits=3,
        align = c("l", "l", "c", "c", "c", "c"),
        col.names = c("LAD", "Region", "$t_0$ estimate", "Std. error", "$R^2$", "Diffusion speed"))
```

# References {#references .unnumbered}