### 2025-01-21 src/help_memory_log.Rmd

```{r}
#mem <- 
read_delim("memory.log", delim = "\\t") %>%
  rename(t = 1) %>% 
  filter(!str_detect(t, 'total')) %>% 
  separate(t, sep = ":", into = c("type", "b")) %>% 
  filter(type == "Mem") %>% 
  separate(b, sep = "\\S*\\K\\s+", into = c("total", "used", "free", "shared", "buff/cache", "available")) %>% 
  arrange(shared)
  #extract(b, c("total", "used", "free", "shared", "buff/cache", "available"), "(\\d+)\\s(\\d+)\\s(\\d+)\\s(\\d+)\\s(\\d+)\\s(\\d+)\\s")

read_delim("memory.log", delim = "\\t") %>%
  rename(t = 1) %>% 
  filter(!str_detect(t, 'total')) %>% 
  separate(t, sep = ":", into = c("type", "b")) %>% 
  filter(type != "Mem") %>% 
  separate(b, sep = "\\S*\\K\\s+", into = c("total", "used", "free")) %>% 
  arrange(desc(free))
```