### 2025-03-12 paper/submission2/anonymised.qmd

---
title: "A multi-scale story of the diffusion of a new technology: the web"
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE) # By default, hide code; set to TRUE to see code
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment

#insert libraries here
library(rprojroot)
library(tidyverse)
library(knitr)
library(kableExtra)
library(sf)
library(cowplot)
library(patchwork)

# This is the project path
path <- find_rstudio_root_file()
```

# Introduction {#sec-introduction}

<!-- geographers lack of interest for diffusion + gap -->
Economic geographers were always interested in how new technologies and
innovations diffuse across space and time with the work of
@hagerstrand1968innovation, which demonstrated the importance of
communications as a mechanism for the diffusion of innovations, being
the most impactful. It was followed by various empirical studies -- see
for instance @ormrod1990 and @iso2005. However, the torch of exploring
and modelling such diffusion processes had been passed to other
disciplines such as economics, business studies and sociology well
before the 'cultural turn' of economic geography
[@perkins2005international; @ding2010modeling].
<!-- NEW spatial granularity --> As a consequence, relevant studies
did not investigate diffusion mechanisms at granular geographical scales to 
provide local policy insights, for instance, regarding infrastructural 
development and local technology policies. 
A potential explanation of the lack of geographical studies exploring the 
diffusion of new and, more specifically for this paper, digital technologies 
across both space and time can be attributed to the scarcity of relevant and 
granular enough data, a problem also highlighted by @iso2005 and
@kemeny2011international. As @zook2022mapping pointed, digital
activities are hardly ever captured in official data.

<!-- contribution -->
This paper offers such a contribution: an economic geography study
revealing the importance of spatial mechanisms for the diffusion of a
digital technology that is the Web in the UK at a high level
of spatial granularity 
<!-- NEW TO JUSTIFY DATA TIME PERIOD-->
from the very early days until its maturity (1996-2012).
<!-- NEW DATA CLARIFICATION -->
It specifically focuses on the diffusion of commercial websites.
It does so by employing a novel source of big data and machine learning
algorithms to model the *active* engagement with Web -- that is creating and 
maintaining instead of just browsing a website -- during that period
and tests how different spatial and temporal mechanisms shaped its
diffusion. This paper also exemplifies how the combination of data
sources and state-of-the-art research methods which escape the
traditional social science domain can offer new lenses to geographical
research regarding the understanding of technological diffusion.

<!-- motivation: stakeholders -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffuse over space and time and use this knowledge to make predictions
about the diffusion of related future technologies. As per
@leibowicz2016representing, historical studies agree that technologies
diffuse differently in terms of times, rates, and geographies and can be
driven by related policies [@victor1993]. @meade2021modelling highlight
that a variety of actors have a direct interest in gaining such
knowledge including network equipment suppliers, network operators,
regulatory and local authorities. These processes and their effects vary
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places [@leibowicz2016representing]. Despite the spatial heterogeneity
of such diffusion mechanisms and the policy relevance, there is a
scarcity of studies analysing the diffusion of new technologies
at a detailed geographical level because of the lack of detailed enough
data and adequate research methods.

<!-- motivation: regional development -->
This gap becomes even more prominent if we consider the importance of
technology and its adoption in economic geography and macroeconomic
growth theories. Simply put, differences in technological adoption and
sophistication may lead to uneven development as new technologies can
increase productivity and allow firms to introduce new products
[@solow1957technical; @aghion1990model; @kemeny2011international].
Linked to the subject of this paper, the Web and, more specifically, 
commercial websites can be considered as the underpinning
technology for the transformative process identified by
@capello2024nexus as the *digital service economy*. Importantly, the
focus here is not the Web in general, but instead commercial websites.
After all, diffusion together with invention and innovation are
considered the pillars of technological change [@das2022diffusion].
Findings from previous studies illustrated that early engagement with
digital technologies such as the Web led to longstanding positive
productivity gains for these regions that hosted early adopters
[@tranosuk]. Hence, obtaining a robust understanding of the spatial
diffusion mechanisms is crucial in mapping future development trajectories.

<!-- NEW motivation: intangible/digital -->
Another line of motivation has to do with the very nature of the Web as a 
digital technology and how it evolved over time. The digital geographies field 
has debunked premature claims about the Internet's placeless-ness by mapping 
various of its facets and technological layers [@tranos2013death], but the lack 
of relevant long time-series data prevented researchers from observing digital
phenomena overtime. Most of these studies focused on its hard infrastructure 
[e.g. @malecki2002economic; @moss2000internet], on Internet subscribers 
[e.g. @blank2018local], or on social media users [e.g. @crampton2013beyond].
Exceptions include early work from @zook2000web, @zook2001old and 
@moss1997tracking, which utilised domain name registration data and use it to
map and illustrate the spatial footprint of the web content. Pivotal as these 
contributions have been in understanding the economic geographies of the 
Internet, they were based on cross-sectional, often spatially detailed, mapping 
and on short temporal comparisons (e.g. 1993-1998) and did not observe the 
evolution and diffusion of the Web until its maturity. As we know from the 
relevant literature, different technologies diffuse at different rates and 
there is no theory that can explain these differences 
[@leibowicz2016representing]. Hence, mapping and modelling the diffusion of the
Web can offer useful insights.

<!-- early engagement with the web -->
This paper focuses on the diffusion of the Web as new technology during
the 1996-2012 period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
Internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer.[^1] Also, only
9 per cent of UK's households had access to the Internet in 1998
[@ons2018], the web included mostly static webpages, there were no
social media and web browsing involved exclusively a desktop PCs as
there were no smartphones [@tranosuk]. Hence, it is fair to say that the
study period captures the very early stages of the diffusion of a new
technology that is the Web until its maturity. The former is a key point
in the lifecycle of a new technology. Firstly, during this period new
technologies are expensive, crude and imperfect
[@rosenberg1994exploring; @wilson201281]. A simple comparison between
Web 1.0 and Web 3.0 applications clearly illustrates this
[@tranos2020social]: while a static website facilitates one-way
information dissemination, a platform like `github` enables cooperation
between users and the creation of new information, meaning, and even
knowledge [@faraj2016special; @barassi2012does]. During this period the
performance of a new technology is the main attraction and not the cost
to access and use it [@wilson2011lessons]. There is a broader
theoretical discussion in the literature about the early adoption
motives. As summarised by @perkins2005international, on one hand,
epidemic models highlight the role of interpersonal contacts as a way
for new technologies to diffuse. On the other hand, economic models
underline the importance of heterogeneity. Different firms have
different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories [@neffke2011regions; @simmie2014new]. This
argument is even more relevant when the focus is on digital technologies
because of the commonly found lag between investment and economic
returns as reflected in the Solow paradox [@acemoglu2014return;
@brynjolfsson2018artificial].

[^1]: <https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft>

<!-- data -->
The data used here depict the active engagement with the Web as they
contain geolocated and time-stamped counts of commercial websites from
the UK. Instead of adopting metrics reflecting the passive engagement
with the Internet such as Internet adoption and speeds [e.g.
@blank2018local; @destefano2022fuzzy], this paper observes yearly counts
of geolocated, commercial websites, which reflect the creation and 
maintenance of such websites. It needs to be highlighted here that although 
there are various types of underpinning web technologies -- namely markup,
stylesheet, client- and server-side coding languages as well as
different web browsers and other related programming language -- the
focus here is on the diffusion of the Web as an overarching technology
instead of more specific web technologies. For a such a detailed and 
cross-country technological diffusion study, see @PAPAGIANNIDIS2015308.

<!-- description of the methods and results -->
To analyse these data, a diverse set of methods are employed. Firstly,
cumulative adoption curves are estimated for the UK as a whole and for
the Local Authority Districts (LAD), illustrating, for the first time,
such S-shaped diffusion patterns for local areas. The exploratory spatial data
analysis offers some first indications of the importance of the different 
spatial diffusion mechanisms in shaping the diffusion of the Web in the UK.
Importantly, the analysis adopts two distinct geographical scales: the LAD and 
the most detailed scale of UK census areas, the Output Areas (OA). Then, a
novel methodological framework based on Machine Learning (ML) algorithms
is developed to test the role of such mechanisms at these two scales. The 
results reveal, for the first time, the importance of different spatial 
diffusion mechanisms in the diffusion of a digital technology at very granular
geographical scales. The analysis also reveals that the diffusion of the Web 
is characterised by spatial heterogeneity, well-established spatial patterns 
and both stability and volatility.

<!-- contents -->
The structure of the paper goes as follows. @sec-litreview reviews the
mechanisms that shape the diffusion of a new technology and sets up the two 
hypotheses that are empirically tested in @sec-results. @sec-datamethods 
describes the data and methods used in this paper. The discussion of the 
results and the conclusions of the paper are offered in @sec-conclusions.

# The mechanisms that shape the diffusion of a new technology {#sec-litreview}

<!-- The fundamentals of technology diffusion -->
The diffusion of new technologies and innovations is driven by three
different mechanisms [@grubler1990rise; @morrill2020spatial]: (i)
distance and proximity, (ii) urban hierarchies in the form of a
centrifugal forces as new technologies diffuse from core to periphery,
and (iii) over time the cumulative level of adoption follows an S-shaped
pattern.
<!-- neighbourhood -->
@hagerstrand1968innovation was the first to identify diffusion as a
geographical process. The starting point was the idea that diffusion is
based on passing information through social networks, which themselves
tend to be geographically bounded. Hence, he identified a
'neighbourhood' mechanism of how information, and consequently,
innovations and new technologies diffuse. This perspective draws
similarities with epidemics: new technologies and innovations, just like
pathogens, spread because of contagion and, consequently, proximity and
exposure [@hivner2003facilitating].

<!-- hierarchy -->
@hagerstrand1968innovation also emphasised the role of the top-down
hierarchy of urban systems and highlighted that new technologies and
innovations are firstly adopted in larger cities and then diffuse to
second tier ones. This is a sequential instead of a simultaneous
process, which resembles the 'lead-lag' spatial acceleration effect in
market research [@bento2018time; @PERES201091]. Schmidt's Law
empirically illustrates a similar pattern [@grubler1990rise]. Core and
highly agglomerated regions is where new technologies are invented and
commercially deployed. This is where the first adopters tend to be
based. Then, technologies spread to the rim and eventually to the
periphery. Although adoption pace might be higher when new technologies
finally arrive to the periphery, the saturation levels may never
reach the ones in the core because of the lack of infrastructure or
other necessary institutions [@leibowicz2016representing].

<!-- time -->
Diffusion is naturally a dynamic process and Hägerstrand recognised the
role time plays in the diffusion of innovations and new technologies: an
early-pioneering period, a middle fast accelerating period and a final
saturation one [@morrill2020spatial]. This was further explored by
@rogers2010diffusion, who identified five types of adopters of new
technologies and innovations: innovators, early adopters, early
majority, late majority and laggards. Hence, plotting the cumulative
adoption of a new technology over time often leads to an S-shaped line
plot [@grubler1990rise].

Although the above processes have been studied in innovation studies,
the distancing of economic geographers from research questions about the
diffusion of new technologies and innovations resulted to a knowledge gap
regarding the explanatory power of these mechanisms at granular, local
scales [@ding2010modeling]. As discussed in @sec-introduction, such 
spatially explicit knowledge can directly inform local technology
and infrastructure policies. Indeed, while the spatial dimension is
present in the following studies, the geographical level of analysis is
often too coarse to ascertain the role of the above mechanisms at the
local scale. @beardsell1999spatial studied the evolution of the computer
industry in 317 US metropolitan areas during the 1977-1992 period using
employment data. Their analysis indicated that the relative size
distribution holds for urban computer employment and also urban
heterogeneity is essential in explaining this distribution. In a recent
study, @bednarz2020pulled focused on wind turbines and modelled their
spatial diffusion across 402 German regions during 1970-2015. Their key
finding is that local demand than local supply was the main driving
factor. @haller2011determinants employed firm panel data for Ireland
during 2003-2005 and illustrated that when firms are located in a region
or industry with high share of firms having a website, they are more
likely to have a website too. At a global scale @perkins2011internet
explored whether the adoption of previous communication technologies
that is mail, telegrams and telephones were shaped by similar
socioeconomic factors as the Internet. Their results indicated common
patterns regarding the drivers behind the adoption of different
communication technologies. @feldman2015rknowledge focused on inventions
and used patent data to map how the the rDNA technology diffused among
inventors across 366 US metropolitan areas. @ding2010modeling modelled
the spatial diffusion of mobile telecommunications across the highest
level of administrative division in China (29 regions in total). Their
analysis indicated that socioeconomic characteristics are important
determinants of the timing, speed and the level of mobile diffusion
within China. @bakher2013diffusion compared different growth curves
as a means to model the diffusion of online shopping in Australia during
the 1998-2009 period, but they did not consider any spatial mechanisms.
At a global scale @PAPAGIANNIDIS2015308 modelled the diffusion of
different web technologies and practices across countries. Using similar
but less extensive data as the one used here, their analysis illustrated
how the diffusion of different web technologies and practices follows an
S-shaped pattern as well as the different diffusion rates of the
different technologies and practices. The most detailed consideration of
the spatial footprint of the diffusion of a new technology comes from
@lengyel2020role, who analysed the adoption and churn of a Hungarian
online social network. Their results were in agreement with the early
theoretical and empirical contributions reviewed here: assortativity,
urban scaling and distance are the key drivers of spatial diffusion.
Based on the above, the main empirical hypothesis that is tested in this paper
states:

> H~1~: The three diffusion mechanisms namely distance, urban hierarchy
and the S-shaped pattern of the cumulative level of adoption shaped
the diffusion of the commercial Web in the UK even at granular geographical 
scales.

<!-- EARLY/LATE adoption-->
Understanding the diffusion of a new technology in multiple, detailed spatial
scales is of interest to economic geography as potential local economic 
benefits may occur because of concentration of early adopters or through 
technological leapfrogging. On the one hand, early adopters are often rewarded 
because of their attitudes towards new technologies and innovations. 
@rogers2010diffusion identifies them as 'knowledgeable risk takers' and 
@griliches1957 as 'profit maximisers' [@ding2010modeling]. On the other hand,
latecomers can adopt and benefit from new technologies that have been developed
elsewhere without incurring the hefty initial R&D costs [@teece2008firm]. At a 
global scale @perkins2005international explored whether the diffusion rate of 
new technologies is driven by a latecomer advantage and the engagement with the 
global economy via foreign direct investments and trade. Their results 
illustrate that indeed latecomers and developing countries experience diffusion 
of new technologies more rapidly than early adopters and developed countries. 
At a regional scale, previous research highlighted the long term and 
sustained productivity benefit of the early adopters of web technologies in 
the UK [@tranosuk]. Although the early adoption and leapfrogging literature has 
not paid much attention on cities and regions [@yu2018sustainability], the 
potential local economic benefits justify the need to investigate the
dynamics of the web diffusion in different scales in the UK. Hence, the second 
empirical hypothesis states:

> H~2~: The diffusion of the commercial Web is shaped by both early adoption 
and latecomer effects.

These two hypotheses are empirically tested in @sec-results employing 
the data and methods described in the next section.

<!-- Technological diffusion, which is by definition an aggregated process, -->
<!-- can be discussed in parallel with individual adoption mechanisms. On the -->
<!-- one hand,  -->
<!-- On the other hand, @perkins2011internet attribute diffusion -->
<!-- to two processes: (i) epidemic-like mechanisms, which are governed by -->
<!-- distance, proximity and social interactions, and (ii) economic -->
<!-- mechanisms as new innovations are adopted by users as they become more -->
<!-- profitable, valueable and useful. -->

# Data and Methods {#sec-datamethods}

<!-- main data -->
To capture the diffusion of the commercial Web, a website density metric is
developed for two different geographical scales that are discussed later in 
this section. The main data are counts of commercial websites that are 
calculated using data from the Internet Archive[^2] and, specifically, the 
JISC UK Web Domain Dataset [@ukwebarchive]. The Internet Archive is one of the 
most complete and oldest archive of webpages in the world operating since
1996 [@ainsworth2011much; @holzmann2016dawn]. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage it
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK country code top level
domain (ccTLD - .uk) from the 1996–2012 period. In essence, this is a
long list of 2.5 billion URLs of archived webpages including also the
archival timestamp.

[^2]: [https://archive.org/](https://archive.org/).

Instead of using the whole .uk ccTLD, this paper focuses on its
commercial subset, the .co.uk second level domain (SLD). This choice
is aligned with the topic of the paper -- the diffusion of the Web for 
commercial functions -- and decreases the heterogeneity of the web data as such 
commercial websites have specific aims: they are used to diffuse information, 
support online transactions and share opinions [@THELWALL2000441; 
@blazquez2018big]. Although a UK company can adopt a generic TLD such as .com 
and these cases escape the data used here, such omissions should not affect the
validity of the results given the popularity of the .uk ccTLD
[@tranosuk]: UK consumers prefer to visit a .uk website when they are
searching for products or services [@hope]; and anecdotal evidence
indicates that during the first half of 2000, three .co.uk domains were
registered every minute [@oecd_coms]. Importantly, previous studies
illustrated that .co.uk is the most popular UK SLD [@tranosuk]. Henceforward 
any reference to websites will refer to commercial -- .co.uk -- websites.

The text from these webpages was scanned using a regular expression
(regex) to identify strings of text which resemble UK postcodes and one
fifth of them included a mention to a postcode [@BL2013geo]. This
information allows the geolocation of the data and the creation of the
LAD and OA counts. To prevent false positives, the resulted postcodes
were matched against official postcode lookup files [@ons].

The data cleaning process includes an aggregation step, through which
the archived webpages are aggregated to the parent websites. This
website reconstruction process lend itself to the construction of
*website* instead of *webpage* density metrics. Websites tend to
represent specific organisations or entities, and, arguably, are more
meaningful observational units than webpages, which can ignore the
upstream dependency of the website they belong to. Based on the
following example, all three webpages are part of the same website
(<http://www.website.co.uk>). Only the overall website and not the
nested webpages are considered for the counts as otherwise the density
metrics would have been biased towards large, place specific, websites.
It is important to note here that websites commonly contain place references, 
with previous research estimating this share to be around 70% 
[@hill2009georeferencing].

-   <http://www.website.co.uk/webpage_a> B15 2TT

-   <http://www.website.co.uk/webpage_b> BS8 1TH

-   <http://www.website.co.uk/webpage_c> B15 2TT

\noindent What is challenging is that this aggregation approach, which
has been used elsewhere [@tranosuk; @shoreditch] may lead to websites
with multiple postcodes. As per the above example,
[www.website.co.uk](http://www.website.co.uk) includes two unique
postcodes: B15 2TT and BS8 1SS. The distribution of the number of
postcodes per website for 2000 is presented in Table \ref{f2000}, which
illustrates a wide range. At the left end of the distribution, there are
websites anchored to a unique location (72% of all the reconstructed
websites in 2000), which may represent a small company with a single
trading location. At the right end, there are websites with thousands of
different postcodes. Considering the time period of the analysis, such
cases can represent directories which used to be popular in the
pre-search engines early times of the commercial Internet as well as
real estate websites [@tranosuk]. Previous research which used the same 
methodology for a small subset of these data illustrates the robustness of the 
above approach: its results matched in terms of accuracy past rounds of 
research based on extensive qualitative work, interviews and observational 
studies for the same study area [@shoreditch].

The analysis is firstly conducted on websites, which only contain one
unique postcode. As a robustness check, the analysis is replicated for
an extended subset of websites, which include up to 10 unique postcodes
to capture commercial websites which point to multiple locations. They
are geolocated by equally attaching them to these locations. This
extended sample includes 94 percent of all the archived websites in
2000 (see Table \ref{f2000}).

```{r eval=T, echo=FALSE, results='asis'}

# load 2000 co.uk
# path.2000 <- "C:/Users/nw19521/OneDrive - University of Bristol/projects/archive/nuts/all2000couk.csv"
# path.2000 <- "/Users/nw19521/Library/CloudStorage/OneDrive-UniversityofBristol/projects/archive/nuts/all2000couk.csv"
path.2000 <- paste0(path, "/data/temp/all2000couk.csv")

all2000.duplicates <- data.table::fread(path.2000) #
# dim(all2000.duplicates) # 3336162
# it includes duplicated URLs: if one webpages includes multiple postcodes
# then it appears multiple times. This is ok for the nuts aggregation, but
# not for the frequencies
# This is only .co.uk

# one line for every host
all2000 <- unique(all2000.duplicates, by = c("host")) # 57897

# unique postcodes per website f table
f.websites.pc <- DescTools::Freq(all2000$V1, breaks = c(0, 1,2, 10,100,1000,10000,100000), ord = "desc")
f.websites.pc$level <- factor(f.websites.pc$level, levels = c("[0,1]","(1,2]", "(2,10]", "(10,100]",
                                                              "(100,1e+03]", "(1e+03,1e+04]",
                                                              "(1e+04,1e+05]"))
levels(f.websites.pc$level) <- c("(0,1]","(1,2]", "(2,10]", "(10,100]", "(100,1000]", "(1000,10000]", "(10000,100000]")

f.websites.pc <- f.websites.pc %>% 
  dplyr::select(-cumfreq) %>% 
  rename(Postcodes = level,
         F = freq,
         'F (%)' = perc,
         'cumulative F' = cumperc)

kable(f.websites.pc,
      format = "latex",
      digits = 3,
      booktabs = T,
      format.args = list(big.mark = ","),
      caption = "Number of unique postcodes per .co.uk website, 2000.\\label{f2000}") %>%
  kableExtra::footnote(general = "Tranos et al. 2021",
                       general_title = "Source: ",
                       footnote_as_chunk = T)  
   # kable_classic(full_width = F, html_font = "Cambria")
```

Another data cleaning step deals with some extreme outliers. Figure
\ref{correct} plots the website counts for the top 1000 postcodes. Some
obvious outliers can be observed for the 2002-2006 period for a handful
of postcodes, which can be attributed to link farms [@BL2013links]. The
website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1,000
websites pointing to them, were replaced with predicted values based on
a simple panel regression model with postcode fixed effects and yearly
dummy variables. These postcodes refer to a residential area with a
small park in South London, residential areas outside Warwick and
Manchester, a rural area in the south border of England and Wales and
small business parks in Crew and Derby. Hence, it is difficult to
believe that these postcodes genuinely hosted such a large amounts of
websites, which were later extensively declined. To put the magnitude of
the data imputation into perspective, this process affected 6 out of the
557,808 postcodes included in the data. As @sec-results illustrates,
this led to a small increase of the model predictive power.

```{r echo=FALSE, message=FALSE, fig.cap="\\label{correct}Yearly website counts per postcode (top 1000) before (top) and after (down) data imputation", out.height="70%"}
path.image1 <- paste0(path, "/outputs/pc_year_1.png")
path.image2 <- paste0(path, "/outputs/pc_year_1_corrected.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)

par(mar=rep(0,4)) # no margins

p1 + plot_spacer() + p2 + plot_layout(ncol = 2)#, widths = c(1, -0.8, 1))
```

<!-- Multi-scale -->
To empirically test the two hypotheses stated in @sec-litreview, the website 
density metric was calculated for two different scales the Local Authority
Districts (LAD) and the Output Areas (OA). The former is the lowest
administrative unit in the UK and there are 374 LAD. The latter is the smallest
census-based geographical unit and there are c. 230,000 of them in the UK. To 
create the website density metric, the yearly website counts at the LAD level 
are standardised by the number of firms in LAD to avoid biases associated with 
LAD hosting a large number of firms. Given that there is no such readily 
available statistic for the OA, the actual OA level counts are used. However, 
because of the the consistent spatial definition of OA (they host 40-250
households),[^3] website counts in OA are interpreted as a density metric too.

[^3]: <https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies>

<!-- S -->
These website density metrics are used for the different stages of the analysis. 
Starting with the first hypothesis (H~1~), to test whether the S-curve 
accurately describes the cumulative adoption of the Web at a local scale, the
following logistic function (Equation \ref{eq:s}) is estimated for the whole of 
the UK and then for each LAD separately:

\begin{align}
y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
\end{align}

\noindent $k$ is the asymptote or, in other words, the saturation level,
$b$ the overall growth rate, and $t_{0}$ the *inflection point* of
maximum growth at $k/2$, where the logistic function is symmetrical
[@wilson201281]. The $t_0$ of each LAD is compared against the $t_0$ of
the UK to delineate whether a LAD reached that point faster or slower
than the country average. This heterogeneity is discussed further when testing 
the second hypothesis (H~2~). Importantly, an accuracy criterion was imposed
and only LAD with $R^2 > 0.9$ were included in this analysis. To
estimate Equation \ref{eq:s} a self-starting logistic growth model was
employed using the `nls` and `SSlogis` functions in `R`. This logistic
growth model depicts the initial period of slow diffusion of a new
technology, which is followed by the fast, exponential growth stage to
then eventually slow down and saturate [@wilson201281;
@grubler1999dynamics].

<!-- **minimum R2 = 95%** in @wilson201281, see also @grubler1990rise -->
<!-- The literature usually uses the saturation level as the asymptote. I am -->
<!-- using the total number of websites as we cannot compute a rate. -->

<!-- ESDA -->
Next, exploratory spatial data analysis provides some first insights of whether 
the other two drivers of spatial diffusion -- namely neighbourhood and 
hierarchy -- underpin the diffusion of the Web in the UK. To capture the former 
and following @ding2010modeling the Moran's I and the Local Index of Spatial
Association (LISA) are estimated for the website density. To capture the
hierarchy mechanism, the Gini coefficient -- an established metric of
inequality -- is calculated. All the above are computed and plotted
longitudinally both for the LAD and the OA.

<!-- conceptual model -->
To formally test H~1~ a modelling framework is developed to assess the 
above diffusion mechanisms. The overarching aim is to build a model that can 
test how well these three mechanisms can predict website density:

\begin{align}
Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S\text-curve_{t}\label{eq:rf.generic}
\end{align}

<!-- <!-- conceptual model --> 
<!-- The models include variables that capture the three processes that the -->
<!-- relevant literature and the descriptive analysis highlighted. Namely, -->
<!-- the models capture: (i) a hierarchical mechanism with diffusion running -->
<!-- from main centres to secondary ones, (ii) a neighborhood mechanism -->
<!-- according to which diffusion first hits nearby locations, and (iii) the -->
<!-- rather canonical pattern of diffusion over time as reflected in the -->
<!-- S-shaped pattern in the cumulative level of adoption. -->

<!-- detailed features -->
To capture the hierarchical mechanism the model include as predictors a
one year lag of website density in London, the largest city in the UK, and a
one year lag of the website density in the nearest city. Both variables clearly
illustrate the urban hierarchy in the UK. Especially for the OA models, a 
third predictor is added to capture hierarchy at a much more granular scale:
a one year lag of the website density in the the nearest retail centre [^4]. As
the literature illustrates, retail centres are key components of urban systems
[@dennis2002central; @jones2021reframing]. Due to the small sizes of the retail 
centres, the latter is only relevant for and included in the OA models and not 
in the LAD models. In addition, the model includes the distance to London, the 
nearest city and the nearest retail centre. The underlying logic is that the 
level of website adoption in a spatial unit depends on the level of the 
adoption in places further up in the urban hierarchy the previous year. The
inclusion of the distance variables incorporates spatial structure into
the hierarchy argument. To depict the neighbourhood mechanism, the
website density of the neighbouring spatial units in the previous year
is employed. This is calculated using rook continuity. Again, the
underpinning rationale is that the level of web adoption within a
spatial unit depends on the level of web adoption in the neighbouring
spatial units the year before. This represents the 'hitting nearby
locations first' argument. Therefore, the spatial and temporal lag of
the website density in LAD and OA is calculated. Lastly, the temporal
pattern of the cumulative adoption is captured by a time trend variable.
Hence, the model will follow the following generic form (Eq.
\ref{eq:model}):

[^4]: The retail centres data have been obtained by the Consumer Data Research Centre, an ESRC data investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1: <https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators>.

\begin{align} \label{eq:model}
Website\,Density_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}

<!-- 4 models -->
Specifically, four different models are estimated. Firstly, all the data points 
for the OA and LAD are utilised in order to build two models and assess their
capacity to predict the adoption of the Web at the two different scales. These 
two models will reveal the predictive capacity of the diffusion mechanisms and 
also demonstrate how the importance of such variables changes across
different scales. The next two sets of models are again trained at these
two scales: OA and LAD. However, instead of using all the data points,
the OA and the LAD from one of the twelve UK regions are held out for
the model training. Then, the trained model is used to predict website
density for the OA or the LAD of the held-out region. This process takes
place recursively for all twelve UK regions. The regional differences in
the predictive capacity of the different samples will reveal how
dissimilar these spatial process are across regions and, importantly, at
different scales.

<!-- RF -->
Equation \ref{eq:model} is estimated using Random Forest
(RF). This is a popular ML algorithm for both regression and
classification problems [@biau2012analysis]. It was introduced by
@breiman2001random and has become a go-to data science tool. RF can
effectively handle skewed distributions and outliers, model non-linear
relationships, require minimal hyperparameter tuning, exhibit low
sensitivity to these parameters, and have relatively short training
times [@Caruana2008; @liaw2002classification; @yan2020using]. These
attributes match well with the website density data characteristics
including skewness especially for the OA. Also, the large data size (c.
230k data points for each of the 17 years) calls for fast training
times. Importantly, RF predictions tend to be more accurate than those
from single regression trees and outperform Ordinary Least Squares in
out-of-sample predictions, even with moderate-sized training data and a
small number of predictors [@mullainathan2017machine; @athey2019machine;
@sulaiman2011intelligent; @pourebrahim2019trip; @biau2012analysis].

<!-- RF -->
RF is a tree-based ensemble learning algorithm [@breiman2001random]. It
begins by generating random samples of the training data, which are then
used to grow regression trees to predict the dependent variable. Data
points and predictors are randomly sampled for the different trees. The
trees are trained in parallel using their own bootstrapped samples of
the training data. A crucial feature of RF is their ability not to
overfit, meaning they can generalize well to unseen test data. While
each tree may overfit individually, the ensemble of trees does not
because the errors of individual trees are averaged, reducing the
overall variance and preventing overfitting [@last2002improving]. For
regression problems, RF predictions are made by averaging the
predictions of all decision trees.

RF have been widely employed to address regression research problems.
@pourebrahim2019trip combined a spatial interaction modelling framework
with ML algorithms including RF to predict commuting flows in New York
City. @sinha2019assessing advocated for adopting spatial ensemble
learning approaches, such as RF, to model spatial data with high
autocorrelation and heterogeneity. @creditspatial predicted employment
density in Los Angeles using spatially explicit RF.
@guns2014recommending used RF to build a recommendation system for
research collaborations. @ren2019predicting trained RF to predict the
socio-economic status of cities using various online and mobility
predictors. @tranos2023using utilised hyperlinks data and RF to make
out-of-sample predictions of interregional trade. @zhou2023geography
employed such a framework to assess whether key predictors of obesity
differ across English cities.

<!-- CV -->
It needs to be highlighted here that the cross validation (CV) for all
models is spatially and temporally sensitive. Instead of using 10 random
and space- and time-agnostic folds, the `CAST` package is employed which
allows holding back data points from specific years and spatial units (regions)
and use them for testing in order to estimate the model performance
[@meyer2018improving].

<!-- assessment -->
To assess the predictive power of the model, three broadly utilised
metrics are employed: the coefficient of determination ($R^2$), mean
absolute error (MAE) and root mean square error (RMSE):

\begin{align}
R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
\end{align}
\begin{align}
MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
\end{align}
\begin{align}
RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
\end{align}

\noindent $y_{k}$ is the $k^{th}$ observation of the dataset, which
consists of $N$ observations in total. $\hat{y_{k}}$ is the $k_{th}$
predicted value for the dependent variable and $\overline{y_{k}}$ is the
average value of $y$. The last two metrics are expressed in the same
units as the dependent variable -- websites per firm for the LAD modes
and the number of websites for the OA models -- while the first one is
the coefficient of determination between the observed and the predicted
values of website adoption. Regarding $MAE$, it is the absolute
difference between the observed and the predicted website adoption.
While $MAE$ does not penalise for large errors, $RSME$ does, as it is
proportional to the squared difference between the observed and the
predicted trade flows. Hence larger errors weigh more for $RMSE$
[@pontius2008components].

An alternative specification could model the growth of website density. However, 
such a strategy is problematic because of the large frequency of 
OA without websites, which is to be expected given their small size and high 
granularity. Still, as a robustness text, the results of a growth model for
LAD are presented in the Appendix.

<!-- H2 -->
Lastly, to test the second hypothesis (H~2~) about the early adoption and the
latecomer effects, a different research framework is developed. Firstly, the 
S-curve and the estimated $t_0$ for every LAD are employed and mapped in order 
to depict the spatial heterogeneity and specific patterns of the speed of the 
diffusion of the Web. Then, the volatility and the stability of the diffusion 
across the different LAD is analysed using tools from the complex systems domain 
[@Batty2009] namely changes in rankings of LAD based on website density. Both 
the S-curve estimation and the volatility analysis focus only on LAD as the 
very large number of OA would have made such analysis difficult to visualise 
and interpret.

# Results {#sec-results}

## The three diffusion mechanisms at granular geographical scales

<!-- S-shaped -->
Figure \ref{s_uk} plots the cumulative adoption of the Web in the UK as a whole
during the 1996-2012 period. In line with the theoretical expectations discussed
in the previous section, the cumulative adoption of the Web is well represented 
by a curve that resembles the letter S. The vertical line in the beginning of 
2003 (2003.158) illustrates the point when the modelled cumulative adoption was 
equal to 50% of the maximum. This $t_0$ point is known as the *inflection* point 
and signals the maximum adoption speed. The results remain unchanged when the 
extended subset of websites, which include up to 10 unique postcodes is used
(see Figure \ref{s_uk10} in the Appendix). Importantly though, the extended
subset of websites indicates a saturation level close to 100% at the end of 
the study period.

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk}Cumulative adoption of the Web, UK"}
path.image <- paste0(path, "/outputs/s/s_uk_per_firm.png")
knitr::include_graphics(path.image)
```

What escaped from previous studies was εexploring whether such a pattern also 
describes cumulative adoption of new technologies at sub-national scales. To 
test this, the same  logistic function is estimated for each of the 374 LAD. The
detailed results are presented in Table \ref{table.s.lads} in the Appendix. 
The key finding is that for 342 LAD (91% of all LAD) the $R^2 > 0.9$ for the 
logistic function, which indicates that even at local scale the cumulative 
adoption of the Web is characterised by a pattern similar to the one observed 
for countries. Graphically, all these LAD curves are plotted in Figure 
\ref{s_lad_uk}. The different colour lines illustrate whether a LAD reached $t_0$ 
faster or slower than the UK in total. This heterogeneity is further discussed 
later in this section.

This is the first time that such cumulative adoption curves have been estimated
for areal units as small as the UK LAD. Understanding and modelling the dynamics
of technological diffusion and, ultimately, being able to predict diffusion 
rates at such small local scales can directly support the deployment of 
infrastructural networks (e.g. 5G) and local technology policies. The above are 
in line with previous studies exploring the diffusion of web technologies 
[@PAPAGIANNIDIS2015308], social media [@lengyel2020role] and online
shopping [@bakher2013diffusion].

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_lad_uk}Cumulative adoption of the Web, LAD"}
path.image <- paste0(path, "/outputs/s/lad_per_firm_all.png")
knitr::include_graphics(path.image)
```

<!-- Neighbourhood effect -->
To assess the second diffusion mechanisms, Figures \ref{morani} and \ref{lisa} 
offer a first insight into whether a neighbourhood or, in other words, a 
distance mechanism underpins the diffusion of the Web in the UK as they present 
the Moran's I and the LISA maps of website density respectively for LAD and OA.
Starting from the former, spatial autocorrelation was higher in the beginning of 
the study period and then after 2000 dropped slightly and stabilised around 0.2. 
This reflects the
early concentration of high website density around London, which over
time diffused as high website density clusters can be seen in other
parts of the country away from London (Figure \ref{lisa}). An almost
reverse pattern can be observed for OA. At the beginning of the study
period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
Because of the very small size of OA, at the early stages of the
diffusion of web technologies their adoption was spatially scattered.
This is reflected in the lack of any significant clusters in 1996
(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
such clusters of high website density were formed and this is reflected
both in the Moran's I and the LISA maps in Figures \ref{morani} and
\ref{lisa}. A similar pattern is observed for the extended dataset (see
Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
spatial autocorrelation is slightly higher illustrating a stronger
spatial clustering mechanism. All in all, the exploratory spatial data
analysis advocates towards an underpinning neighbourhood mechanism, which 
is present at different scales. Interestingly, the magnitude of Moran's I
is similar to the ones revealed by @ding2010modeling about mobile phone
adoption in China albeit the much more detailed spatial scale of this
paper.

```{r, morani, echo=FALSE, message=FALSE, fig.cap="\\label{morani}Website density Moran's I"}

path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa.csv")

la.m <- read_csv(path.in.la) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'LAD')
oa.m <- read_csv(path.in.oa) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'OA')

bind_rows(la.m, oa.m) %>% 
  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Moran\'s I') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa}\\centering Website density LISA maps", out.height="80%"}

path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la1996.png")
path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2000.png")
path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la2012.png")
path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa1996.png")
path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2000.png")
path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc1_oa2012.png")
path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)  
p3 <- ggdraw() + draw_image(path.image3)  
p4 <- ggdraw() + draw_image(path.image4) 
p5 <- ggdraw() + draw_image(path.image5)
p6 <- ggdraw() + draw_image(path.image6)  
p7 <- ggdraw() + draw_image(path.image7)

par(mar=rep(0,4)) # no margins

library(patchwork)
# (p1 | p4)/(p2 | p5)/(p3 | p6)/ p7 +
#   #plot_layout(widths = 1)
#     plot_layout(widths = c(1, -0.5))

# p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 / p7 +
#   plot_layout(ncol = 3, widths = c(1, -0.85, 1), heights = c(1,1,1,2))

p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))

# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)
```

To illustrate whether a hierarchical process also underpins the
diffusion of web technologies -- the third diffusion mechanism that is tested 
here -- the Gini coefficient is calculated yearly
both for LAD and OA. As a metric of inequality, the Gini coefficient
demonstrates whether website density is concentrated in a small number
of LAD or OA, or whether it is more equally spread across the country.
Both scales of analysis in Figure \ref{gini} illustrate the same
picture. At the beginning of the commercial Internet, website density
was extremely unequal, or, in other words, only a few places had
websites associated with them. Inequality dropped and plateaued after
2000 for both scales. This is illustrative of a hierarchical diffusion
mechanism that led over time to a more equal spread. Interestingly, the
year 2000 is again a period of change for this diffusion mechanism as it
was for the neighbourhood process. There is a substantial difference
between the Gini coefficient magnitude for LAD and OA, but this is
expected as the very small size of OA equates to a lot of polygons
without any websites pointing to them -- for example residential OA.

```{r, gini, echo=FALSE, message=FALSE, fig.cap="\\label{gini}Website density Gini coefficient"}

path.in.la <- paste0(path, "/outputs/gini/gini_la.csv")
path.in.oa <- paste0(path, "/outputs/gini/gini_oa.csv")

la.gini <- read_csv(path.in.la) %>% 
  mutate(spatial.unit = 'LAD')
oa.gini <- read_csv(path.in.oa) %>% 
  mutate(spatial.unit = 'OA')

bind_rows(la.gini, oa.gini) %>%
  rename(Gini = 'gini(a$n)') %>% 
  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Gini') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

After discussing the descriptive findings, the validity of the above diffusion
mechanisms at the local scale is assessed with a predictive modelling framework
described in @sec-datamethods. The reader is reminded that four models are 
estimated: for the OA and LAD, firstly all data points are employed to 
assess the capacity of the three mechanisms to to predict the adoption of the 
Web at these two scales; then, the LAD/OA from one of the twelve regions are
recursively held out to assess the spatial heterogeneity of these mechanisms
across the different UK regions.

@tbl-model-metrics presents the model performance for the first set of
models, for which all data points are employed for training and testing
via CV. The first one is trained and tested on 374 LAD and the second on
232,296 OA, both for the 16 year period (1997-2012). The results are
remarkably good considering that the are the outcome of space and time
sensitive CV, so the the model does not suffer from overfitting. At the
LAD level the model predicts 83% of the variation of website density.
Both error metrics indicate that the model error is fraction of a website 
per firm (0.020 and 0.029). At the OA the $R^2$ drops down to 33%. Considering 
its granularity, this is still a remarkable performance. To contextualise it, 
the model results in a $MAE$ equal to 1.2 websites for areas small enough to 
host 40-250 households. Because of the small size of the spatial units, the
distribution is highly skewed and a significant part of them is not
linked to any websites. In 1997 only 1% of the UK OA were associated
with at least one website. This should not come as a surprise given that this
was the very beginning of the commercial Internet and any activities
with a digital footprint were concentrated in a handful of areas, as 
illustrated in Figure \ref{lisa}. At the end of the study
period almost half of the UK OA were not associated with a website.
Again, given the granularity of the data this was expected.

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.029  | 0.824   | 0.020 |
| Output Areas      | 3.218  | 0.333   | 1.233 |

: Model metrics {#tbl-model-metrics}

The results in @tbl-model-metrics are robust against different
specifications. As noted earlier, the data imputation process resulted in
a modest improvement of the predictive capacity of the model. As per
@tbl-model-metrics-no-correction, all accuracy metrics are only slightly
worst when the data imputation is not applied. As expected, the
improvement is higher for the more detailed scale of analysis. Then,
@tbl-model-metrics-10 presents the model results for the extended
dataset, which includes websites with up to 10 unique postcodes. While
the results for LAD are only slightly less accurate, the $R^{2}$ of the
OA model is much higher in comparison to the base results in
@tbl-model-metrics. In addition, @tbl-model-growth-metrics in the Appendix
presents the accuracy results of a LAD model, the dependent variable of which 
is the yearly growth rate of website density. Even for such a specification
(see Equation \ref{model.growth} in the Appendix) the predictive capacity of
the model is exceptionally high ($R^2 = 0.5$). All in all, the modelling exercise
revealed how well the three diffusion mechanisms predict the diffusion of the 
Web as a new technology. Importantly, this predictive power is highly robust
against against different scales and delineations of the diffusion of
the web.

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.032  | 0.781   | 0.021 |
| Output Areas      | 4.644  | 0.244   | 1.237 |

: Model metrics without data imputation {#tbl-model-metrics-no-correction}

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities |  0.190 | 0.749   | 0.129 |
| Output Areas      | 21.591 | 0.455   | 6.845 |

: Model metrics for the extended dataset {#tbl-model-metrics-10}

Figure \ref{var.imp} plots the importance of the different predictors.
When the focus is on LAD, the website density in the neighbouring LAD, in the
nearest city and in London the year before are the most important predictors. 
They are followed by the yearly trend, while the spatial configuration as 
reflected in distances to London or to the nearest city only plays a minor 
role. This can be attributed to the relatively coarse spatial scale of LAD.
Nevertheless, all previously discussed spatial processes are at play in the
diffusion of the Web at the LAD level: the first two predictors depict 
the hierarchical mechanism, the spatial and temporal lag of website density the
neighbourhood mechanism, and the yearly trend the time-sensitive
cumulative adoption pattern.

At the much more granular scale of OA, the picture is
almost reversed. The two most important predictors are the distance to
London and to the nearest city followed by the spatial and temporal lag
of website density and the distance to the nearest retail centre. They
still depict a hierarchical mechanism, but proximity to the different
population centres is more important than their lagged web densities in
predicting website diffusion. The neighbourhood mechanism is still
strong but less important at this scale. What is interesting is the
almost negligible role of the yearly trend and London's website density.
While the former probably illustrates the large heterogeneity in how the
Web has been adopted at this very fine scale, the latter highlights that
the importance of past web adoption rates in large population centres is
surpassed by proximity to them and spatial configuration at this scale.

The above pattern of variable importance is largely in accordance with
the models estimated for the extended dataset. A comparison with Figure
\ref{var.imp10} illustrates that the main difference is the importance
of the neighbourhood mechanism for the OA as the space and time lag of
the website density is by far the most important predictor when the
extended dataset is employed. For the LAD, the importance of the distance to
the nearest city increases in favour to its website density the previous
year. But still, the variable importance patterns for the two datasets 
are not dissimilar.

```{r varimp, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp}Variable importance", out.height="60%"}
path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected.csv")
path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected.csv")

importance.oa <- read_csv(path.importance.oa) %>% 
  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
  mutate(variable = str_to_sentence(variable))
importance.lad <- read_csv(path.importance.lad) %>% 
  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
  mutate(variable = str_to_sentence(variable))

importance.oa %>% 
  rename(OA = importance) %>% 
  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
  dplyr::select(-variable.y, -id) %>% 
  rename(variable = variable.x) %>% 
  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
  mutate(order = ifelse(variable == "Distance to london", 1,
                        ifelse(variable == "London's website density, t-1", 2,
                               ifelse(variable == "Distance to the nearest city", 3,
                                      ifelse(variable == "Nearest city's website density, t-1", 4,
                                             ifelse(variable == "Distance to the nearest retail centre", 5,
                                                    ifelse(variable == "Nearest retail centre's website density, t-1", 6,
                                                           ifelse(variable == "Spatial and temporal lag of website density", 7, 8)))))))) %>% 
  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
         fill = spatial.unit)) +
  geom_col(position = 'dodge') +
  ylab('') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit: ', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous('Variable importance') +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7))
```

Table \ref{table.regions} presents the results of the recursive hold out
models and allows to (i) highlight the regional heterogeneity of the
spatial processes behind the diffusion of the Web, and (ii) compare this
heterogeneity across the two different scales of analysis. It illustrates both 
regional similarities and differences. At the top of Table \ref{table.regions}
the regions of South East and the West Midlands have the highest similarity with 
the rest of the country. In other words, the web density for the LAD of these 
regions can be very well predicted by models trained on data from the LAD from 
the rest of the country and, therefore, the spatial diffusion mechanisms for 
these regions represent quite well the UK's average. This should not come as a 
surprise as the spatial structure of both regions resembles the broader UK: 
both regions contain some highly urbanised LAD (South East is well integrated 
with the Greater London and West Midlands is the home of the city of Birmingham), 
but also smaller commuter towns as well as more rural settlements. At the other 
end of the spectrum, the spatial diffusion mechanisms for Northern Ireland, 
Scotland and North West diverge the most from the country average. Rurality and 
remoteness seem to be common attributes for these regions. This hierarchy though 
is not consistent across scales. For the more granular OA models, the spatial 
diffusion mechanisms for South West, Yorkshire and The Humber and North West 
regions are closer to the country average. Interestingly, these regions are 
well-known tourist destinations (Lake District, Peak District and 
Cornwall-Devon) and the exploratory analysis revealed clusters of high web 
density only visible at the OA scale (Figure \ref{lisa}). Then, Scotland, 
Northern Ireland and London have the lowest similarity with the country 
average. Again, rurality, peripherality, but also the highest level or 
urbanisation make the spatial diffusion mechanisms of these regions more 
distinct.

```{r regions}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(caption= "Regional differences\\label{table.regions}", 
        digits=3,
        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
```

Comparing the above with Table \ref{table.regions.10} in the Appendix, which 
presents the results for the the extended dataset with websites with up to 10 
unique postcodes, the results do not differentiate much, signalling the 
robustness of the analysis. The most salient difference has to do with Northern 
Ireland. According to the extended sample at the granular OA level, Northern 
Ireland's spatial diffusion mechanisms are closer to the UK's average than the 
results based on the websites with a unique postcode. While the latter are 
websites and, consequently, economic activities solely based in Northern 
Ireland, websites with multiple postcodes also include websites with postcodes 
in Britain and, therefore, can be characterised by slightly different spatial 
diffusion processes.

## Early adoption and latecomers

<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
<!-- and latecomer effects. -->

To test the second hypothesis (H~2~), the spatial heterogeneity of Web adoption 
speeds is firstly analysed. As highlighted in the literature [@risk_perceptions], 
different agents have different perceptions about and levels of acceptance of 
the risks and the potential economic returns associated with the adoption of 
new technologies -- see for instance the seminal work of @venkatesh2000theoretical.
To reveal such patterns, Figure \ref{s_map} maps whether LAD reached their 
inflection point earlier or later than the UK in total and based on this are 
labelled as *fast* or *slow* accordingly. Inflection points have been used in
the relevant literature to compare diffusion processes across different settings 
and technologies [@cherp2021national; @woo2017exploring] as they can enable more 
insightful comparisons than, for example, the average growth over the 1996-2012
study period.

The map illustrates a clear high concentration of fast adopting LAD in the 
South and East of England, as expected. There are also some concentrations in 
Wales, Scotland and the North West of England. At a more detailed level, there 
are some expected cases of LAD with relevant industrial backgrounds delineated 
as fast: the City of London, a world-renowned cluster of finance industries
[@cook2007role], and Reading, a town with high-tech service industries
in proximity to London and its main airport, Heathrow
[@england2005polynet]. However, there are also LAD which were expected
to appear as fast -- e.g. Hackney in central London and Bristol, a
well-established creative cluster [@oatley1999cultural;
@bassett2002cultural] -- but are delineated as slow. Similar to 
@beardsell1999spatial, who analysed the spatial evolution of computer industry 
employment across 317 US metro areas, spatial heterogeneity is evident in the 
diffusion speed when the scale of analysis is as detailed as the LAD.

Importantly, there is obvious spatial clustering of fast and slow LAD. Out of 
the ten fastest LAD based on when they reached their inflection point (see 
detailed Table \ref{table.s.lads} in the Appendix), nine were located in the 
South East of England and London and one in Cambridge. The above observations 
remain almost unchanged when the analysis includes websites with up to 10 unique
postcodes -- see Figure \ref{s_map10} in the Appendix -- demonstrating the
robustness of the analysis. Given (i) the expectation that 
early adopters of new technologies are often rewarded [@rogers2010diffusion; 
@griliches1957; @ding2010modeling], and (ii) the previous findings 
illustrating the long term and sustained productivity benefit for the early 
adopters of Web in the UK [@tranosuk], the observed spatial clustering of LAD 
with fast diffusion rates can further reinforce existing and rooted patterns of 
unequal economic productivity in the UK [@mccann2022productivity].

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map}LAD Web diffusion speed, 1996-2012"}
path.image <- paste0(path, "/outputs/s/speed_map.png")
knitr::include_graphics(path.image)
```

To reveal patterns of early adoption and leapfrogging, the stability and 
volatility of the above patterns is analysed. Figure \ref{rank} plots the 
relative ranking of LAD based on website density at the beginning and the end 
of the study period. To decrease noise, the average ranking of 1996-1998 and 
2010-2012 is plotted instead of the individual years. Each line represents a 
single LAD and the colours depict the 
5\textsuperscript{th} and 95\textsuperscript{th} percentile of the absolute
difference of the LAD ranking between 1996-1998 and 2010-2012. In other words, 
the 5\textsuperscript{th} percentile includes the LAD, the relative
position of which remained unchanged between the beginning and the end of 
the study period, both at the top and at the bottom of the hierarchy (small
absolute difference in ranking). The 95th percentile includes LAD that either
increased or decreased a lot their relative positions over time (high absolute 
difference in ranking). Thus, Figure \ref{rank} reveals cases of 
*leapfrogging* since the LAD at the top-right part of the figure managed to 
jump at the top of the hierarchy at the end of the study period. 
<!-- which are often associated with latecomers benefits in the  -->
<!-- literature. -->

```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank}Dynamics of Web diffusion based on LAD rankings"}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av.png")
knitr::include_graphics(path.image)
```

To further analyse the above, Table \ref{table.late} juxtaposes the 
volatility of the LAD Web diffusion against the early conditions as it reports 
the $t_0$ estimate (inflection point) for the leapfrogging LAD identified in
Figure \ref{rank}. Indeed, more than half of the LAD that drastically increased 
their relative position at the end of the study period reached the inflection 
point $t_0$ later than the country in total and, therefore, the Web diffusion 
speed has been relatively slow for them. The fact that these LAD experienced
relatively faster diffusion rates at the end of the study period, which is 
aligned with the maturity period of the web, can lead to potential positive 
economic gains for these LAD. There is extensive literature describing the 
benefits of such technological leapfrogging and the underpinning argument is 
that latecomers can adopt and benefit from new technologies that have been 
developed elsewhere without incurring the hefty initial R&D costs 
[@teece2008firm].

Based on Figure \ref{rank} and Table \ref{table.late}, the opportunity 
to catch up that @perkins2005international illustrated at the country level, is 
also visible at the much more detailed LAD scale. Although the relevant 
literature did not pay attention to cities and regions [@yu2018sustainability], 
the above finding advocates towards further exploring leapfrogging effects at 
such a granular scale. As previously, this is consistent with including websites 
with up to 10 postcodes (see Figure \ref{rank10}).

All in all, the diffusion of the Web at the detailed scale of LAD is 
characterised by spatial heterogeneity, well-established spatial patterns and 
both stability and volatility. Although the relevant literature neglected local
scales, the above findings illustrated that such local dynamics exists and 
worth further consideration and investigation especially regarding potential 
local economic effects.

```{r late/early}
late.path <- paste0(path, "/outputs/ranks/late.csv")

read_csv(late.path) %>%
  dplyr::select(-'early/late') %>% 
  rename('$t_{0}$ estimate' = estimate,
         "Diffusion speed" = fast) %>% 
  kable(caption= "Late adopters\\label{table.late}", 
        digits=3,
        #col.names = c("LAD", "$t_{0}$ estimate", "Diffusion speed", "Early/late"),
        align = c("l", "c", "c"),
        escape = F)  
  #kableExtra::row_spec(10, hline_after = TRUE)
```

# Conclusions {#sec-conclusions}

<!-- ADD MORE ABOUT LOCAL POLICIES: E.G. USE MY RESULTS TO MODEL WHERE INFRASTRUCTURE OR GSM ANTENNAS SHOULD BE LOCALTED -->

The paper offers an economic geography study of the diffusion of a new 
technology -- the Web. Contrary to previous studies, it does so by adopting 
(i) a high level of spatial granularity, and (ii) a long study period that
captures the very early stages of this technology until its maturity.
This is the first time that the importance of granular spatial
mechanisms for the diffusion of a new technology is exposed. 
Importantly, the empirical framework and the data employed here allows
to observe the active engagement with the Web that is creating and
maintaining a commercial website instead of the more passive action of
browsing webpages and accessing the Internet.

<!-- H1 -->
Specifically, two hypotheses are tested. Firstly, whether the three diffusion 
mechanisms namely distance, urban hierarchy and the S-shaped pattern of the 
cumulative level of adoption shaped the diffusion of the commercial Web in 
the UK at granular geographical scales. The analysis firstly illustrates how 
well the S-pattern of cumulative adoption fits at the local scale and offers 
some descriptive insights about the importance of the distance and 
hierarchical diffusion mechanisms. Then, the modelling exercise reveals the 
remarkably good predictive capacity of these mechanisms at granular geographical 
scales. The modelling framework exposes that the relative importance of 
these spatial mechanisms differ across different scales. It also reveals 
spatial heterogeneity. While for some UK regions the importance of the spatial 
diffusion mechanisms are very similar to the rest of the country, other regions
diverge significantly from the country average.

To do the above, a robust and novel ML modelling framework was designed to 
predict website diffusion over space and time using as predictors the spatial 
diffusion mechanisms. Space- and time-sensitive CV was designed for the model
to avoid overfitting. Importantly, the results are robust against different 
specifications and subsets of the data. This is the first time that the power 
of these mechanisms in predicting the diffusion of a new technology was
empirically tested, importantly, at very fine scales.

<!-- H2 -->
<!-- H~2~: The diffusion of the commercial Web is shaped by both early adoption  -->
<!-- and latecomer effects. -->
The second hypothesis assessed whether the diffusion of the web is shaped by 
both early adoption and latecomer effects. On the one hand, the analysis 
highlighted the heterogeneity of adoption rates between LAD as well as the 
spatial clustering of LAD with similar adoption rates. Importantly, the 
observed spatial clustering of LAD with fast diffusion rates can
reinforce existing and rooted patterns of unequal economic productivity 
in the UK [@mccann2022productivity]. Previous research had already indicated 
the long-term productivity gains of the early adoption of such technologies 
[@tranosuk]. On the other hand, the analysis also exposed instances of 
remarkable stability as well as volatility of the Web adoption, with the 
latter being indicative of leapfrogging. While the literature had 
discussed the benefits of technological leapfrogging at the national scale 
[@teece2008firm; @perkins2005international], local scales had escaped such 
studies [@yu2018sustainability]. Thus the leapfrogging evidence presented here 
advocate towards further exploring such effects at granular scales.

The paper offers a number of different contributions. Firstly, it
assesses, for the first time, the role of these spatial diffusion
mechanisms at such granular spatial scales. While previous studies had
highlighted these mechanisms [@hagerstrand1968innovation;
@rogers2010diffusion; @grubler1990rise], their role had never been
tested before at such granular scale. The findings talk directly to this
literature [e.g. @beardsell1999spatial; @haller2011determinants;
@lengyel2020role]. More broadly, the paper offers a robust
methodological framework, which utilises novel data and modern ML
algorithms and techniques to model the diffusion of a new technology
across space and time. It directly addresses the data problem which
prevented economic geographers in engaging with diffusion studies
[@iso2005; @kemeny2011international; @zook2022mapping]. Thus, the paper
seeks to renew economic geography's interest in the diffusion of new
technologies. Current data science advances including the availability
of granular enough data and adequate methods can enable economic
geographers to reveal interesting spatial patterns of diffusion of new
technologies as this paper has demonstrated. This is important as (i)
various stakeholders are interested in predicting localised demand for
new technologies for planning purposes, for example 5G networks
[@leibowicz2016representing;@meade2021modelling], and (ii) the adoption 
of such digital technologies is associated with local economic gains 
[@solow1957technical; @aghion1990model; @kemeny2011international; @tranosuk;
@capello2024nexus]. 
<!-- In addition, while most related studies focused on -->
<!-- tangible technologies, this paper sheds light on the diffusion of a -->
<!-- truly weightless digital technology that is the Web. Thus, it adds much -->
<!-- more granular findings to the limited related studies -->
<!-- [@sinai2004geography; @haller2011determinants].  -->
Importantly, although the geographical focus of this paper is the UK, the 
empirical framework developed here is easily transferable to other geographical 
contexts.

<!-- reset the figure, table and section numbering -->

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

# Appendix {#appendix}

<!-- #sec-appendix -->

<!-- .appendix -->

\counterwithin{figure}{section}
\counterwithin{table}{section}

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_uk10}Cumulative adoption of the Web, UK; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/s/s_uk_per_firm_10.png")
knitr::include_graphics(path.image)
```

```{r echo=FALSE, message=FALSE, fig.cap="\\label{s_map10}LAD Web diffusion speed, 1996-2012; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/s/speed_map_10.png")
knitr::include_graphics(path.image)
```

```{r echo=FALSE, message=FALSE, fig.cap="\\label{rank10}Dynamics of Web diffusion based on LAD rankings; up to 10 postcodes per website"}
path.image <- paste0(path, "/outputs/ranks/web_per_firm2000_2012_only0595_av_10.png")
knitr::include_graphics(path.image)
```

```{r, morani10, echo=FALSE, message=FALSE, fig.cap="\\label{morani10}Website density Moran's I; up to 10 postcodes per website"}

path.in.la <- paste0(path, "/outputs/lisa/corrected/LA/morani_la_10.csv")
path.in.oa <- paste0(path, "/outputs/lisa/corrected/oa/morani_oa_10.csv")

la.m <- read_csv(path.in.la) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'LAD')
oa.m <- read_csv(path.in.oa) %>% 
  filter(p < 0.001) %>% # all values were significant
  mutate(spatial.unit = 'OA')

bind_rows(la.m, oa.m) %>% 
  ggplot(aes(x=year, y=morani, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Moran\'s I') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r, echo=FALSE, message=FALSE, fig.cap="\\label{lisa10}\\centering Website density LISA maps; up to 10 postcodes per website", out.height="80%"}

path.image1 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la1996.png")
path.image2 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2000.png")
path.image3 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc11_la2012.png")
path.image4 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa1996.png")
path.image5 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2000.png")
path.image6 <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc11_oa2012.png")
path.image7 <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc1_la_legend.png")

p1 <- ggdraw() + draw_image(path.image1)  
p2 <- ggdraw() + draw_image(path.image2)  
p3 <- ggdraw() + draw_image(path.image3)  
p4 <- ggdraw() + draw_image(path.image4) 
p5 <- ggdraw() + draw_image(path.image5)
p6 <- ggdraw() + draw_image(path.image6)  
p7 <- ggdraw() + draw_image(path.image7)

par(mar=rep(0,4)) # no margins

p <- p1 / plot_spacer() / p4 / p2 / plot_spacer() / p5 / p3 / plot_spacer() / p6 +
  plot_layout(ncol = 3, widths = c(1, -0.8, 1), heights = c(1,1,1))

# p + inset_element(p7, left = 0.45, bottom = -0.5, right = 0.6, top = 0.1)
p + inset_element(p7, left = 0.29, bottom = 0.8, right = 0.49, top = 1.2)

```

```{r, gini10, echo=FALSE, message=FALSE, fig.cap="\\label{gini10}Website density Gini coefficient; up to 10 postcodes per website"}

path.in.la <- paste0(path, "/outputs/gini/gini_la_10.csv")
path.in.oa <- paste0(path, "/outputs/gini/gini_oa_10.csv")

la.gini <- read_csv(path.in.la) %>% 
  mutate(spatial.unit = 'LAD')
oa.gini <- read_csv(path.in.oa) %>% 
  mutate(spatial.unit = 'OA')

bind_rows(la.gini, oa.gini) %>%
  rename(Gini = 'gini(a$n)') %>% 
  ggplot(aes(x=year, y=Gini, fill= spatial.unit)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  ylab('Gini') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous(NULL, labels = as.character(1996:2012), breaks = 1996:2012) +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
```

```{r varimp10, eval = T, echo=FALSE, message = F, warning=FALSE, fig.cap="\\label{var.imp10}Variable importance; up to 10 postcodes per website", out.height="60%"}
path.importance.lad <- paste0(path, "/outputs/rf/figures/varimp_LA_corrected_10.csv")
path.importance.oa <- paste0(path, "/outputs/rf/figures/varimp_OA_corrected_10.csv")

importance.oa <- read_csv(path.importance.oa) %>% 
  mutate(id = c("LD", "NC", "NR", "STL", "Y", "DL", "DNC", "DNR")) %>% 
  mutate(variable = str_to_sentence(variable))
importance.lad <- read_csv(path.importance.lad) %>% 
  mutate(id = c("LD", "NC", "STL", "Y", "DL", "DNC")) %>% 
  mutate(variable = str_to_sentence(variable))

importance.oa %>% 
  rename(OA = importance) %>% 
  left_join(importance.lad %>% rename(LAD = importance), by = "id") %>%
  dplyr::select(-variable.y, -id) %>% 
  rename(variable = variable.x) %>% 
  pivot_longer(!variable, names_to = "spatial.unit", values_to = "importance") %>% 
  mutate(order = ifelse(variable == "Distance to london", 1,
                        ifelse(variable == "London's website density, t-1", 2,
                               ifelse(variable == "Distance to the nearest city", 3,
                                      ifelse(variable == "Nearest city's website density, t-1", 4,
                                             ifelse(variable == "Distance to the nearest retail centre", 5,
                                                    ifelse(variable == "Nearest retail centre's website density, t-1", 6,
                                                           ifelse(variable == "Spatial and temporal lag of website density", 7, 8)))))))) %>% 
  ggplot(aes(x=importance, y=reorder(variable, desc(order)), #forcats::fct_reorder(variable, importance), 
         fill = spatial.unit)) +
  geom_col(position = 'dodge') +
  ylab('') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(vjust = 0.5, hjust=1)) +
  scale_fill_manual('Spatial unit: ', 
                    values = c("LAD" = "grey80",
                                "OA"="grey48"),
                     limits = force) +
  scale_x_continuous('Variable importance') +
  theme(legend.position="bottom",
        text = element_text(size = 7),
        legend.text=element_text(size=7),
        legend.title=element_text(size=7),
        axis.title=element_text(size=7))
```

```{r, echo=FALSE, message = F, error=FALSE}

path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- st_read(path.geo, quiet = T)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- st_transform(la, 4326)

# regions
path.regions <- paste0(path, "/data/raw/LAD21_RGN21_EN_LU.csv")
regions <- read_csv(path.regions)

s.la.path <- paste0(path, "/data/temp/s_la_per_firm.csv")

read.csv(s.la.path) %>% 
  left_join(la %>% as_tibble() %>% select(LAD21CD, LAD21NM),
            by = c("ladcd" = "LAD21CD")) %>% 
  arrange(estimate) %>%
  left_join(regions %>% select(LAD21CD, RGN21NM), by = c("ladcd" = "LAD21CD")) %>% 
  select(LAD21NM, RGN21NM, estimate, std.error, r2, fast) %>% 
  rename(LAD = LAD21NM,
         Region = RGN21NM,
         "t_0 estimate" = "estimate",
         "R-squared" = "r2",
         "Std. error" = "std.error",
         "Diffusion speed" = "fast") %>% 
  kable(caption= "S-curve estiamtes for LAD\\label{table.s.lads}", 
        digits=3,
        align = c("l", "l", "c", "c", "c", "c"),
        col.names = c("LAD", "Region", "$t_0$ estimate", "Std. error", "$R^2$", "Diffusion speed"))
```

```{r regions_no_imp}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA_corrected_10.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA_corrected_10.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(caption= "Regional differences; up to 10 postcodes per website\\label{table.regions.10}", 
        digits=3,
        col.names = c("Region", "$R^2$ LAD", "Rank LAD", "$R^2$ OA", "Rank OA")) 
```

|                   | $RMSE$ | $R^{2}$ | $MAE$ |
|-------------------|--------|---------|-------|
| Local Authorities | 0.286  | 0.506   | 0.223 |

: Model metrics for growth model {#tbl-model-growth-metrics}

@tbl-model-growth-metrics presents the results of a Random Forest model that 
was trained based on Equation \ref{model.growth} using all LAD data points. In
essence, this is a model similar to the one estimated based on Equation
\ref{eq:model} and presented in @tbl-model-metrics, with the main difference 
being the dependent variable which is the year growth rate of website 
density for each LAD. To avoid cases with LAD have no websites, 
$t \in [2000, 2012]$. The high predictive capability of the model reflects 
the robustness of the modelling approach.

\begin{align} \label{model.growth}
Website\,Density\,Growth_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}

# References {#references .unnumbered}