### 2025-03-12 paper/submission2/anonymised.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A multi-scale story of the diffusion of a new technology: the web},
  pdfkeywords={diffusion, technologies, Web, digital},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{A multi-scale story of the diffusion of a new technology: the
web}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper investigates the spatial diffusion of a new technology that
is the Web in the UK. It employs novel data and machine learning methods
to model the influence of well-established diffusion mechanisms.
Contrary to previous studies, it adopts multiple scales, high spatial
granularity and a long study period that captures the early stages of
the Web until its maturity (1996-2012). Findings reveal the importance
of such spatial mechanisms (namely distance, urban hierarchy and the
S-shaped pattern of the cumulative level of adoption) even at granular
scales. They also highlight spatial heterogeneity and instances of
leapfrogging.
\end{abstract}

\begin{keyword}
    diffusion \sep technologies \sep Web \sep 
    digital
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{sec-introduction}

Economic geographers were always interested in how new technologies and
innovations diffuse across space and time with the work of
\citet{hagerstrand1968innovation}, which demonstrated the importance of
communications as a mechanism for the diffusion of innovations, being
the most impactful. It was followed by various empirical studies -- see
for instance \citet{ormrod1990} and \citet{iso2005}. However, the torch
of exploring and modelling such diffusion processes had been passed to
other disciplines such as economics, business studies and sociology well
before the `cultural turn' of economic geography
\citep{perkins2005international, ding2010modeling}. As a consequence,
relevant studies did not investigate diffusion mechanisms at granular
geographical scales to provide local policy insights, for instance,
regarding infrastructural development and local technology policies. A
potential explanation of the lack of geographical studies exploring the
diffusion of new and, more specifically for this paper, digital
technologies across both space and time can be attributed to the
scarcity of relevant and granular enough data, a problem also
highlighted by \citet{iso2005} and \citet{kemeny2011international}. As
\citet{zook2022mapping} pointed, digital activities are hardly ever
captured in official data.

This paper offers such a contribution: an economic geography study
revealing the importance of spatial mechanisms for the diffusion of a
digital technology that is the Web in the UK at a high level of spatial
granularity from the very early days until its maturity (1996-2012). It
specifically focuses on the diffusion of commercial websites. It does so
by employing a novel source of big data and machine learning algorithms
to model the \emph{active} engagement with Web -- that is creating and
maintaining instead of just browsing a website -- during that period and
tests how different spatial and temporal mechanisms shaped its
diffusion. This paper also exemplifies how the combination of data
sources and state-of-the-art research methods which escape the
traditional social science domain can offer new lenses to geographical
research regarding the understanding of technological diffusion.

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffuse over space and time and use this knowledge to make predictions
about the diffusion of related future technologies. As per
\citet{leibowicz2016representing}, historical studies agree that
technologies diffuse differently in terms of times, rates, and
geographies and can be driven by related policies \citep{victor1993}.
\citet{meade2021modelling} highlight that a variety of actors have a
direct interest in gaining such knowledge including network equipment
suppliers, network operators, regulatory and local authorities. These
processes and their effects vary a lot across scales: although the
diffusion of a new technology might not be optimal at a local level, it
might be beneficial from a global perspective as it could lead to faster
diffusion to less advantaged places \citep{leibowicz2016representing}.
Despite the spatial heterogeneity of such diffusion mechanisms and the
policy relevance, there is a scarcity of studies analysing the diffusion
of new technologies at a detailed geographical level because of the lack
of detailed enough data and adequate research methods.

This gap becomes even more prominent if we consider the importance of
technology and its adoption in economic geography and macroeconomic
growth theories. Simply put, differences in technological adoption and
sophistication may lead to uneven development as new technologies can
increase productivity and allow firms to introduce new products
\citep{solow1957technical, aghion1990model, kemeny2011international}.
Linked to the subject of this paper, the Web and, more specifically,
commercial websites can be considered as the underpinning technology for
the transformative process identified by \citet{capello2024nexus} as the
\emph{digital service economy}. Importantly, the focus here is not the
Web in general, but instead commercial websites. After all, diffusion
together with invention and innovation are considered the pillars of
technological change \citep{das2022diffusion}. Findings from previous
studies illustrated that early engagement with digital technologies such
as the Web led to longstanding positive productivity gains for these
regions that hosted early adopters \citep{tranosuk}. Hence, obtaining a
robust understanding of the spatial diffusion mechanisms is crucial in
mapping future development trajectories.

Another line of motivation has to do with the very nature of the Web as
a digital technology and how it evolved over time. The digital
geographies field has debunked premature claims about the Internet's
placeless-ness by mapping various of its facets and technological layers
\citep{tranos2013death}, but the lack of relevant long time-series data
prevented researchers from observing digital phenomena overtime. Most of
these studies focused on its hard infrastructure
\citep[e.g.][]{malecki2002economic, moss2000internet}, on Internet
subscribers \citep[e.g.][]{blank2018local}, or on social media users
\citep[e.g.][]{crampton2013beyond}. Exceptions include early work from
\citet{zook2000web}, \citet{zook2001old} and \citet{moss1997tracking},
which utilised domain name registration data and use it to map and
illustrate the spatial footprint of the web content. Pivotal as these
contributions have been in understanding the economic geographies of the
Internet, they were based on cross-sectional, often spatially detailed,
mapping and on short temporal comparisons (e.g.~1993-1998) and did not
observe the evolution and diffusion of the Web until its maturity. As we
know from the relevant literature, different technologies diffuse at
different rates and there is no theory that can explain these
differences \citep{leibowicz2016representing}. Hence, mapping and
modelling the diffusion of the Web can offer useful insights.

This paper focuses on the diffusion of the Web as new technology during
the 1996-2012 period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
Internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer.\footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}
Also, only 9 per cent of UK's households had access to the Internet in
1998 \citep{ons2018}, the web included mostly static webpages, there
were no social media and web browsing involved exclusively a desktop PCs
as there were no smartphones \citep{tranosuk}. Hence, it is fair to say
that the study period captures the very early stages of the diffusion of
a new technology that is the Web until its maturity. The former is a key
point in the lifecycle of a new technology. Firstly, during this period
new technologies are expensive, crude and imperfect
\citep{rosenberg1994exploring, wilson201281}. A simple comparison
between Web 1.0 and Web 3.0 applications clearly illustrates this
\citep{tranos2020social}: while a static website facilitates one-way
information dissemination, a platform like \texttt{github} enables
cooperation between users and the creation of new information, meaning,
and even knowledge \citep{faraj2016special, barassi2012does}. During
this period the performance of a new technology is the main attraction
and not the cost to access and use it \citep{wilson2011lessons}. There
is a broader theoretical discussion in the literature about the early
adoption motives. As summarised by \citet{perkins2005international}, on
one hand, epidemic models highlight the role of interpersonal contacts
as a way for new technologies to diffuse. On the other hand, economic
models underline the importance of heterogeneity. Different firms have
different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories \citep{neffke2011regions, simmie2014new}. This
argument is even more relevant when the focus is on digital technologies
because of the commonly found lag between investment and economic
returns as reflected in the Solow paradox
\citep{acemoglu2014return, brynjolfsson2018artificial}.

The data used here depict the active engagement with the Web as they
contain geolocated and time-stamped counts of commercial websites from
the UK. Instead of adopting metrics reflecting the passive engagement
with the Internet such as Internet adoption and speeds
\citep[e.g.][]{blank2018local, destefano2022fuzzy}, this paper observes
yearly counts of geolocated, commercial websites, which reflect the
creation and maintenance of such websites. It needs to be highlighted
here that although there are various types of underpinning web
technologies -- namely markup, stylesheet, client- and server-side
coding languages as well as different web browsers and other related
programming language -- the focus here is on the diffusion of the Web as
an overarching technology instead of more specific web technologies. For
a such a detailed and cross-country technological diffusion study, see
\citet{PAPAGIANNIDIS2015308}.

To analyse these data, a diverse set of methods are employed. Firstly,
cumulative adoption curves are estimated for the UK as a whole and for
the Local Authority Districts (LAD), illustrating, for the first time,
such S-shaped diffusion patterns for local areas. The exploratory
spatial data analysis offers some first indications of the importance of
the different spatial diffusion mechanisms in shaping the diffusion of
the Web in the UK. Importantly, the analysis adopts two distinct
geographical scales: the LAD and the most detailed scale of UK census
areas, the Output Areas (OA). Then, a novel methodological framework
based on Machine Learning (ML) algorithms is developed to test the role
of such mechanisms at these two scales. The results reveal, for the
first time, the importance of different spatial diffusion mechanisms in
the diffusion of a digital technology at very granular geographical
scales. The analysis also reveals that the diffusion of the Web is
characterised by spatial heterogeneity, well-established spatial
patterns and both stability and volatility.

The structure of the paper goes as follows. Section~\ref{sec-litreview}
reviews the mechanisms that shape the diffusion of a new technology and
sets up the two hypotheses that are empirically tested in
Section~\ref{sec-results}. Section~\ref{sec-datamethods} describes the
data and methods used in this paper. The discussion of the results and
the conclusions of the paper are offered in
Section~\ref{sec-conclusions}.

\section{The mechanisms that shape the diffusion of a new
technology}\label{sec-litreview}

The diffusion of new technologies and innovations is driven by three
different mechanisms \citep{grubler1990rise, morrill2020spatial}: (i)
distance and proximity, (ii) urban hierarchies in the form of a
centrifugal forces as new technologies diffuse from core to periphery,
and (iii) over time the cumulative level of adoption follows an S-shaped
pattern. \citet{hagerstrand1968innovation} was the first to identify
diffusion as a geographical process. The starting point was the idea
that diffusion is based on passing information through social networks,
which themselves tend to be geographically bounded. Hence, he identified
a `neighbourhood' mechanism of how information, and consequently,
innovations and new technologies diffuse. This perspective draws
similarities with epidemics: new technologies and innovations, just like
pathogens, spread because of contagion and, consequently, proximity and
exposure \citep{hivner2003facilitating}.

\citet{hagerstrand1968innovation} also emphasised the role of the
top-down hierarchy of urban systems and highlighted that new
technologies and innovations are firstly adopted in larger cities and
then diffuse to second tier ones. This is a sequential instead of a
simultaneous process, which resembles the `lead-lag' spatial
acceleration effect in market research
\citep{bento2018time, PERES201091}. Schmidt's Law empirically
illustrates a similar pattern \citep{grubler1990rise}. Core and highly
agglomerated regions is where new technologies are invented and
commercially deployed. This is where the first adopters tend to be
based. Then, technologies spread to the rim and eventually to the
periphery. Although adoption pace might be higher when new technologies
finally arrive to the periphery, the saturation levels may never reach
the ones in the core because of the lack of infrastructure or other
necessary institutions \citep{leibowicz2016representing}.

Diffusion is naturally a dynamic process and Hägerstrand recognised the
role time plays in the diffusion of innovations and new technologies: an
early-pioneering period, a middle fast accelerating period and a final
saturation one \citep{morrill2020spatial}. This was further explored by
\citet{rogers2010diffusion}, who identified five types of adopters of
new technologies and innovations: innovators, early adopters, early
majority, late majority and laggards. Hence, plotting the cumulative
adoption of a new technology over time often leads to an S-shaped line
plot \citep{grubler1990rise}.

Although the above processes have been studied in innovation studies,
the distancing of economic geographers from research questions about the
diffusion of new technologies and innovations resulted to a knowledge
gap regarding the explanatory power of these mechanisms at granular,
local scales \citep{ding2010modeling}. As discussed in
Section~\ref{sec-introduction}, such spatially explicit knowledge can
directly inform local technology and infrastructure policies. Indeed,
while the spatial dimension is present in the following studies, the
geographical level of analysis is often too coarse to ascertain the role
of the above mechanisms at the local scale. \citet{beardsell1999spatial}
studied the evolution of the computer industry in 317 US metropolitan
areas during the 1977-1992 period using employment data. Their analysis
indicated that the relative size distribution holds for urban computer
employment and also urban heterogeneity is essential in explaining this
distribution. In a recent study, \citet{bednarz2020pulled} focused on
wind turbines and modelled their spatial diffusion across 402 German
regions during 1970-2015. Their key finding is that local demand than
local supply was the main driving factor. \citet{haller2011determinants}
employed firm panel data for Ireland during 2003-2005 and illustrated
that when firms are located in a region or industry with high share of
firms having a website, they are more likely to have a website too. At a
global scale \citet{perkins2011internet} explored whether the adoption
of previous communication technologies that is mail, telegrams and
telephones were shaped by similar socioeconomic factors as the Internet.
Their results indicated common patterns regarding the drivers behind the
adoption of different communication technologies.
\citet{feldman2015rknowledge} focused on inventions and used patent data
to map how the the rDNA technology diffused among inventors across 366
US metropolitan areas. \citet{ding2010modeling} modelled the spatial
diffusion of mobile telecommunications across the highest level of
administrative division in China (29 regions in total). Their analysis
indicated that socioeconomic characteristics are important determinants
of the timing, speed and the level of mobile diffusion within China.
\citet{bakher2013diffusion} compared different growth curves as a means
to model the diffusion of online shopping in Australia during the
1998-2009 period, but they did not consider any spatial mechanisms. At a
global scale \citet{PAPAGIANNIDIS2015308} modelled the diffusion of
different web technologies and practices across countries. Using similar
but less extensive data as the one used here, their analysis illustrated
how the diffusion of different web technologies and practices follows an
S-shaped pattern as well as the different diffusion rates of the
different technologies and practices. The most detailed consideration of
the spatial footprint of the diffusion of a new technology comes from
\citet{lengyel2020role}, who analysed the adoption and churn of a
Hungarian online social network. Their results were in agreement with
the early theoretical and empirical contributions reviewed here:
assortativity, urban scaling and distance are the key drivers of spatial
diffusion. Based on the above, the main empirical hypothesis that is
tested in this paper states:

\begin{quote}
H\textsubscript{1}: The three diffusion mechanisms namely distance,
urban hierarchy and the S-shaped pattern of the cumulative level of
adoption shaped the diffusion of the commercial Web in the UK even at
granular geographical scales.
\end{quote}

Understanding the diffusion of a new technology in multiple, detailed
spatial scales is of interest to economic geography as potential local
economic benefits may occur because of concentration of early adopters
or through technological leapfrogging. On the one hand, early adopters
are often rewarded because of their attitudes towards new technologies
and innovations. \citet{rogers2010diffusion} identifies them as
`knowledgeable risk takers' and \citet{griliches1957} as `profit
maximisers' \citep{ding2010modeling}. On the other hand, latecomers can
adopt and benefit from new technologies that have been developed
elsewhere without incurring the hefty initial R\&D costs
\citep{teece2008firm}. At a global scale
\citet{perkins2005international} explored whether the diffusion rate of
new technologies is driven by a latecomer advantage and the engagement
with the global economy via foreign direct investments and trade. Their
results illustrate that indeed latecomers and developing countries
experience diffusion of new technologies more rapidly than early
adopters and developed countries. At a regional scale, previous research
highlighted the long term and sustained productivity benefit of the
early adopters of web technologies in the UK \citep{tranosuk}. Although
the early adoption and leapfrogging literature has not paid much
attention on cities and regions \citep{yu2018sustainability}, the
potential local economic benefits justify the need to investigate the
dynamics of the web diffusion in different scales in the UK. Hence, the
second empirical hypothesis states:

\begin{quote}
H\textsubscript{2}: The diffusion of the commercial Web is shaped by
both early adoption and latecomer effects.
\end{quote}

These two hypotheses are empirically tested in Section~\ref{sec-results}
employing the data and methods described in the next section.

\section{Data and Methods}\label{sec-datamethods}

To capture the diffusion of the commercial Web, a website density metric
is developed for two different geographical scales that are discussed
later in this section. The main data are counts of commercial websites
that are calculated using data from the Internet Archive\footnote{\url{https://archive.org/}.}
and, specifically, the JISC UK Web Domain Dataset \citep{ukwebarchive}.
The Internet Archive is one of the most complete and oldest archive of
webpages in the world operating since 1996
\citep{ainsworth2011much, holzmann2016dawn}. It is a web crawler, which
discovers webpages by following the hyperlinks of every webpage it
archives. This dataset, which is curated by the British Library,
contains all the archived webpages from the UK country code top level
domain (ccTLD - .uk) from the 1996--2012 period. In essence, this is a
long list of 2.5 billion URLs of archived webpages including also the
archival timestamp.

Instead of using the whole .uk ccTLD, this paper focuses on its
commercial subset, the .co.uk second level domain (SLD). This choice is
aligned with the topic of the paper -- the diffusion of the Web for
commercial functions -- and decreases the heterogeneity of the web data
as such commercial websites have specific aims: they are used to diffuse
information, support online transactions and share opinions
\citep{THELWALL2000441, blazquez2018big}. Although a UK company can
adopt a generic TLD such as .com and these cases escape the data used
here, such omissions should not affect the validity of the results given
the popularity of the .uk ccTLD \citep{tranosuk}: UK consumers prefer to
visit a .uk website when they are searching for products or services
\citep{hope}; and anecdotal evidence indicates that during the first
half of 2000, three .co.uk domains were registered every minute
\citep{oecd_coms}. Importantly, previous studies illustrated that .co.uk
is the most popular UK SLD \citep{tranosuk}. Henceforward any reference
to websites will refer to commercial -- .co.uk -- websites.

The text from these webpages was scanned using a regular expression
(regex) to identify strings of text which resemble UK postcodes and one
fifth of them included a mention to a postcode \citep{BL2013geo}. This
information allows the geolocation of the data and the creation of the
LAD and OA counts. To prevent false positives, the resulted postcodes
were matched against official postcode lookup files \citep{ons}.

The data cleaning process includes an aggregation step, through which
the archived webpages are aggregated to the parent websites. This
website reconstruction process lend itself to the construction of
\emph{website} instead of \emph{webpage} density metrics. Websites tend
to represent specific organisations or entities, and, arguably, are more
meaningful observational units than webpages, which can ignore the
upstream dependency of the website they belong to. Based on the
following example, all three webpages are part of the same website
(\url{http://www.website.co.uk}). Only the overall website and not the
nested webpages are considered for the counts as otherwise the density
metrics would have been biased towards large, place specific, websites.
It is important to note here that websites commonly contain place
references, with previous research estimating this share to be around
70\% \citep{hill2009georeferencing}.

\begin{itemize}
\item
  \url{http://www.website.co.uk/webpage_a} B15 2TT
\item
  \url{http://www.website.co.uk/webpage_b} BS8 1TH
\item
  \url{http://www.website.co.uk/webpage_c} B15 2TT
\end{itemize}

\noindent What is challenging is that this aggregation approach, which
has been used elsewhere \citep{tranosuk, shoreditch} may lead to
websites with multiple postcodes. As per the above example,
\href{http://www.website.co.uk}{www.website.co.uk} includes two unique
postcodes: B15 2TT and BS8 1SS. The distribution of the number of
postcodes per website for 2000 is presented in Table \ref{f2000}, which
illustrates a wide range. At the left end of the distribution, there are
websites anchored to a unique location (72\% of all the reconstructed
websites in 2000), which may represent a small company with a single
trading location. At the right end, there are websites with thousands of
different postcodes. Considering the time period of the analysis, such
cases can represent directories which used to be popular in the
pre-search engines early times of the commercial Internet as well as
real estate websites \citep{tranosuk}. Previous research which used the
same methodology for a small subset of these data illustrates the
robustness of the above approach: its results matched in terms of
accuracy past rounds of research based on extensive qualitative work,
interviews and observational studies for the same study area
\citep{shoreditch}.

The analysis is firstly conducted on websites, which only contain one
unique postcode. As a robustness check, the analysis is replicated for
an extended subset of websites, which include up to 10 unique postcodes
to capture commercial websites which point to multiple locations. They
are geolocated by equally attaching them to these locations. This
extended sample includes 94 percent of all the archived websites in 2000
(see Table \ref{f2000}).

\begin{table}

\caption{Number of unique postcodes per .co.uk website, 2000.\label{f2000}}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Postcodes & F & F (\%) & cumulative F\\
\midrule
(0,1] & 41,596 & 0.718 & 0.718\\
(1,2] & 6,451 & 0.111 & 0.830\\
(2,10] & 6,163 & 0.106 & 0.936\\
(10,100] & 2,975 & 0.051 & 0.988\\
(100,1000] & 646 & 0.011 & 0.999\\
\addlinespace
(1000,10000] & 62 & 0.001 & 1.000\\
(10000,100000] & 4 & 0.000 & 1.000\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Source: } Tranos et al. 2021}\\
\end{tabular}
\end{table}

Another data cleaning step deals with some extreme outliers. Figure
\ref{correct} plots the website counts for the top 1000 postcodes. Some
obvious outliers can be observed for the 2002-2006 period for a handful
of postcodes, which can be attributed to link farms \citep{BL2013links}.
The website counts for these postcodes (SE24 9HP, CV8 2ED, GL16 7YA, CW1
6GL, M28 2SL, DE21 7BF), which in 2004 or 2005 had more than 1,000
websites pointing to them, were replaced with predicted values based on
a simple panel regression model with postcode fixed effects and yearly
dummy variables. These postcodes refer to a residential area with a
small park in South London, residential areas outside Warwick and
Manchester, a rural area in the south border of England and Wales and
small business parks in Crew and Derby. Hence, it is difficult to
believe that these postcodes genuinely hosted such a large amounts of
websites, which were later extensively declined. To put the magnitude of
the data imputation into perspective, this process affected 6 out of the
557,808 postcodes included in the data. As Section~\ref{sec-results}
illustrates, this led to a small increase of the model predictive power.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{anonymised_files/figure-pdf/unnamed-chunk-2-1.pdf}

}

\caption{\label{correct}Yearly website counts per postcode (top 1000)
before (top) and after (down) data imputation}

\end{figure}%

To empirically test the two hypotheses stated in
Section~\ref{sec-litreview}, the website density metric was calculated
for two different scales the Local Authority Districts (LAD) and the
Output Areas (OA). The former is the lowest administrative unit in the
UK and there are 374 LAD. The latter is the smallest census-based
geographical unit and there are c.~230,000 of them in the UK. To create
the website density metric, the yearly website counts at the LAD level
are standardised by the number of firms in LAD to avoid biases
associated with LAD hosting a large number of firms. Given that there is
no such readily available statistic for the OA, the actual OA level
counts are used. However, because of the the consistent spatial
definition of OA (they host 40-250 households),\footnote{\url{https://www.ons.gov.uk/methodology/geography/ukgeographies/statisticalgeographies}}
website counts in OA are interpreted as a density metric too.

These website density metrics are used for the different stages of the
analysis. Starting with the first hypothesis (H\textsubscript{1}), to
test whether the S-curve accurately describes the cumulative adoption of
the Web at a local scale, the following logistic function (Equation
\ref{eq:s}) is estimated for the whole of the UK and then for each LAD
separately:

\begin{align}
y = k /(1 + e^{-(t-t_{o})})\label{eq:s}
\end{align}

\noindent \(k\) is the asymptote or, in other words, the saturation
level, \(b\) the overall growth rate, and \(t_{0}\) the \emph{inflection
point} of maximum growth at \(k/2\), where the logistic function is
symmetrical \citep{wilson201281}. The \(t_0\) of each LAD is compared
against the \(t_0\) of the UK to delineate whether a LAD reached that
point faster or slower than the country average. This heterogeneity is
discussed further when testing the second hypothesis
(H\textsubscript{2}). Importantly, an accuracy criterion was imposed and
only LAD with \(R^2 > 0.9\) were included in this analysis. To estimate
Equation \ref{eq:s} a self-starting logistic growth model was employed
using the \texttt{nls} and \texttt{SSlogis} functions in \texttt{R}.
This logistic growth model depicts the initial period of slow diffusion
of a new technology, which is followed by the fast, exponential growth
stage to then eventually slow down and saturate
\citep{wilson201281, grubler1999dynamics}.

Next, exploratory spatial data analysis provides some first insights of
whether the other two drivers of spatial diffusion -- namely
neighbourhood and hierarchy -- underpin the diffusion of the Web in the
UK. To capture the former and following \citet{ding2010modeling} the
Moran's I and the Local Index of Spatial Association (LISA) are
estimated for the website density. To capture the hierarchy mechanism,
the Gini coefficient -- an established metric of inequality -- is
calculated. All the above are computed and plotted longitudinally both
for the LAD and the OA.

To formally test H\textsubscript{1} a modelling framework is developed
to assess the above diffusion mechanisms. The overarching aim is to
build a model that can test how well these three mechanisms can predict
website density:

\begin{align}
Website\,Density_{t} \sim Hierarchy_{t-1} + Neighbourhood_{t-1} + S\text-curve_{t}\label{eq:rf.generic}
\end{align}

To capture the hierarchical mechanism the model include as predictors a
one year lag of website density in London, the largest city in the UK,
and a one year lag of the website density in the nearest city. Both
variables clearly illustrate the urban hierarchy in the UK. Especially
for the OA models, a third predictor is added to capture hierarchy at a
much more granular scale: a one year lag of the website density in the
the nearest retail centre \footnote{The retail centres data have been
  obtained by the Consumer Data Research Centre, an ESRC data
  investment, under project ID CDRC 498-01, ES/L011840/1; ES/L011891/1:
  \url{https://data.cdrc.ac.uk/dataset/retail-centre-boundaries-and-open-indicators}.}.
As the literature illustrates, retail centres are key components of
urban systems \citep{dennis2002central, jones2021reframing}. Due to the
small sizes of the retail centres, the latter is only relevant for and
included in the OA models and not in the LAD models. In addition, the
model includes the distance to London, the nearest city and the nearest
retail centre. The underlying logic is that the level of website
adoption in a spatial unit depends on the level of the adoption in
places further up in the urban hierarchy the previous year. The
inclusion of the distance variables incorporates spatial structure into
the hierarchy argument. To depict the neighbourhood mechanism, the
website density of the neighbouring spatial units in the previous year
is employed. This is calculated using rook continuity. Again, the
underpinning rationale is that the level of web adoption within a
spatial unit depends on the level of web adoption in the neighbouring
spatial units the year before. This represents the `hitting nearby
locations first' argument. Therefore, the spatial and temporal lag of
the website density in LAD and OA is calculated. Lastly, the temporal
pattern of the cumulative adoption is captured by a time trend variable.
Hence, the model will follow the following generic form (Eq.
\ref{eq:model}):

\begin{align} \label{eq:model}
Website\,Density_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}

Specifically, four different models are estimated. Firstly, all the data
points for the OA and LAD are utilised in order to build two models and
assess their capacity to predict the adoption of the Web at the two
different scales. These two models will reveal the predictive capacity
of the diffusion mechanisms and also demonstrate how the importance of
such variables changes across different scales. The next two sets of
models are again trained at these two scales: OA and LAD. However,
instead of using all the data points, the OA and the LAD from one of the
twelve UK regions are held out for the model training. Then, the trained
model is used to predict website density for the OA or the LAD of the
held-out region. This process takes place recursively for all twelve UK
regions. The regional differences in the predictive capacity of the
different samples will reveal how dissimilar these spatial process are
across regions and, importantly, at different scales.

Equation \ref{eq:model} is estimated using Random Forest (RF). This is a
popular ML algorithm for both regression and classification problems
\citep{biau2012analysis}. It was introduced by \citet{breiman2001random}
and has become a go-to data science tool. RF can effectively handle
skewed distributions and outliers, model non-linear relationships,
require minimal hyperparameter tuning, exhibit low sensitivity to these
parameters, and have relatively short training times
\citep{Caruana2008, liaw2002classification, yan2020using}. These
attributes match well with the website density data characteristics
including skewness especially for the OA. Also, the large data size (c.
230k data points for each of the 17 years) calls for fast training
times. Importantly, RF predictions tend to be more accurate than those
from single regression trees and outperform Ordinary Least Squares in
out-of-sample predictions, even with moderate-sized training data and a
small number of predictors
\citep{mullainathan2017machine, athey2019machine, sulaiman2011intelligent, pourebrahim2019trip, biau2012analysis}.

RF is a tree-based ensemble learning algorithm
\citep{breiman2001random}. It begins by generating random samples of the
training data, which are then used to grow regression trees to predict
the dependent variable. Data points and predictors are randomly sampled
for the different trees. The trees are trained in parallel using their
own bootstrapped samples of the training data. A crucial feature of RF
is their ability not to overfit, meaning they can generalize well to
unseen test data. While each tree may overfit individually, the ensemble
of trees does not because the errors of individual trees are averaged,
reducing the overall variance and preventing overfitting
\citep{last2002improving}. For regression problems, RF predictions are
made by averaging the predictions of all decision trees.

RF have been widely employed to address regression research problems.
\citet{pourebrahim2019trip} combined a spatial interaction modelling
framework with ML algorithms including RF to predict commuting flows in
New York City. \citet{sinha2019assessing} advocated for adopting spatial
ensemble learning approaches, such as RF, to model spatial data with
high autocorrelation and heterogeneity. \citet{creditspatial} predicted
employment density in Los Angeles using spatially explicit RF.
\citet{guns2014recommending} used RF to build a recommendation system
for research collaborations. \citet{ren2019predicting} trained RF to
predict the socio-economic status of cities using various online and
mobility predictors. \citet{tranos2023using} utilised hyperlinks data
and RF to make out-of-sample predictions of interregional trade.
\citet{zhou2023geography} employed such a framework to assess whether
key predictors of obesity differ across English cities.

It needs to be highlighted here that the cross validation (CV) for all
models is spatially and temporally sensitive. Instead of using 10 random
and space- and time-agnostic folds, the \texttt{CAST} package is
employed which allows holding back data points from specific years and
spatial units (regions) and use them for testing in order to estimate
the model performance \citep{meyer2018improving}.

To assess the predictive power of the model, three broadly utilised
metrics are employed: the coefficient of determination (\(R^2\)), mean
absolute error (MAE) and root mean square error (RMSE):

\begin{align}
R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
\end{align} \begin{align}
MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
\end{align} \begin{align}
RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
\end{align}

\noindent \(y_{k}\) is the \(k^{th}\) observation of the dataset, which
consists of \(N\) observations in total. \(\hat{y_{k}}\) is the
\(k_{th}\) predicted value for the dependent variable and
\(\overline{y_{k}}\) is the average value of \(y\). The last two metrics
are expressed in the same units as the dependent variable -- websites
per firm for the LAD modes and the number of websites for the OA models
-- while the first one is the coefficient of determination between the
observed and the predicted values of website adoption. Regarding
\(MAE\), it is the absolute difference between the observed and the
predicted website adoption. While \(MAE\) does not penalise for large
errors, \(RSME\) does, as it is proportional to the squared difference
between the observed and the predicted trade flows. Hence larger errors
weigh more for \(RMSE\) \citep{pontius2008components}.

An alternative specification could model the growth of website density.
However, such a strategy is problematic because of the large frequency
of OA without websites, which is to be expected given their small size
and high granularity. Still, as a robustness text, the results of a
growth model for LAD are presented in the Appendix.

Lastly, to test the second hypothesis (H\textsubscript{2}) about the
early adoption and the latecomer effects, a different research framework
is developed. Firstly, the S-curve and the estimated \(t_0\) for every
LAD are employed and mapped in order to depict the spatial heterogeneity
and specific patterns of the speed of the diffusion of the Web. Then,
the volatility and the stability of the diffusion across the different
LAD is analysed using tools from the complex systems domain
\citep{Batty2009} namely changes in rankings of LAD based on website
density. Both the S-curve estimation and the volatility analysis focus
only on LAD as the very large number of OA would have made such analysis
difficult to visualise and interpret.

\section{Results}\label{sec-results}

\subsection{The three diffusion mechanisms at granular geographical
scales}\label{the-three-diffusion-mechanisms-at-granular-geographical-scales}

Figure \ref{s_uk} plots the cumulative adoption of the Web in the UK as
a whole during the 1996-2012 period. In line with the theoretical
expectations discussed in the previous section, the cumulative adoption
of the Web is well represented by a curve that resembles the letter S.
The vertical line in the beginning of 2003 (2003.158) illustrates the
point when the modelled cumulative adoption was equal to 50\% of the
maximum. This \(t_0\) point is known as the \emph{inflection} point and
signals the maximum adoption speed. The results remain unchanged when
the extended subset of websites, which include up to 10 unique postcodes
is used (see Figure \ref{s_uk10} in the Appendix). Importantly though,
the extended subset of websites indicates a saturation level close to
100\% at the end of the study period.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm.png}

}

\caption{\label{s_uk}Cumulative adoption of the Web, UK}

\end{figure}%

What escaped from previous studies was εexploring whether such a pattern
also describes cumulative adoption of new technologies at sub-national
scales. To test this, the same logistic function is estimated for each
of the 374 LAD. The detailed results are presented in Table
\ref{table.s.lads} in the Appendix. The key finding is that for 342 LAD
(91\% of all LAD) the \(R^2 > 0.9\) for the logistic function, which
indicates that even at local scale the cumulative adoption of the Web is
characterised by a pattern similar to the one observed for countries.
Graphically, all these LAD curves are plotted in Figure \ref{s_lad_uk}.
The different colour lines illustrate whether a LAD reached \(t_0\)
faster or slower than the UK in total. This heterogeneity is further
discussed later in this section.

This is the first time that such cumulative adoption curves have been
estimated for areal units as small as the UK LAD. Understanding and
modelling the dynamics of technological diffusion and, ultimately, being
able to predict diffusion rates at such small local scales can directly
support the deployment of infrastructural networks (e.g.~5G) and local
technology policies. The above are in line with previous studies
exploring the diffusion of web technologies
\citep{PAPAGIANNIDIS2015308}, social media \citep{lengyel2020role} and
online shopping \citep{bakher2013diffusion}.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/lad_per_firm_all.png}

}

\caption{\label{s_lad_uk}Cumulative adoption of the Web, LAD}

\end{figure}%

To assess the second diffusion mechanisms, Figures \ref{morani} and
\ref{lisa} offer a first insight into whether a neighbourhood or, in
other words, a distance mechanism underpins the diffusion of the Web in
the UK as they present the Moran's I and the LISA maps of website
density respectively for LAD and OA. Starting from the former, spatial
autocorrelation was higher in the beginning of the study period and then
after 2000 dropped slightly and stabilised around 0.2. This reflects the
early concentration of high website density around London, which over
time diffused as high website density clusters can be seen in other
parts of the country away from London (Figure \ref{lisa}). An almost
reverse pattern can be observed for OA. At the beginning of the study
period Moran's I was around 0.5 and it plateaued after 2000 around 0.2.
Because of the very small size of OA, at the early stages of the
diffusion of web technologies their adoption was spatially scattered.
This is reflected in the lack of any significant clusters in 1996
(Figure \ref{lisa}). Eventually, as the adoption rate increased, more
such clusters of high website density were formed and this is reflected
both in the Moran's I and the LISA maps in Figures \ref{morani} and
\ref{lisa}. A similar pattern is observed for the extended dataset (see
Figures\ref{morani10} and \ref{lisa10}). In this case, the magnitude of
spatial autocorrelation is slightly higher illustrating a stronger
spatial clustering mechanism. All in all, the exploratory spatial data
analysis advocates towards an underpinning neighbourhood mechanism,
which is present at different scales. Interestingly, the magnitude of
Moran's I is similar to the ones revealed by \citet{ding2010modeling}
about mobile phone adoption in China albeit the much more detailed
spatial scale of this paper.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani-1.pdf}

}

\caption{\label{morani}Website density Moran's I}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-5-1.pdf}

}

\caption{\label{lisa}\centering Website density LISA maps}

\end{figure}%

To illustrate whether a hierarchical process also underpins the
diffusion of web technologies -- the third diffusion mechanism that is
tested here -- the Gini coefficient is calculated yearly both for LAD
and OA. As a metric of inequality, the Gini coefficient demonstrates
whether website density is concentrated in a small number of LAD or OA,
or whether it is more equally spread across the country. Both scales of
analysis in Figure \ref{gini} illustrate the same picture. At the
beginning of the commercial Internet, website density was extremely
unequal, or, in other words, only a few places had websites associated
with them. Inequality dropped and plateaued after 2000 for both scales.
This is illustrative of a hierarchical diffusion mechanism that led over
time to a more equal spread. Interestingly, the year 2000 is again a
period of change for this diffusion mechanism as it was for the
neighbourhood process. There is a substantial difference between the
Gini coefficient magnitude for LAD and OA, but this is expected as the
very small size of OA equates to a lot of polygons without any websites
pointing to them -- for example residential OA.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini-1.pdf}

}

\caption{\label{gini}Website density Gini coefficient}

\end{figure}%

After discussing the descriptive findings, the validity of the above
diffusion mechanisms at the local scale is assessed with a predictive
modelling framework described in Section~\ref{sec-datamethods}. The
reader is reminded that four models are estimated: for the OA and LAD,
firstly all data points are employed to assess the capacity of the three
mechanisms to to predict the adoption of the Web at these two scales;
then, the LAD/OA from one of the twelve regions are recursively held out
to assess the spatial heterogeneity of these mechanisms across the
different UK regions.

Table~\ref{tbl-model-metrics} presents the model performance for the
first set of models, for which all data points are employed for training
and testing via CV. The first one is trained and tested on 374 LAD and
the second on 232,296 OA, both for the 16 year period (1997-2012). The
results are remarkably good considering that the are the outcome of
space and time sensitive CV, so the the model does not suffer from
overfitting. At the LAD level the model predicts 83\% of the variation
of website density. Both error metrics indicate that the model error is
fraction of a website per firm (0.020 and 0.029). At the OA the \(R^2\)
drops down to 33\%. Considering its granularity, this is still a
remarkable performance. To contextualise it, the model results in a
\(MAE\) equal to 1.2 websites for areas small enough to host 40-250
households. Because of the small size of the spatial units, the
distribution is highly skewed and a significant part of them is not
linked to any websites. In 1997 only 1\% of the UK OA were associated
with at least one website. This should not come as a surprise given that
this was the very beginning of the commercial Internet and any
activities with a digital footprint were concentrated in a handful of
areas, as illustrated in Figure \ref{lisa}. At the end of the study
period almost half of the UK OA were not associated with a website.
Again, given the granularity of the data this was expected.

\begin{longtable}[]{@{}llll@{}}
\caption{Model metrics}\label{tbl-model-metrics}\tabularnewline
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Local Authorities & 0.029 & 0.824 & 0.020 \\
Output Areas & 3.218 & 0.333 & 1.233 \\
\end{longtable}

The results in Table~\ref{tbl-model-metrics} are robust against
different specifications. As noted earlier, the data imputation process
resulted in a modest improvement of the predictive capacity of the
model. As per Table~\ref{tbl-model-metrics-no-correction}, all accuracy
metrics are only slightly worst when the data imputation is not applied.
As expected, the improvement is higher for the more detailed scale of
analysis. Then, Table~\ref{tbl-model-metrics-10} presents the model
results for the extended dataset, which includes websites with up to 10
unique postcodes. While the results for LAD are only slightly less
accurate, the \(R^{2}\) of the OA model is much higher in comparison to
the base results in Table~\ref{tbl-model-metrics}. In addition,
Table~\ref{tbl-model-growth-metrics} in the Appendix presents the
accuracy results of a LAD model, the dependent variable of which is the
yearly growth rate of website density. Even for such a specification
(see Equation \ref{model.growth} in the Appendix) the predictive
capacity of the model is exceptionally high (\(R^2 = 0.5\)). All in all,
the modelling exercise revealed how well the three diffusion mechanisms
predict the diffusion of the Web as a new technology. Importantly, this
predictive power is highly robust against against different scales and
delineations of the diffusion of the web.

\begin{longtable}[]{@{}llll@{}}
\caption{Model metrics without data
imputation}\label{tbl-model-metrics-no-correction}\tabularnewline
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Local Authorities & 0.032 & 0.781 & 0.021 \\
Output Areas & 4.644 & 0.244 & 1.237 \\
\end{longtable}

\begin{longtable}[]{@{}llll@{}}
\caption{Model metrics for the extended
dataset}\label{tbl-model-metrics-10}\tabularnewline
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Local Authorities & 0.190 & 0.749 & 0.129 \\
Output Areas & 21.591 & 0.455 & 6.845 \\
\end{longtable}

Figure \ref{var.imp} plots the importance of the different predictors.
When the focus is on LAD, the website density in the neighbouring LAD,
in the nearest city and in London the year before are the most important
predictors. They are followed by the yearly trend, while the spatial
configuration as reflected in distances to London or to the nearest city
only plays a minor role. This can be attributed to the relatively coarse
spatial scale of LAD. Nevertheless, all previously discussed spatial
processes are at play in the diffusion of the Web at the LAD level: the
first two predictors depict the hierarchical mechanism, the spatial and
temporal lag of website density the neighbourhood mechanism, and the
yearly trend the time-sensitive cumulative adoption pattern.

At the much more granular scale of OA, the picture is almost reversed.
The two most important predictors are the distance to London and to the
nearest city followed by the spatial and temporal lag of website density
and the distance to the nearest retail centre. They still depict a
hierarchical mechanism, but proximity to the different population
centres is more important than their lagged web densities in predicting
website diffusion. The neighbourhood mechanism is still strong but less
important at this scale. What is interesting is the almost negligible
role of the yearly trend and London's website density. While the former
probably illustrates the large heterogeneity in how the Web has been
adopted at this very fine scale, the latter highlights that the
importance of past web adoption rates in large population centres is
surpassed by proximity to them and spatial configuration at this scale.

The above pattern of variable importance is largely in accordance with
the models estimated for the extended dataset. A comparison with Figure
\ref{var.imp10} illustrates that the main difference is the importance
of the neighbourhood mechanism for the OA as the space and time lag of
the website density is by far the most important predictor when the
extended dataset is employed. For the LAD, the importance of the
distance to the nearest city increases in favour to its website density
the previous year. But still, the variable importance patterns for the
two datasets are not dissimilar.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp-1.pdf}

}

\caption{\label{var.imp}Variable importance}

\end{figure}%

Table \ref{table.regions} presents the results of the recursive hold out
models and allows to (i) highlight the regional heterogeneity of the
spatial processes behind the diffusion of the Web, and (ii) compare this
heterogeneity across the two different scales of analysis. It
illustrates both regional similarities and differences. At the top of
Table \ref{table.regions} the regions of South East and the West
Midlands have the highest similarity with the rest of the country. In
other words, the web density for the LAD of these regions can be very
well predicted by models trained on data from the LAD from the rest of
the country and, therefore, the spatial diffusion mechanisms for these
regions represent quite well the UK's average. This should not come as a
surprise as the spatial structure of both regions resembles the broader
UK: both regions contain some highly urbanised LAD (South East is well
integrated with the Greater London and West Midlands is the home of the
city of Birmingham), but also smaller commuter towns as well as more
rural settlements. At the other end of the spectrum, the spatial
diffusion mechanisms for Northern Ireland, Scotland and North West
diverge the most from the country average. Rurality and remoteness seem
to be common attributes for these regions. This hierarchy though is not
consistent across scales. For the more granular OA models, the spatial
diffusion mechanisms for South West, Yorkshire and The Humber and North
West regions are closer to the country average. Interestingly, these
regions are well-known tourist destinations (Lake District, Peak
District and Cornwall-Devon) and the exploratory analysis revealed
clusters of high web density only visible at the OA scale (Figure
\ref{lisa}). Then, Scotland, Northern Ireland and London have the lowest
similarity with the country average. Again, rurality, peripherality, but
also the highest level or urbanisation make the spatial diffusion
mechanisms of these regions more distinct.

\begin{longtable}[]{@{}lrrrr@{}}
\caption{Regional differences\label{table.regions}}\tabularnewline
\toprule\noalign{}
Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
South East & 0.952 & 1 & 0.122 & 6 \\
West Midlands & 0.926 & 2 & 0.126 & 4 \\
Wales & 0.909 & 3 & 0.122 & 8 \\
Yorkshire and The Humber & 0.908 & 4 & 0.160 & 2 \\
East Midlands & 0.904 & 5 & 0.122 & 7 \\
South West & 0.894 & 6 & 0.185 & 1 \\
North East & 0.887 & 7 & 0.124 & 5 \\
London & 0.880 & 8 & 0.090 & 11 \\
East of England & 0.876 & 9 & 0.106 & 9 \\
North West & 0.837 & 10 & 0.142 & 3 \\
Scotland & 0.774 & 11 & 0.059 & 12 \\
Nortern Ireland & 0.590 & 12 & 0.101 & 10 \\
\end{longtable}

Comparing the above with Table \ref{table.regions.10} in the Appendix,
which presents the results for the the extended dataset with websites
with up to 10 unique postcodes, the results do not differentiate much,
signalling the robustness of the analysis. The most salient difference
has to do with Northern Ireland. According to the extended sample at the
granular OA level, Northern Ireland's spatial diffusion mechanisms are
closer to the UK's average than the results based on the websites with a
unique postcode. While the latter are websites and, consequently,
economic activities solely based in Northern Ireland, websites with
multiple postcodes also include websites with postcodes in Britain and,
therefore, can be characterised by slightly different spatial diffusion
processes.

\subsection{Early adoption and
latecomers}\label{early-adoption-and-latecomers}

To test the second hypothesis (H\textsubscript{2}), the spatial
heterogeneity of Web adoption speeds is firstly analysed. As highlighted
in the literature \citep{risk_perceptions}, different agents have
different perceptions about and levels of acceptance of the risks and
the potential economic returns associated with the adoption of new
technologies -- see for instance the seminal work of
\citet{venkatesh2000theoretical}. To reveal such patterns, Figure
\ref{s_map} maps whether LAD reached their inflection point earlier or
later than the UK in total and based on this are labelled as \emph{fast}
or \emph{slow} accordingly. Inflection points have been used in the
relevant literature to compare diffusion processes across different
settings and technologies \citep{cherp2021national, woo2017exploring} as
they can enable more insightful comparisons than, for example, the
average growth over the 1996-2012 study period.

The map illustrates a clear high concentration of fast adopting LAD in
the South and East of England, as expected. There are also some
concentrations in Wales, Scotland and the North West of England. At a
more detailed level, there are some expected cases of LAD with relevant
industrial backgrounds delineated as fast: the City of London, a
world-renowned cluster of finance industries \citep{cook2007role}, and
Reading, a town with high-tech service industries in proximity to London
and its main airport, Heathrow \citep{england2005polynet}. However,
there are also LAD which were expected to appear as fast -- e.g.~Hackney
in central London and Bristol, a well-established creative cluster
\citep{oatley1999cultural, bassett2002cultural} -- but are delineated as
slow. Similar to \citet{beardsell1999spatial}, who analysed the spatial
evolution of computer industry employment across 317 US metro areas,
spatial heterogeneity is evident in the diffusion speed when the scale
of analysis is as detailed as the LAD.

Importantly, there is obvious spatial clustering of fast and slow LAD.
Out of the ten fastest LAD based on when they reached their inflection
point (see detailed Table \ref{table.s.lads} in the Appendix), nine were
located in the South East of England and London and one in Cambridge.
The above observations remain almost unchanged when the analysis
includes websites with up to 10 unique postcodes -- see Figure
\ref{s_map10} in the Appendix -- demonstrating the robustness of the
analysis. Given (i) the expectation that early adopters of new
technologies are often rewarded
\citep{rogers2010diffusion, griliches1957, ding2010modeling}, and (ii)
the previous findings illustrating the long term and sustained
productivity benefit for the early adopters of Web in the UK
\citep{tranosuk}, the observed spatial clustering of LAD with fast
diffusion rates can further reinforce existing and rooted patterns of
unequal economic productivity in the UK \citep{mccann2022productivity}.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map.png}

}

\caption{\label{s_map}LAD Web diffusion speed, 1996-2012}

\end{figure}%

To reveal patterns of early adoption and leapfrogging, the stability and
volatility of the above patterns is analysed. Figure \ref{rank} plots
the relative ranking of LAD based on website density at the beginning
and the end of the study period. To decrease noise, the average ranking
of 1996-1998 and 2010-2012 is plotted instead of the individual years.
Each line represents a single LAD and the colours depict the
5\textsuperscript{th} and 95\textsuperscript{th} percentile of the
absolute difference of the LAD ranking between 1996-1998 and 2010-2012.
In other words, the 5\textsuperscript{th} percentile includes the LAD,
the relative position of which remained unchanged between the beginning
and the end of the study period, both at the top and at the bottom of
the hierarchy (small absolute difference in ranking). The 95th
percentile includes LAD that either increased or decreased a lot their
relative positions over time (high absolute difference in ranking).
Thus, Figure \ref{rank} reveals cases of \emph{leapfrogging} since the
LAD at the top-right part of the figure managed to jump at the top of
the hierarchy at the end of the study period.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av.png}

}

\caption{\label{rank}Dynamics of Web diffusion based on LAD rankings}

\end{figure}%

To further analyse the above, Table \ref{table.late} juxtaposes the
volatility of the LAD Web diffusion against the early conditions as it
reports the \(t_0\) estimate (inflection point) for the leapfrogging LAD
identified in Figure \ref{rank}. Indeed, more than half of the LAD that
drastically increased their relative position at the end of the study
period reached the inflection point \(t_0\) later than the country in
total and, therefore, the Web diffusion speed has been relatively slow
for them. The fact that these LAD experienced relatively faster
diffusion rates at the end of the study period, which is aligned with
the maturity period of the web, can lead to potential positive economic
gains for these LAD. There is extensive literature describing the
benefits of such technological leapfrogging and the underpinning
argument is that latecomers can adopt and benefit from new technologies
that have been developed elsewhere without incurring the hefty initial
R\&D costs \citep{teece2008firm}.

Based on Figure \ref{rank} and Table \ref{table.late}, the opportunity
to catch up that \citet{perkins2005international} illustrated at the
country level, is also visible at the much more detailed LAD scale.
Although the relevant literature did not pay attention to cities and
regions \citep{yu2018sustainability}, the above finding advocates
towards further exploring leapfrogging effects at such a granular scale.
As previously, this is consistent with including websites with up to 10
postcodes (see Figure \ref{rank10}).

All in all, the diffusion of the Web at the detailed scale of LAD is
characterised by spatial heterogeneity, well-established spatial
patterns and both stability and volatility. Although the relevant
literature neglected local scales, the above findings illustrated that
such local dynamics exists and worth further consideration and
investigation especially regarding potential local economic effects.

\begin{longtable}[]{@{}lcc@{}}
\caption{Late adopters\label{table.late}}\tabularnewline
\toprule\noalign{}
LAD & \(t_{0}\) estimate & Diffusion speed \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
LAD & \(t_{0}\) estimate & Diffusion speed \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Halton & 2007.1 & slow \\
Blackpool & 2002.4 & fast \\
Rutland & 2006.5 & slow \\
Torbay & 2002.2 & fast \\
Isles of Scilly & 2012.2 & slow \\
North Devon & 2002.9 & fast \\
Gosport & 2003.8 & slow \\
Thanet & 2004.0 & slow \\
Na h-Eileanan Siar & NA & NA \\
Conwy & 2002.6 & fast \\
\end{longtable}

\section{Conclusions}\label{sec-conclusions}

The paper offers an economic geography study of the diffusion of a new
technology -- the Web. Contrary to previous studies, it does so by
adopting (i) a high level of spatial granularity, and (ii) a long study
period that captures the very early stages of this technology until its
maturity. This is the first time that the importance of granular spatial
mechanisms for the diffusion of a new technology is exposed.
Importantly, the empirical framework and the data employed here allows
to observe the active engagement with the Web that is creating and
maintaining a commercial website instead of the more passive action of
browsing webpages and accessing the Internet.

Specifically, two hypotheses are tested. Firstly, whether the three
diffusion mechanisms namely distance, urban hierarchy and the S-shaped
pattern of the cumulative level of adoption shaped the diffusion of the
commercial Web in the UK at granular geographical scales. The analysis
firstly illustrates how well the S-pattern of cumulative adoption fits
at the local scale and offers some descriptive insights about the
importance of the distance and hierarchical diffusion mechanisms. Then,
the modelling exercise reveals the remarkably good predictive capacity
of these mechanisms at granular geographical scales. The modelling
framework exposes that the relative importance of these spatial
mechanisms differ across different scales. It also reveals spatial
heterogeneity. While for some UK regions the importance of the spatial
diffusion mechanisms are very similar to the rest of the country, other
regions diverge significantly from the country average.

To do the above, a robust and novel ML modelling framework was designed
to predict website diffusion over space and time using as predictors the
spatial diffusion mechanisms. Space- and time-sensitive CV was designed
for the model to avoid overfitting. Importantly, the results are robust
against different specifications and subsets of the data. This is the
first time that the power of these mechanisms in predicting the
diffusion of a new technology was empirically tested, importantly, at
very fine scales.

The second hypothesis assessed whether the diffusion of the web is
shaped by both early adoption and latecomer effects. On the one hand,
the analysis highlighted the heterogeneity of adoption rates between LAD
as well as the spatial clustering of LAD with similar adoption rates.
Importantly, the observed spatial clustering of LAD with fast diffusion
rates can reinforce existing and rooted patterns of unequal economic
productivity in the UK \citep{mccann2022productivity}. Previous research
had already indicated the long-term productivity gains of the early
adoption of such technologies \citep{tranosuk}. On the other hand, the
analysis also exposed instances of remarkable stability as well as
volatility of the Web adoption, with the latter being indicative of
leapfrogging. While the literature had discussed the benefits of
technological leapfrogging at the national scale
\citep{teece2008firm, perkins2005international}, local scales had
escaped such studies \citep{yu2018sustainability}. Thus the leapfrogging
evidence presented here advocate towards further exploring such effects
at granular scales.

The paper offers a number of different contributions. Firstly, it
assesses, for the first time, the role of these spatial diffusion
mechanisms at such granular spatial scales. While previous studies had
highlighted these mechanisms
\citep{hagerstrand1968innovation, rogers2010diffusion, grubler1990rise},
their role had never been tested before at such granular scale. The
findings talk directly to this literature
\citep[e.g.][]{beardsell1999spatial, haller2011determinants, lengyel2020role}.
More broadly, the paper offers a robust methodological framework, which
utilises novel data and modern ML algorithms and techniques to model the
diffusion of a new technology across space and time. It directly
addresses the data problem which prevented economic geographers in
engaging with diffusion studies
\citep{iso2005, kemeny2011international, zook2022mapping}. Thus, the
paper seeks to renew economic geography's interest in the diffusion of
new technologies. Current data science advances including the
availability of granular enough data and adequate methods can enable
economic geographers to reveal interesting spatial patterns of diffusion
of new technologies as this paper has demonstrated. This is important as
(i) various stakeholders are interested in predicting localised demand
for new technologies for planning purposes, for example 5G networks
\citep{leibowicz2016representing, meade2021modelling}, and (ii) the
adoption of such digital technologies is associated with local economic
gains
\citep{solow1957technical, aghion1990model, kemeny2011international, tranosuk, capello2024nexus}.
Importantly, although the geographical focus of this paper is the UK,
the empirical framework developed here is easily transferable to other
geographical contexts.

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\section{Appendix}\label{appendix}

\counterwithin{figure}{section}
\counterwithin{table}{section}

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/s_uk_per_firm_10.png}

}

\caption{\label{s_uk10}Cumulative adoption of the Web, UK; up to 10
postcodes per website}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/s/speed_map_10.png}

}

\caption{\label{s_map10}LAD Web diffusion speed, 1996-2012; up to 10
postcodes per website}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/ranks/web_per_firm2000_2012_only0595_av_10.png}

}

\caption{\label{rank10}Dynamics of Web diffusion based on LAD rankings;
up to 10 postcodes per website}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/morani10-1.pdf}

}

\caption{\label{morani10}Website density Moran's I; up to 10 postcodes
per website}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=0.8\textheight]{anonymised_files/figure-pdf/unnamed-chunk-11-1.pdf}

}

\caption{\label{lisa10}\centering Website density LISA maps; up to 10
postcodes per website}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{anonymised_files/figure-pdf/gini10-1.pdf}

}

\caption{\label{gini10}Website density Gini coefficient; up to 10
postcodes per website}

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=0.6\textheight]{anonymised_files/figure-pdf/varimp10-1.pdf}

}

\caption{\label{var.imp10}Variable importance; up to 10 postcodes per
website}

\end{figure}%

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.3186}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2212}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1416}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1062}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0619}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1504}}@{}}
\caption{S-curve estiamtes for LAD\label{table.s.lads}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
LAD
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Region
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(t_0\) estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Diffusion speed
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
LAD
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Region
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(t_0\) estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(R^2\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Diffusion speed
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Horsham & South East & 2001.423 & 0.254 & 0.950 & fast \\
Fareham & South East & 2001.443 & 0.346 & 0.925 & fast \\
Kingston upon Thames & London & 2001.513 & 0.354 & 0.933 & fast \\
Kensington and Chelsea & London & 2001.581 & 0.341 & 0.943 & fast \\
Runnymede & South East & 2001.582 & 0.266 & 0.943 & fast \\
Bracknell Forest & South East & 2001.586 & 0.324 & 0.928 & fast \\
Elmbridge & South East & 2001.656 & 0.333 & 0.932 & fast \\
Reigate and Banstead & South East & 2001.665 & 0.253 & 0.964 & fast \\
South Cambridgeshire & East of England & 2001.680 & 0.433 & 0.905 &
fast \\
Walsall & West Midlands & 2001.715 & 0.413 & 0.901 & fast \\
Surrey Heath & South East & 2001.718 & 0.333 & 0.933 & fast \\
Woking & South East & 2001.759 & 0.292 & 0.951 & fast \\
South Norfolk & East of England & 2001.874 & 0.358 & 0.931 & fast \\
City of London & London & 2001.874 & 0.499 & 0.930 & fast \\
Wokingham & South East & 2001.891 & 0.287 & 0.958 & fast \\
Reading & South East & 2001.900 & 0.318 & 0.948 & fast \\
Sevenoaks & South East & 2001.904 & 0.407 & 0.926 & fast \\
Huntingdonshire & East of England & 2001.912 & 0.402 & 0.925 & fast \\
Pendle & North West & 2001.924 & 0.407 & 0.923 & fast \\
St Albans & East of England & 2001.933 & 0.396 & 0.927 & fast \\
Perth and Kinross & NA & 2001.943 & 0.228 & 0.970 & fast \\
Bromley & London & 2001.945 & 0.396 & 0.919 & fast \\
Rushmoor & South East & 2001.962 & 0.518 & 0.912 & fast \\
Powys & NA & 2001.969 & 0.278 & 0.963 & fast \\
Swindon & South West & 2001.996 & 0.390 & 0.916 & fast \\
Chelmsford & East of England & 2002.008 & 0.545 & 0.911 & fast \\
Crawley & South East & 2002.008 & 0.379 & 0.949 & fast \\
Dumfries and Galloway & NA & 2002.012 & 0.251 & 0.967 & fast \\
Inverclyde & NA & 2002.030 & 0.330 & 0.915 & fast \\
North Hertfordshire & East of England & 2002.031 & 0.475 & 0.920 &
fast \\
Moray & NA & 2002.050 & 0.358 & 0.947 & fast \\
Orkney Islands & NA & 2002.070 & 0.283 & 0.958 & fast \\
Fife & NA & 2002.070 & 0.363 & 0.945 & fast \\
Mole Valley & South East & 2002.071 & 0.460 & 0.925 & fast \\
Guildford & South East & 2002.113 & 0.413 & 0.941 & fast \\
Buckinghamshire & South East & 2002.114 & 0.390 & 0.942 & fast \\
Cotswold & South West & 2002.116 & 0.375 & 0.946 & fast \\
Shetland Islands & NA & 2002.120 & 0.332 & 0.949 & fast \\
South Oxfordshire & South East & 2002.146 & 0.367 & 0.950 & fast \\
Watford & East of England & 2002.157 & 0.451 & 0.929 & fast \\
Aberdeenshire & NA & 2002.168 & 0.416 & 0.931 & fast \\
Southampton & South East & 2002.189 & 0.472 & 0.922 & fast \\
Mid Suffolk & East of England & 2002.196 & 0.442 & 0.922 & fast \\
Eden & North West & 2002.196 & 0.334 & 0.949 & fast \\
South Lakeland & North West & 2002.197 & 0.238 & 0.974 & fast \\
Stoke-on-Trent & West Midlands & 2002.203 & 0.419 & 0.927 & fast \\
Babergh & East of England & 2002.209 & 0.449 & 0.927 & fast \\
Bexley & London & 2002.213 & 0.493 & 0.909 & fast \\
Torbay & South West & 2002.215 & 0.374 & 0.942 & fast \\
Brentwood & East of England & 2002.219 & 0.336 & 0.936 & fast \\
Ceredigion & NA & 2002.227 & 0.241 & 0.971 & fast \\
Spelthorne & South East & 2002.232 & 0.451 & 0.933 & fast \\
Bath and North East Somerset & South West & 2002.238 & 0.402 & 0.941 &
fast \\
Falkirk & NA & 2002.245 & 0.367 & 0.938 & fast \\
Broxtowe & East Midlands & 2002.258 & 0.435 & 0.932 & fast \\
East Hertfordshire & East of England & 2002.261 & 0.453 & 0.928 &
fast \\
Stirling & NA & 2002.268 & 0.251 & 0.975 & fast \\
Ealing & London & 2002.294 & 0.523 & 0.922 & fast \\
Mid Sussex & South East & 2002.295 & 0.453 & 0.936 & fast \\
Barrow-in-Furness & North West & 2002.303 & 0.284 & 0.959 & fast \\
West Oxfordshire & South East & 2002.322 & 0.467 & 0.929 & fast \\
Sutton & London & 2002.326 & 0.544 & 0.911 & fast \\
Portsmouth & South East & 2002.356 & 0.479 & 0.921 & fast \\
Great Yarmouth & East of England & 2002.369 & 0.265 & 0.968 & fast \\
East Hampshire & South East & 2002.370 & 0.474 & 0.931 & fast \\
Wyre Forest & West Midlands & 2002.376 & 0.561 & 0.916 & fast \\
Angus & NA & 2002.378 & 0.396 & 0.939 & fast \\
Argyll and Bute & NA & 2002.427 & 0.321 & 0.962 & fast \\
Scottish Borders & NA & 2002.428 & 0.298 & 0.962 & fast \\
East Suffolk & East of England & 2002.431 & 0.533 & 0.919 & fast \\
Blackpool & North West & 2002.443 & 0.294 & 0.962 & fast \\
Cherwell & South East & 2002.448 & 0.533 & 0.923 & fast \\
Wandsworth & London & 2002.449 & 0.570 & 0.921 & fast \\
Tendring & East of England & 2002.456 & 0.456 & 0.935 & fast \\
Westminster & London & 2002.459 & 0.437 & 0.952 & fast \\
Richmond upon Thames & London & 2002.462 & 0.389 & 0.956 & fast \\
Test Valley & South East & 2002.463 & 0.553 & 0.920 & fast \\
Isle of Wight & South East & 2002.478 & 0.377 & 0.953 & fast \\
Vale of White Horse & South East & 2002.496 & 0.408 & 0.949 & fast \\
Croydon & London & 2002.510 & 0.504 & 0.927 & fast \\
Hounslow & London & 2002.521 & 0.426 & 0.952 & fast \\
Copeland & North West & 2002.524 & 0.237 & 0.971 & fast \\
Wiltshire & South West & 2002.528 & 0.450 & 0.940 & fast \\
Gwynedd & NA & 2002.536 & 0.372 & 0.954 & fast \\
Hillingdon & London & 2002.560 & 0.470 & 0.940 & fast \\
Conwy & NA & 2002.564 & 0.372 & 0.954 & fast \\
Highland & NA & 2002.566 & 0.270 & 0.972 & fast \\
Basingstoke and Deane & South East & 2002.572 & 0.461 & 0.947 & fast \\
Harrow & London & 2002.587 & 0.509 & 0.936 & fast \\
West Berkshire & South East & 2002.595 & 0.696 & 0.902 & fast \\
North Lincolnshire & Yorkshire and The Humber & 2002.596 & 0.501 & 0.933
& fast \\
East Dunbartonshire & NA & 2002.598 & 0.528 & 0.925 & fast \\
Bradford & Yorkshire and The Humber & 2002.598 & 0.613 & 0.916 & fast \\
Waverley & South East & 2002.606 & 0.512 & 0.928 & fast \\
Stroud & South West & 2002.608 & 0.559 & 0.916 & fast \\
Herefordshire, County of & West Midlands & 2002.622 & 0.388 & 0.956 &
fast \\
North Norfolk & East of England & 2002.624 & 0.422 & 0.945 & fast \\
Nottingham & East Midlands & 2002.630 & 0.531 & 0.936 & fast \\
Castle Point & East of England & 2002.645 & 0.470 & 0.936 & fast \\
North Warwickshire & West Midlands & 2002.657 & 0.380 & 0.955 & fast \\
Gravesham & South East & 2002.680 & 0.473 & 0.933 & fast \\
Lancaster & North West & 2002.692 & 0.336 & 0.963 & fast \\
West Suffolk & East of England & 2002.699 & 0.536 & 0.930 & fast \\
Stafford & West Midlands & 2002.719 & 0.499 & 0.933 & fast \\
Tunbridge Wells & South East & 2002.720 & 0.401 & 0.958 & fast \\
Oxford & South East & 2002.723 & 0.583 & 0.934 & fast \\
East Lothian & NA & 2002.731 & 0.537 & 0.927 & fast \\
Redcar and Cleveland & North East & 2002.750 & 0.509 & 0.929 & fast \\
Tamworth & West Midlands & 2002.755 & 0.551 & 0.929 & fast \\
South Hams & South West & 2002.773 & 0.422 & 0.952 & fast \\
Allerdale & North West & 2002.784 & 0.376 & 0.958 & fast \\
Central Bedfordshire & East of England & 2002.825 & 0.620 & 0.921 &
fast \\
East Cambridgeshire & East of England & 2002.843 & 0.636 & 0.927 &
fast \\
Havant & South East & 2002.853 & 0.541 & 0.932 & fast \\
South Lanarkshire & NA & 2002.874 & 0.486 & 0.938 & fast \\
Stratford-on-Avon & West Midlands & 2002.879 & 0.470 & 0.945 & fast \\
Arun & South East & 2002.882 & 0.526 & 0.938 & fast \\
Fenland & East of England & 2002.891 & 0.632 & 0.918 & fast \\
Monmouthshire & NA & 2002.893 & 0.501 & 0.935 & fast \\
Denbighshire & NA & 2002.901 & 0.457 & 0.947 & fast \\
Southend-on-Sea & East of England & 2002.903 & 0.551 & 0.932 & fast \\
High Peak & East Midlands & 2002.903 & 0.488 & 0.946 & fast \\
Pembrokeshire & NA & 2002.905 & 0.387 & 0.958 & fast \\
Malvern Hills & West Midlands & 2002.907 & 0.524 & 0.940 & fast \\
Camden & London & 2002.915 & 0.599 & 0.941 & fast \\
North Devon & South West & 2002.926 & 0.426 & 0.950 & fast \\
Wyre & North West & 2002.927 & 0.553 & 0.927 & fast \\
Darlington & North East & 2002.933 & 0.615 & 0.929 & fast \\
Swale & South East & 2002.940 & 0.447 & 0.948 & fast \\
Welwyn Hatfield & East of England & 2002.951 & 0.628 & 0.922 & fast \\
Mid Devon & South West & 2002.960 & 0.485 & 0.946 & fast \\
Lewes & South East & 2002.964 & 0.507 & 0.945 & fast \\
Dorset & South West & 2002.968 & 0.407 & 0.958 & fast \\
Brighton and Hove & South East & 2002.976 & 0.578 & 0.934 & fast \\
Braintree & East of England & 2002.978 & 0.534 & 0.938 & fast \\
Epsom and Ewell & South East & 2002.979 & 0.746 & 0.908 & fast \\
Sefton & North West & 2002.980 & 0.616 & 0.928 & fast \\
Eastleigh & South East & 2002.992 & 0.595 & 0.929 & fast \\
Craven & Yorkshire and The Humber & 2002.993 & 0.379 & 0.965 & fast \\
Hammersmith and Fulham & London & 2002.997 & 0.585 & 0.944 & fast \\
Greenwich & London & 2003.014 & 0.571 & 0.930 & slow \\
Cheltenham & South West & 2003.025 & 0.602 & 0.937 & slow \\
Worcester & West Midlands & 2003.041 & 0.659 & 0.920 & slow \\
Tower Hamlets & London & 2003.043 & 0.660 & 0.926 & slow \\
New Forest & South East & 2003.046 & 0.554 & 0.941 & slow \\
East Renfrewshire & NA & 2003.046 & 0.739 & 0.900 & slow \\
Fylde & North West & 2003.056 & 0.501 & 0.946 & slow \\
Carmarthenshire & NA & 2003.067 & 0.390 & 0.958 & slow \\
West Lancashire & North West & 2003.076 & 0.547 & 0.937 & slow \\
Hastings & South East & 2003.086 & 0.685 & 0.911 & slow \\
Carlisle & North West & 2003.111 & 0.541 & 0.934 & slow \\
Wychavon & West Midlands & 2003.115 & 0.570 & 0.935 & slow \\
Isle of Anglesey & NA & 2003.144 & 0.403 & 0.960 & slow \\
Dudley & West Midlands & 2003.150 & 0.678 & 0.925 & slow \\
Ryedale & Yorkshire and The Humber & 2003.152 & 0.380 & 0.961 & slow \\
Rother & South East & 2003.162 & 0.366 & 0.966 & slow \\
Leicester & East Midlands & 2003.167 & 0.686 & 0.924 & slow \\
Breckland & East of England & 2003.182 & 0.567 & 0.938 & slow \\
Tandridge & South East & 2003.190 & 0.747 & 0.918 & slow \\
Shropshire & West Midlands & 2003.206 & 0.621 & 0.928 & slow \\
West Devon & South West & 2003.228 & 0.489 & 0.949 & slow \\
Milton Keynes & South East & 2003.231 & 0.700 & 0.929 & slow \\
Sandwell & West Midlands & 2003.234 & 0.760 & 0.916 & slow \\
Winchester & South East & 2003.234 & 0.443 & 0.956 & slow \\
Redditch & West Midlands & 2003.238 & 0.706 & 0.923 & slow \\
Rugby & West Midlands & 2003.265 & 0.709 & 0.922 & slow \\
North Somerset & South West & 2003.278 & 0.558 & 0.939 & slow \\
West Lindsey & East Midlands & 2003.285 & 0.674 & 0.924 & slow \\
Richmondshire & Yorkshire and The Humber & 2003.286 & 0.425 & 0.958 &
slow \\
Melton & East Midlands & 2003.292 & 0.534 & 0.946 & slow \\
Chichester & South East & 2003.295 & 0.582 & 0.942 & slow \\
Torridge & South West & 2003.301 & 0.532 & 0.942 & slow \\
Dundee City & NA & 2003.303 & 0.568 & 0.939 & slow \\
South Ayrshire & NA & 2003.303 & 0.626 & 0.931 & slow \\
East Devon & South West & 2003.308 & 0.492 & 0.947 & slow \\
Bolton & North West & 2003.323 & 0.774 & 0.913 & slow \\
Hertsmere & East of England & 2003.326 & 0.726 & 0.924 & slow \\
Bournemouth, Christchurch and Poole & South West & 2003.347 & 0.662 &
0.936 & slow \\
Maldon & East of England & 2003.349 & 0.495 & 0.951 & slow \\
Barnet & London & 2003.352 & 0.766 & 0.919 & slow \\
Brent & London & 2003.389 & 0.762 & 0.925 & slow \\
Midlothian & NA & 2003.413 & 0.732 & 0.921 & slow \\
Boston & East Midlands & 2003.416 & 0.593 & 0.933 & slow \\
Scarborough & Yorkshire and The Humber & 2003.421 & 0.431 & 0.963 &
slow \\
Kirklees & Yorkshire and The Humber & 2003.435 & 0.725 & 0.925 & slow \\
Tameside & North West & 2003.441 & 0.819 & 0.907 & slow \\
Newport & NA & 2003.449 & 0.635 & 0.930 & slow \\
Glasgow City & NA & 2003.451 & 0.759 & 0.922 & slow \\
Clackmannanshire & NA & 2003.454 & 0.585 & 0.936 & slow \\
Three Rivers & East of England & 2003.458 & 0.628 & 0.947 & slow \\
Harborough & East Midlands & 2003.488 & 0.549 & 0.950 & slow \\
Rochdale & North West & 2003.496 & 0.685 & 0.937 & slow \\
Tonbridge and Malling & South East & 2003.497 & 0.710 & 0.927 & slow \\
Somerset West and Taunton & South West & 2003.502 & 0.491 & 0.955 &
slow \\
Wrexham & NA & 2003.512 & 0.553 & 0.942 & slow \\
Worthing & South East & 2003.522 & 0.622 & 0.938 & slow \\
Tewkesbury & South West & 2003.540 & 0.580 & 0.946 & slow \\
City of Edinburgh & NA & 2003.541 & 0.685 & 0.941 & slow \\
Coventry & West Midlands & 2003.592 & 0.564 & 0.948 & slow \\
Northumberland & North East & 2003.607 & 0.508 & 0.954 & slow \\
Mid and East Antrim & NA & 2003.609 & 0.893 & 0.900 & slow \\
Wigan & North West & 2003.617 & 0.755 & 0.929 & slow \\
Lichfield & West Midlands & 2003.629 & 0.625 & 0.939 & slow \\
Ashford & South East & 2003.633 & 0.535 & 0.944 & slow \\
Cornwall & South West & 2003.638 & 0.443 & 0.963 & slow \\
Mendip & South West & 2003.652 & 0.505 & 0.955 & slow \\
Stevenage & East of England & 2003.665 & 0.525 & 0.958 & slow \\
Blaenau Gwent & NA & 2003.666 & 0.643 & 0.933 & slow \\
Lincoln & East Midlands & 2003.666 & 0.634 & 0.940 & slow \\
South Somerset & South West & 2003.676 & 0.553 & 0.948 & slow \\
Charnwood & East Midlands & 2003.677 & 0.699 & 0.934 & slow \\
North Ayrshire & NA & 2003.682 & 0.607 & 0.942 & slow \\
Folkestone and Hythe & South East & 2003.684 & 0.792 & 0.905 & slow \\
Eastbourne & South East & 2003.684 & 0.697 & 0.930 & slow \\
Blackburn with Darwen & North West & 2003.685 & 0.680 & 0.938 & slow \\
Dover & South East & 2003.699 & 0.574 & 0.948 & slow \\
Adur & South East & 2003.735 & 0.551 & 0.954 & slow \\
Haringey & London & 2003.739 & 0.667 & 0.943 & slow \\
Gosport & South East & 2003.751 & 0.600 & 0.934 & slow \\
Trafford & North West & 2003.752 & 0.640 & 0.949 & slow \\
Norwich & East of England & 2003.756 & 0.742 & 0.939 & slow \\
West Lothian & NA & 2003.758 & 0.652 & 0.943 & slow \\
North East Lincolnshire & Yorkshire and The Humber & 2003.811 & 0.858 &
0.920 & slow \\
Bristol, City of & South West & 2003.828 & 0.831 & 0.928 & slow \\
Rushcliffe & East Midlands & 2003.839 & 0.591 & 0.954 & slow \\
Hinckley and Bosworth & East Midlands & 2003.857 & 0.749 & 0.934 &
slow \\
Torfaen & NA & 2003.861 & 0.841 & 0.922 & slow \\
Swansea & NA & 2003.875 & 0.650 & 0.947 & slow \\
Wolverhampton & West Midlands & 2003.882 & 0.793 & 0.931 & slow \\
Kingston upon Hull, City of & Yorkshire and The Humber & 2003.885 &
0.757 & 0.936 & slow \\
Calderdale & Yorkshire and The Humber & 2003.889 & 0.714 & 0.939 &
slow \\
North Northamptonshire & East Midlands & 2003.902 & 0.746 & 0.935 &
slow \\
Stockton-on-Tees & North East & 2003.902 & 0.955 & 0.916 & slow \\
Wealden & South East & 2003.906 & 0.736 & 0.938 & slow \\
Thanet & South East & 2003.953 & 0.667 & 0.947 & slow \\
Broadland & East of England & 2003.962 & 0.727 & 0.938 & slow \\
West Northamptonshire & East Midlands & 2003.972 & 0.736 & 0.940 &
slow \\
North Kesteven & East Midlands & 2003.976 & 0.633 & 0.942 & slow \\
Neath Port Talbot & NA & 2003.981 & 0.767 & 0.933 & slow \\
Rhondda Cynon Taf & NA & 2003.991 & 0.769 & 0.925 & slow \\
Enfield & London & 2003.992 & 0.588 & 0.954 & slow \\
East Lindsey & East Midlands & 2004.010 & 0.577 & 0.952 & slow \\
Teignbridge & South West & 2004.015 & 0.681 & 0.941 & slow \\
York & Yorkshire and The Humber & 2004.025 & 0.739 & 0.944 & slow \\
Harlow & East of England & 2004.030 & 0.803 & 0.936 & slow \\
Vale of Glamorgan & NA & 2004.033 & 0.592 & 0.951 & slow \\
Newark and Sherwood & East Midlands & 2004.034 & 0.668 & 0.946 & slow \\
Merton & London & 2004.041 & 0.932 & 0.922 & slow \\
Hambleton & Yorkshire and The Humber & 2004.065 & 0.636 & 0.953 &
slow \\
Bedford & East of England & 2004.068 & 0.852 & 0.933 & slow \\
Stockport & North West & 2004.075 & 0.757 & 0.942 & slow \\
South Ribble & North West & 2004.126 & 0.674 & 0.948 & slow \\
Uttlesford & East of England & 2004.144 & 0.905 & 0.926 & slow \\
Rossendale & North West & 2004.160 & 0.976 & 0.923 & slow \\
Oldham & North West & 2004.162 & 0.801 & 0.938 & slow \\
Sheffield & Yorkshire and The Humber & 2004.189 & 0.770 & 0.944 &
slow \\
Ribble Valley & North West & 2004.204 & 0.718 & 0.945 & slow \\
Maidstone & South East & 2004.222 & 0.854 & 0.932 & slow \\
Cheshire West and Chester & North West & 2004.234 & 0.742 & 0.947 &
slow \\
Thurrock & East of England & 2004.241 & 0.841 & 0.924 & slow \\
Telford and Wrekin & West Midlands & 2004.241 & 1.024 & 0.917 & slow \\
Cardiff & NA & 2004.246 & 0.823 & 0.937 & slow \\
County Durham & North East & 2004.250 & 0.626 & 0.953 & slow \\
Plymouth & South West & 2004.252 & 0.785 & 0.934 & slow \\
Bridgend & NA & 2004.256 & 0.547 & 0.956 & slow \\
Rotherham & Yorkshire and The Humber & 2004.273 & 0.722 & 0.938 &
slow \\
Hyndburn & North West & 2004.298 & 0.881 & 0.932 & slow \\
Newcastle upon Tyne & North East & 2004.305 & 0.906 & 0.932 & slow \\
Wakefield & Yorkshire and The Humber & 2004.314 & 0.905 & 0.923 &
slow \\
Havering & London & 2004.345 & 0.822 & 0.938 & slow \\
South Kesteven & East Midlands & 2004.353 & 1.026 & 0.915 & slow \\
Canterbury & South East & 2004.354 & 0.602 & 0.960 & slow \\
Broxbourne & East of England & 2004.363 & 0.817 & 0.944 & slow \\
Peterborough & East of England & 2004.382 & 0.937 & 0.932 & slow \\
Redbridge & London & 2004.411 & 1.028 & 0.925 & slow \\
Bury & North West & 2004.431 & 0.735 & 0.953 & slow \\
Harrogate & Yorkshire and The Humber & 2004.433 & 0.971 & 0.927 &
slow \\
Birmingham & West Midlands & 2004.435 & 1.182 & 0.912 & slow \\
South Tyneside & North East & 2004.446 & 0.636 & 0.948 & slow \\
Chorley & North West & 2004.462 & 1.093 & 0.913 & slow \\
North East Derbyshire & East Midlands & 2004.483 & 0.703 & 0.953 &
slow \\
Lambeth & London & 2004.485 & 0.958 & 0.937 & slow \\
Doncaster & Yorkshire and The Humber & 2004.496 & 0.672 & 0.953 &
slow \\
South Gloucestershire & South West & 2004.499 & 0.881 & 0.937 & slow \\
Caerphilly & NA & 2004.499 & 0.819 & 0.938 & slow \\
East Staffordshire & West Midlands & 2004.501 & 1.132 & 0.910 & slow \\
North Tyneside & North East & 2004.509 & 0.864 & 0.933 & slow \\
South Holland & East Midlands & 2004.534 & 0.851 & 0.941 & slow \\
Gateshead & North East & 2004.535 & 0.869 & 0.934 & slow \\
St.~Helens & North West & 2004.542 & 1.097 & 0.923 & slow \\
Flintshire & NA & 2004.549 & 0.764 & 0.951 & slow \\
Selby & Yorkshire and The Humber & 2004.549 & 0.716 & 0.947 & slow \\
Belfast & NA & 2004.556 & 1.332 & 0.906 & slow \\
Sedgemoor & South West & 2004.613 & 0.540 & 0.966 & slow \\
Exeter & South West & 2004.617 & 0.964 & 0.936 & slow \\
South Derbyshire & East Midlands & 2004.629 & 0.934 & 0.938 & slow \\
Derbyshire Dales & East Midlands & 2004.668 & 0.648 & 0.963 & slow \\
East Riding of Yorkshire & Yorkshire and The Humber & 2004.683 & 0.728 &
0.955 & slow \\
Basildon & East of England & 2004.691 & 1.067 & 0.925 & slow \\
Barking and Dagenham & London & 2004.800 & 1.198 & 0.916 & slow \\
Causeway Coast and Glens & NA & 2004.834 & 1.024 & 0.928 & slow \\
North Lanarkshire & NA & 2004.839 & 0.878 & 0.945 & slow \\
Chesterfield & East Midlands & 2004.887 & 1.173 & 0.927 & slow \\
Newcastle-under-Lyme & West Midlands & 2004.933 & 1.172 & 0.925 &
slow \\
East Ayrshire & NA & 2004.934 & 1.280 & 0.903 & slow \\
Nuneaton and Bedworth & West Midlands & 2004.956 & 0.950 & 0.941 &
slow \\
Blaby & East Midlands & 2004.959 & 1.059 & 0.929 & slow \\
Leeds & Yorkshire and The Humber & 2004.979 & 0.836 & 0.956 & slow \\
Erewash & East Midlands & 2004.987 & 0.939 & 0.944 & slow \\
Bassetlaw & East Midlands & 2004.998 & 0.982 & 0.944 & slow \\
Hackney & London & 2005.026 & 1.157 & 0.934 & slow \\
Barnsley & Yorkshire and The Humber & 2005.055 & 0.912 & 0.942 & slow \\
Amber Valley & East Midlands & 2005.105 & 1.017 & 0.941 & slow \\
Colchester & East of England & 2005.144 & 1.250 & 0.927 & slow \\
Ashfield & East Midlands & 2005.176 & 1.073 & 0.942 & slow \\
Wirral & North West & 2005.184 & 1.084 & 0.939 & slow \\
Knowsley & North West & 2005.186 & 1.160 & 0.923 & slow \\
Windsor and Maidenhead & South East & 2005.195 & 1.746 & 0.902 & slow \\
Bromsgrove & West Midlands & 2005.494 & 1.144 & 0.941 & slow \\
Liverpool & North West & 2005.526 & 1.276 & 0.936 & slow \\
Staffordshire Moorlands & West Midlands & 2005.583 & 1.426 & 0.924 &
slow \\
Islington & London & 2005.970 & 1.223 & 0.955 & slow \\
Warrington & North West & 2006.015 & 1.387 & 0.942 & slow \\
Newham & London & 2006.078 & 1.908 & 0.907 & slow \\
Bolsover & East Midlands & 2006.188 & 1.143 & 0.957 & slow \\
Lisburn and Castlereagh & NA & 2006.193 & 1.970 & 0.903 & slow \\
Manchester & North West & 2006.449 & 1.434 & 0.950 & slow \\
Sunderland & North East & 2006.479 & 1.679 & 0.928 & slow \\
South Staffordshire & West Midlands & 2006.490 & 1.837 & 0.922 & slow \\
Rutland & East Midlands & 2006.529 & 0.986 & 0.967 & slow \\
Cannock Chase & West Midlands & 2006.700 & 2.231 & 0.917 & slow \\
Mansfield & East Midlands & 2006.889 & 1.579 & 0.946 & slow \\
Ards and North Down & NA & 2006.947 & 2.485 & 0.907 & slow \\
Halton & North West & 2007.069 & 0.932 & 0.967 & slow \\
Burnley & North West & 2007.372 & 2.286 & 0.919 & slow \\
Middlesbrough & North East & 2010.074 & 5.297 & 0.900 & slow \\
Isles of Scilly & South West & 2012.163 & 4.257 & 0.953 & slow \\
\end{longtable}

\begin{longtable}[]{@{}lrrrr@{}}
\caption{Regional differences; up to 10 postcodes per
website\label{table.regions.10}}\tabularnewline
\toprule\noalign{}
Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Region & \(R^2\) LAD & Rank LAD & \(R^2\) OA & Rank OA \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Yorkshire and The Humber & 0.864 & 1 & 0.191 & 5 \\
West Midlands & 0.860 & 2 & 0.148 & 8 \\
South East & 0.848 & 3 & 0.171 & 7 \\
North West & 0.827 & 4 & 0.221 & 1 \\
London & 0.826 & 5 & 0.118 & 9 \\
South West & 0.801 & 6 & 0.194 & 4 \\
Wales & 0.746 & 7 & 0.114 & 11 \\
East Midlands & 0.732 & 8 & 0.117 & 10 \\
East of England & 0.731 & 9 & 0.178 & 6 \\
North East & 0.695 & 10 & 0.215 & 2 \\
Scotland & 0.588 & 11 & 0.052 & 12 \\
Nortern Ireland & 0.488 & 12 & 0.197 & 3 \\
\end{longtable}

\begin{longtable}[]{@{}llll@{}}
\caption{Model metrics for growth
model}\label{tbl-model-growth-metrics}\tabularnewline
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& \(RMSE\) & \(R^{2}\) & \(MAE\) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Local Authorities & 0.286 & 0.506 & 0.223 \\
\end{longtable}

Table~\ref{tbl-model-growth-metrics} presents the results of a Random
Forest model that was trained based on Equation \ref{model.growth} using
all LAD data points. In essence, this is a model similar to the one
estimated based on Equation \ref{eq:model} and presented in
Table~\ref{tbl-model-metrics}, with the main difference being the
dependent variable which is the year growth rate of website density for
each LAD. To avoid cases with LAD have no websites,
\(t \in [2000, 2012]\). The high predictive capability of the model
reflects the robustness of the modelling approach.

\begin{align} \label{model.growth}
Website\,Density\,Growth_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}

\renewcommand\refname{References}
  \bibliography{bibliography}

\end{document}