### 2025-07-29 paper/submission4/title.qmd

---
title: "A multi-scale story of the diffusion of a new technology: the web"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol
        department: School of Geographical Sciences
        address: University Road, BS8 1SS
        city: Bristol
        country: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
    This paper investigates the spatial diffusion of a new technology that is the Web in the UK. It employs novel data and machine learning methods to model the influence of well-established diffusion mechanisms. Contrary to previous studies, it adopts multiple scales, high spatial granularity and a long study period that captures the early stages of the Web until its maturity (1996-2012). Findings reveal the importance of such spatial mechanisms (namely distance, urban hierarchy and the S-shaped pattern of the cumulative level of adoption) even at granular scales. They also highlight spatial heterogeneity and instances of leapfrogging. 
keywords: 
  - diffusion
  - technologies
  - Web
  - digital
date: last-modified
format:
  elsevier-pdf:
    keep-tex: true
    text: |
      \usepackage[demo]{graphicx} % "demo" option just for this example
      \usepackage{subcaption}
    journal:
      name: Journal of Economic Geography
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

# Acknowledgements

Many thanks to Dr Levi Wolf and Professor Dani Arribas-Bel for their constructive 
comments. Also many thanks to the participants of the 6th Global Conference on 
Economic Geography and the Winter Seminar of GfR. Then, many thanks to the 
organisers of the seminar series and the members of the Quantitative Spatial 
Science (QuSS) Research Group at the University of Bristol, the Centre for 
Business Prosperity Research Group at Aston University (Dr Tasos Kitsos), and 
the Centre for Innovation Research (Professor Tom Broekel) at the University of 
Stavanger. Finally, many thanks to the editor and the three anonymous reviewers 
for the very helpful comments and suggestions.

# Funding

No external funding was received for this work.